---
title: "Depredation FSSgams"
author: "Nicole Hamre"
---
# Set up
## Librarys
```{r librarys}
# install.packages('mgcv')
# install.packages('MuMIn')
# install.packages('gamm4')

library(tidyverse)
library(FSSgam)
library(mgcv)
library(MuMIn)
library(gamm4)
library(corrplot)

```
## Directories
```{r dir, echo = FALSE}
w.dir <- "/Users/23088313/Documents/git_repos/Analysis-Hamre-Bioeconomic" # working directory
d.dir <- paste(w.dir, "Ningaloo_RUM/rumIgnore", sep='/') # data directory
m.dir <- paste(w.dir, "Ningaloo_RUM/rumDP/models", sep = '/') # model directory
f.dir <- paste(w.dir, "Ningaloo_RUM/rumFunc", sep = '/')
dpGAM <- paste(m.dir, "dpGAM", sep='/') # depredation gam output directory
crGAM <- paste(m.dir, "crGAM", sep='/') # catch rate gam output directory
```

```{r function}
source(paste(f.dir, "cleaningFunc.R", sep = '/'))
```

## Read data
```{r read data}
# dpdat <- read.csv(paste(d.dir, "dpDat_v1.csv", sep='/'))
# 
# Ning <- read.csv(paste(d.dir, "Ning_v2.csv", sep = '/'))
 
N <- Ning %>% 
  filter(Agreement == "Yes") 
# Extractive data
exdat <- N %>%
  dplyr::filter(Activity == "Fishing")

# Boat based extractive
bf <- N %>% 
  dplyr::filter(SiteType == "Boat" & ActivityType %in% c("Extractive", "Both"))

dpdat <- bf %>% 
  dplyr::filter(FishingType %in% c("Trolling", "Demersal"))

dpdat <- readRDS(paste(d.dir, "dpdat_v1.csv", sep = "/"))

dpdat2 <- dpdat %>% 
  filter(nDP > 21)
```
## Prep data
```{r}
# Making all character vars factors
dpdat <- dpdat %>%
  mutate_if(is.character, as.factor) %>% 
  mutate(facYear = as.factor(facYear))
```

# GAMMs
## Assign 
## Check for correlations
```{r corrolations}
# Need to put this package in the libs
# ASk charlotte if you do this with just continuous variables of if you do factors to
# If you put them all in then stick to teh format below 
# Need to check for correlations above 0.95, any higher should be ruled out.
# Figure out which one gets ruled out

 round(cor(dpdat[,c("numYear","nTimesLast24m","exDecMedianTime","Depth", "DistNearestBRkm", "DistNearestInfra", "UseLat","UseLong", "DecFishingHr","MaxHook","BoatLength",
                    "b_kdens", "b_5km")], use='complete.obs'), 2)

round(cor(dpdat[,c("nDP", "nHooked")], use='complete.obs'), 2)
```

## Assign variables
```{r assign cont}
DPcont <- c("nTimesLast24m","exDecMedianTime",
           "Depth","DistNearestBRkm", "DistNearestInfra", "DecFishingHr","MaxHook","BoatLength", "b_kdens", "b_5km") # No coords in this for transformations - will be added later 

DPfac <- c("Month", "Side", "facYear", "Resident", "Habitat", "LunarPhase", "FishingType") # factors
```

## Checking data
### Removing Nas from data set
```{r distrobution}

dpdat2 <- dpdat %>% 
  mutate(nTimesLast24m = ifelse(is.na(nTimesLast24m), 0, nTimesLast24m)) %>% 
  filter(!(ID %in% c(502, 548, 565))) %>% 
  mutate(MaxHook = ifelse(!is.na(Depth) & Depth > 0 & FishingType %in% "Demersal", Depth, MaxHook)) %>% 
  drop_na(MaxHook) %>% 
  mutate(BoatLength = ifelse(is.na(BoatLength), round(mean(BoatLength, na.rm = T)), BoatLength)) %>% 
  mutate(DecFishingHr = ifelse(is.na(DecFishingHr), round(mean(DecFishingHr, na.rm = T)), DecFishingHr)) %>% 
# turn to decimal subtract /add then turn back  into right format 
  mutate(temp_Start = StartTime) %>% 
  mutate(temp_Start = gsub("[:]", ".", temp_Start)) %>%
  mutate(temp_Start = as.numeric(temp_Start)) %>% 
  mutate(temp_Stop = StopTime) %>% 
  mutate(temp_Stop = gsub("[:]", ".", temp_Stop)) %>% # will substitute all : with . within temp_start
  mutate(temp_Stop = as.numeric(temp_Stop)) %>% 
  mutate(temp_Start = ifelse(is.na(temp_Start), temp_Stop - DecFishingHr, temp_Start)) %>% 
  mutate(temp_Stop = ifelse(is.na(temp_Stop), temp_Start + DecFishingHr, temp_Stop)) %>% 
  rowwise() %>%
  mutate(temp_Med = median(c(temp_Start, temp_Stop))) %>%
  ungroup() %>% 
  mutate(temp_Med = gsub("[.]", ":", temp_Med)) %>% 
  mutate(Resident = as.factor(Resident)) %>% 
  mutate(temp_Med = ifelse(str_length(temp_Med)==2, str_c(temp_Med, ":00"), temp_Med)) %>% # if string is 2 digits long it will ass ":00" to end of colum temp_Med
  mutate(temp_Med = ifelse(str_length(temp_Med)==3, str_c("0", temp_Med, "0"), temp_Med)) %>% # if string is 3 digits long it will add a 0 to the strat and end of whats in temp_Med
  mutate(temp_Med = ifelse(str_length(temp_Med)==4, str_c(temp_Med, "0"), temp_Med)) %>% 
  mutate(exDecMedianTime = ifelse(!is.na(temp_Med), hhmm2dec(temp_Med), exDecMedianTime)) %>% 
  mutate(exDecMedianTime = round(exDecMedianTime, digits = 2)) %>% # rounding to 2 digits after decimal 
  mutate(exDecMedianTime = ifelse(is.na(exDecMedianTime), 12, exDecMedianTime)) %>% 
  mutate(exDecMedianTime = as.numeric(exDecMedianTime)) %>% 
  select(PersonID, nDP, nHooked, nTimesLast24m, exDecMedianTime,
           Depth,DistNearestBRkm, DistNearestInfra, UseLat,
           UseLong, DecFishingHr,MaxHook, BoatLength, b_kdens, b_5km, Month, Side, facYear, Resident, Habitat, LunarPhase, FishingType) %>% 
  mutate(Depth = Depth*-1) %>% 
  mutate(Depth = ifelse(PersonID %in% 190, 0, Depth)) %>% 
  as.data.frame()

# nTimesLast24m_sqrt - assuming in non entries are due to people not entering 0s if people haven't been to Exmouth before - hence filled with 0 
# Depth_logxplusone - removed 3 rows 
# MaxHook - if fishing was demersal made max hook equal to depth
# BoatLength - filled in NAs with average boat length
# DecFishingHr - filled nas with average fishing hours (4)
# exDecMedianTime - if there was a start time i added the average fishing time (4) to get and end time and caculated the median from those times - and vice versa for when theres an end time and no start time. cases where there were neither a start or end time were filled with the average median time (12)
```

### Distribution
Looking at data through various transformations to see which transformation provides the most normal distribution and is homoscedastic. 
```{r distrobution}
# need to run in console
par(mfrow=c(3,4))
for(i in DPcont) {
  x <- dpdat2[,i]
  x = as.numeric(unlist(x))
  hist((x))
  plot((x), main=paste(i))
  hist(sqrt(x))
  plot(sqrt(x))
  hist(log(x+1))
  plot(log(x+1))
  hist((x))
  plot((x), main=paste(i))
  hist((x)^2)
  plot((x)^2)
  hist((x)^3)
  plot((x)^3)
}

# nTimesLast24m: x
# exDecMedianTime: log(x+1)
# Depth: log(x+1)
# DistNearestBR:log(x+1)
# DistNearestInfra: sqrt
# UseLat: x
# UseLong:x
# DecFishingHr:sqrt
# MaxHook: sqrt
# BoatLength: x
# b_kdens: sqrt
# b_5km: log(x+1)


# Transform variables into dataset
dpdat3$exDecMedianTime_sqrt <- sqrt(dpdat3$exDecMedianTime)
dpdat3$Depth_logxplusone <- log(dpdat3$Depth +1)
dpdat3$DistNearestBRkm_logxplusone <- log(dpdat3$DistNearestBRkm +1)
dpdat3$DistNearestInfra_sqrt <- sqrt(dpdat3$DistNearestInfra)
dpdat3$b_kdens_sqrt <- sqrt(dpdat3$b_kdens)
dpdat3$b_5km_logxplusone <- log(dpdat3$b_5km +1)
dpdat3$MaxHook_sqrt <- sqrt(dpdat3$MaxHook)
dpdat3$DecFishingHr_sqrt <- sqrt(dpdat3$DecFishingHr)


# Reset continuous predictor names and add coords
DPcont <- c("nTimesLast24m","exDecMedianTime_sqrt",
           "Depth_logxplusone","DistNearestBRkm_logxplusone", "DistNearestInfra_sqrt", "UseLat",
           "UseLong", "DecFishingHr_sqrt","MaxHook_sqrt",
           "BoatLength", "b_kdens_sqrt", "b_5km_logxplusone")

which(is.na(dpdat2)) # make sure no nas
which(is.nan(dpdat2$DPcont)) # make sure no nans

dpdat2 %>% 
  select(-UseLat) %>% 
    filter(if_any(where(is.numeric), ~ .x <0)) # make sure there are no negatives

# resenting to see if negatives in use lat are problem 
DPcont <- c("nTimesLast24m","exDecMedianTime_sqrt",
           "Depth_logxplusone","DistNearestBRkm_logxplusone","DecFishingHr_sqrt","MaxHook_sqrt",
           "BoatLength", "b_kdens_sqrt")

DPfac <- c("Month", "Side", "facYear", "Habitat",  "FishingType") # factors
```

### Zeros
Aiming to have less than 80% zeros in resposne
```{r zeros}
unique.vars=unique(as.character(dpdat2$nDP))
unique.vars.use=character()
for(i in 1:length(unique.vars)){
  temp.dat=dpdat2[which(dpdat2$nDP==unique.vars[i]),]
  if(length(which(temp.dat$response==0))/nrow(temp.dat)<0.8){
    unique.vars.use=c(unique.vars.use,unique.vars[i])}
}
unique.vars.use  

# 60% zeros - all good
```

## Setting up models
```{r GAMMs}

out.all <- list()
var.imp <- list()
fss.all=list()
top.all=list()
resp.vars=unique.vars.use

# k is knows determines out wigglgying we allow the line to be 
# cr is structure we give to n caught
# re = random effect

model1 <- gam(nDP~s(nHooked, k=3, bs='cr')+
                s(PersonID, bs='re'), family='poisson', data = dpdat2)

model.set <- generate.model.set(use.dat = dpdat2, 
test.fit = model1, 
pred.vars.cont = DPcont, 
pred.vars.fact = DPfac, 
max.predictors = 3, 
k=3, 
null.terms = "s(nHooked, k=5,
                                bs='cr')+s(PersonID, bs='re')", 
smooth.smooth.interactions = F, 
factor.factor.interactions=F)

out.list <- fit.model.set(model.set, max.models = 600, parallel = T)

names(out.list)
#names(out.list)
 # examine the list of failed models
 #out.list$failed.models
 #out.list$success.models
 fss.all=c(fss.all,list(out.list))
 mod.table=out.list$mod.data.out
 mod.table=mod.table[order(mod.table$AICc),]
 out.i=mod.table
 out.all=c(out.all,list(out.i))
 var.imp=c(var.imp,list(out.list$variable.importance$aic$variable.weights.raw))
 all.less.2AICc=mod.table[which(mod.table$delta.AICc<2),]
 top.all=c(top.all,list(all.less.2AICc))

all.mod.fits=do.call("rbind",out.all)
all.var.imp=do.call("rbind",var.imp)
top.mod.fits=do.call("rbind",top.all)
```

## Top Model & Predictions
```{r}
######### Predict and plot ..#########
# gam.site<-gam(response~s(bathymetry,k=3,bs='cr')+s(cube.Aspect,k=5,bs='cr') + status +
#                  s(site,bs="re"), family=tw(),data=legal.dat)

gam1 <- gam(nDP ~ s(Depth_logxplusone, by = facYear, k = 3, bs = "cr") + s(DistNearestBRkm_logxplusone, by = facYear, k = 3, bs = "cr") + facYear + 
              s(nHooked, k = 5, bs ="cr") + s(PersonID, bs = "re"), family = "poisson", data = dpdat2)

# predict bathymetry from model
mod<-gam1
testdata <- expand.grid(Depth_logxplusone=seq(min(dpdat2$Depth_logxplusone),max(dpdat2$Depth_logxplusone), length.out = 10),
                        DistNearestBRkm_logxplusone=seq(min(dpdat2$DistNearestBRkm_logxplusone),max(dpdat2$DistNearestBRkm_ogxplusone), length.out=10),
                        nHooked=seq(min(dpdat2$nHooked),max(dpdat2$nHooked), length.out = 10),
                        facYear=(mod$model$facYear),
                        PersonID=(mod$model$PersonID)) %>%
  distinct()%>%
  glimpse()

fits <- predict.gam(mod, newdata=testdata, type='response', se.fit=T)
```

### Year Predictions
```{r}
predicts.year = testdata%>%data.frame(fits)%>%
  dplyr::group_by(facYear)%>% #only change here
  dplyr::summarise(response=mean(fit),se.fit=mean(se.fit))%>%
  dplyr::ungroup()
write.csv(predicts.year, paste(dpGAM, "predict.year.csv", sep = "/")) #there is some BUG in dplyr - that this fixes
predicts.year<-read.csv(paste(dpGAM, "predict.year.csv", sep = "/")) %>%
  dplyr::mutate(facYear = as.factor(facYear)) %>%
  glimpse()

# Plot year
ggmod.gam.year<- ggplot(aes(x = facYear, y = response), data = predicts.year) +
  ylab("Predicted number of fish depredated") + 
  xlab('Year') +
  # geom_violin(data = predicts.year, aes()) +
  scale_x_discrete(limits = rev(levels(predicts.year$year)))+
  geom_errorbar(data = predicts.year, aes(ymin = response - se.fit, ymax = response + se.fit), width = 0.5) +
  theme_classic()
ggmod.gam.year
```

### Depth Predictions
```{r}
# Calculating standard error
predicts.depth = testdata%>%data.frame(fits)%>%
  dplyr::group_by(Depth_logxplusone, facYear)%>% #only change here
  dplyr::summarise(response=mean(fit),se.fit=mean(se.fit))%>%
  dplyr::ungroup() 
write.csv(predicts.depth, paste(dpGAM, "predict.depth.csv", sep = "/")) #there is some BUG in dplyr - that this fixes
predicts.depth<-read.csv(paste(dpGAM, "predict.depth.csv", sep = "/")) %>%
  dplyr::mutate(facYear = as.factor(facYear)) %>% 
  glimpse()

# Plot depth
ggmod.gam.depth <- ggplot() +
  ylab("Predicted number of fish depredated")+
  xlab("Depth (m, log + 1)")+
  geom_ribbon(data = predicts.depth, aes(x = Depth_logxplusone,y = response, ymax= response + se.fit, ymin =
                                            response - se.fit, xmin = 0, xmax = 6, group = facYear,
                                            fill = facYear, colour = facYear),alpha = 0.2, show.legend = T) +
  geom_point(data = predicts.depth, aes(x = Depth_logxplusone, y = response, colour = facYear),  alpha =
               0.75, size=2, show.legend = F) +
  geom_line(data=predicts.depth,aes(x = Depth_logxplusone,y = response, group = facYear, color = facYear), alpha = 0.75, show.legend = F) +
  scale_fill_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  scale_color_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  theme_classic()
  # annotate("text", x = -Inf, y=Inf, label = "(c)",vjust = 1, hjust = -.1,size=5)
ggmod.gam.depth
```


### Distance to Nearest Boat Ramp Predictions
```{r}
# Calculating standard error
predicts.br = testdata%>%data.frame(fits)%>%
  dplyr::group_by(DistNearestBRkm_logxplusone, facYear)%>% #only change here
  dplyr::summarise(response=mean(fit),se.fit=mean(se.fit))%>%
  dplyr::ungroup()
write.csv(predicts.br, paste(dpGAM, "predict.br.csv", sep = "/")) #there is some BUG in dplyr - that this fixes
predicts.br<-read.csv(paste(dpGAM, "predict.br.csv", sep = "/")) %>%
  dplyr::mutate(facYear = as.factor(facYear)) %>%
  glimpse()

# Plot distnace to nearest boat ramp
ggmod.gam.br<- ggplot() +
  ylab("Predicted number of fish depredated")+
  xlab('Distance to nearest boat ramp (km, log + 1)')+
  geom_ribbon(data = predicts.br, aes(x = DistNearestBRkm_logxplusone,y = response, ymax= response + se.fit, ymin =
                                            response - se.fit, xmin = 0, xmax = 6, group = facYear,
                                            fill = facYear, colour = facYear),alpha = 0.2, show.legend = T) +
  geom_point(data = predicts.br, aes(x = DistNearestBRkm_logxplusone, y = response, colour = facYear),  alpha =
               0.75, size=2, show.legend = F) +
  geom_line(data=predicts.br,aes(x = DistNearestBRkm_logxplusone,y = response, group = facYear, color = facYear), alpha = 0.75, show.legend = F) +
  scale_fill_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  scale_color_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  theme_classic()
ggmod.gam.br
```
```{r}
write.csv(all.var.imp, paste(dpGAM, "all.var.imp.csv", sep= "/"))

dat.var.imp <-read.csv(paste(dpGAM, "all.var.imp.csv", sep= "/"))%>% #from local copy
  rename(resp.var=X)%>%
  gather(key=predictor,value=importance,2:ncol(.))%>%
  mutate(Model = "Depredation") %>% 
  glimpse()

dat.var.label<-dat.var.imp%>%
  mutate(label=NA)%>%
  mutate(label=ifelse(predictor=="facYear", "X",label))%>%
  mutate(label=ifelse(predictor=="Depth_logxplusone", "X",label))%>%
  mutate(label=ifelse(predictor=="DistNearestBRkm_logxplusone", "X",label))%>%
  glimpse()

legend_title <- "Importance"

re <- colorRampPalette(c("#97c1a9","#cce2bc"))(200)

# Plot gg.importance.scores
gg.importance.scores <- ggplot(dat.var.label, aes(x=predictor, y=Model,fill=importance)) +
  geom_tile(show.legend=T) +
  scale_fill_gradientn(legend_title, colours=c("white", re), na.value = "grey98",
                       limits = c(0, max(dat.var.label$importance))) +
  scale_x_discrete(limits=c("nTimesLast24m","exDecMedianTime_sqrt","Depth_logxplusone","DistNearestBRkm_logxplusone","DecFishingHr_sqrt","MaxHook_sqrt", "BoatLength",
                            "b_kdens_sqrt", "Month", "Side", "facYear", "Habitat",  "FishingType"),
                   labels=c("Visits in last 24 months","Time of Day Fishing (sqrt)","Depth (m, log + 1 )","Distance to boat ramp (km, log + 1)","Hours fishing (sqrt)","Maximum hook depth (sqrt)", "Boat Length (m)",
                            "Boat Kernal Density (sqrt)", "Month", "Side", "Year", "Habitat",  "Fishing Type"), guide = guide_axis(angle = 45)) +
  xlab(NULL) +
  ylab(NULL) +
  theme_classic() +
  geom_text(aes(label=label))
  
gg.importance.scores
```
## Model 2 & Predictions

```{r}
# Reset continuous predictor names and add coords
DPcont <- c("nTimesLast24m","exDecMedianTime_sqrt",
           "Depth_logxplusone","DistNearestBRkm_logxplusone", "DistNearestInfra_sqrt", "UseLat",
           "UseLong", "DecFishingHr_sqrt","MaxHook_sqrt",
           "BoatLength", "b_kdens_sqrt", "b_5km_logxplusone")


dpdat2 %>% 
  select(-UseLat) %>% 
    filter(if_any(where(is.numeric), ~ .x <0)) # make sure there are no negatives

dem <- dpdat2 %>% 
  filter(FishingType %in% "Demersal")
# resenting to see if negatives in use lat are problem 
DPcont <- c("nTimesLast24m","exDecMedianTime_sqrt",
           "DistNearestBRkm_logxplusone", "DistNearestInfra_sqrt","DecFishingHr_sqrt","MaxHook_sqrt",
           "BoatLength", "b_kdens_sqrt")

DPfac <- c("Month", "facYear", "Habitat") # factors

out.all <- list()
var.imp <- list()
fss.all=list()
top.all=list()
resp.vars=unique.vars.use

# k is knows determines out wigglgying we allow the line to be 
# cr is structure we give to n caught
# re = random effect

model1 <- gam(nDP~s(nHooked, k=3, bs='cr')+
                s(PersonID, bs='re'), family='poisson', data = dem)

model.set <- generate.model.set(use.dat = dem, test.fit = model1, pred.vars.cont = DPcont, pred.vars.fact = DPfac, max.predictors = 5, k=3, null.terms = "s(nHooked, k=5,
                                bs='cr')+s(PersonID, bs='re')", smooth.smooth.interactions = F, factor.factor.interactions=F)

out.list <- fit.model.set(model.set, max.models = 600, parallel = T)

names(out.list)
#names(out.list)
 # examine the list of failed models
 #out.list$failed.models
 #out.list$success.models
 fss.all=c(fss.all,list(out.list))
 mod.table=out.list$mod.data.out
 mod.table=mod.table[order(mod.table$AICc),]
 out.i=mod.table
 out.all=c(out.all,list(out.i))
 var.imp=c(var.imp,list(out.list$variable.importance$aic$variable.weights.raw))
 all.less.2AICc=mod.table[which(mod.table$delta.AICc<2),]
 top.all=c(top.all,list(all.less.2AICc))

all.mod.fits=do.call("rbind",out.all)
all.var.imp=do.call("rbind",var.imp)
top.mod.fits=do.call("rbind",top.all)
```

```{r}
gam2 <- gam(nDP ~ s(DecFishingHr_sqrt, by = Month, k = 3, bs = "cr") + s(DistNearestInfra_sqrt, by = Month, k = 3, bs = "cr") + s(MaxHook_sqrt, by = Month, k = 3, bs = "cr") + s(nTimesLast24m, by = Month, k = 3, bs = "cr") + Month + s(nHooked, k = 5, bs = "cr") + s(PersonID, bs = "re"), family = "poisson", data = dem)

# predict bathymetry from model
mod<-gam2
testdata <- expand.grid(DecFishingHr_sqrt=seq(min(dem$DecFishingHr_sqrt),max(dem$DecFishingHr_sqrt), length.out = 4),
                        DistNearestInfra_sqrt=seq(min(dem$DistNearestInfra_sqrt),max(dem$DistNearestInfra_sqrt), length.out=4),
                        MaxHook_sqrt=seq(min(dem$MaxHook_sqrt),max(dem$MaxHook_sqrt), length.out=4),
                        nTimesLast24m=seq(min(dem$nTimesLast24m),max(dem$nTimesLast24m), length.out=4),
                        nHooked=seq(min(dem$nHooked),max(dem$nHooked), length.out = 4),
                        Month=(mod$model$Month),
                        PersonID=(mod$model$PersonID)) %>%
  distinct()%>%
  glimpse()

fits <- predict.gam(mod, newdata=testdata, type='response', se.fit=T)
```

### Year Predictions
```{r}
predicts.month = testdata%>%data.frame(fits)%>%
  dplyr::group_by(Month)%>% #only change here
  dplyr::summarise(response=mean(fit),se.fit=mean(se.fit))%>%
  dplyr::ungroup()
write.csv(predicts.month, paste(dpGAM, "predict.month.csv", sep = "/")) #there is some BUG in dplyr - that this fixes
predicts.month<-read.csv(paste(dpGAM, "predict.month.csv", sep = "/")) %>%
  dplyr::mutate(month = as.factor(Month)) %>%
  glimpse()

# Plot month
ggmod.gam.month<- ggplot(aes(x = month, y = response), data = predicts.month) +
  ylab("Predicted number of fish depredated") + 
  xlab('Month') +
  # geom_violin(data = predicts.month, aes()) +
  scale_x_discrete(limits = rev(levels(predicts.month$Month)))+
  geom_errorbar(data = predicts.month, aes(ymin = response - se.fit, ymax = response + se.fit), width = 0.5) +
  theme_classic()
ggmod.gam.month
```

### Depth Predictions
```{r}
# Calculating standard error
predicts.DecFishingHr = testdata%>%data.frame(fits)%>%
  dplyr::group_by(DecFishingHr_sqrt, Month)%>% #only change here
  dplyr::summarise(response=mean(fit),se.fit=mean(se.fit))%>%
  dplyr::ungroup() %>% 
  glimpse()
write.csv(predicts.DecFishingHr, paste(dpGAM, "predict.DecFishingHr.csv", sep = "/")) #there is some BUG in dplyr - that this fixes
predicts.DecFishingHr<-read.csv(paste(dpGAM, "predict.DecFishingHr.csv", sep = "/")) %>%
  dplyr::mutate(Month = as.factor(Month)) %>% 
  glimpse()

# Plot depth
ggmod.gam.DecFishingHr <- ggplot() +
  ylab("Predicted number of fish depredated")+
  xlab("Time spent fishing")+
  geom_ribbon(data = predicts.DecFishingHr, aes(x = DecFishingHr_sqrt,y = response, ymax= response + se.fit, ymin =
                                            response - se.fit, xmin = 0, xmax = 6, group = Month,
                                            fill =  Month, colour = Month),alpha = 0.2, show.legend = T) +
  geom_point(data = predicts.DecFishingHr, aes(x = DecFishingHr_sqrt, y = response, colour = Month),  alpha =
               0.75, size=2, show.legend = F) +
  geom_line(data=predicts.DecFishingHr,aes(x = DecFishingHr_sqrt,y = response, group = Month, color = Month), alpha = 0.75, show.legend = F) +
  # scale_fill_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  # scale_color_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  theme_classic()
  # annotate("text", x = -Inf, y=Inf, label = "(c)",vjust = 1, hjust = -.1,size=5)
ggmod.gam.DecFishingHr
```


### Distance to Nearest Boat Ramp Predictions
```{r}
# Calculating standard error
predicts.infra = testdata%>%data.frame(fits)%>%
  dplyr::group_by(DistNearestInfra_sqrt, Month)%>% #only change here
  dplyr::summarise(response=mean(fit),se.fit=mean(se.fit))%>%
  dplyr::ungroup()
write.csv(predicts.infra, paste(dpGAM, "predict.infra.csv", sep = "/")) #there is some BUG in dplyr - that this fixes
predicts.infra<-read.csv(paste(dpGAM, "predict.infra.csv", sep = "/")) %>%
  dplyr::mutate(Month = as.factor(Month)) %>%
  glimpse()

# Plot distnace to nearest boat ramp
ggmod.gam.infra<- ggplot() +
  ylab("Predicted number of fish depredated")+
  xlab('Distance to nearest infrastructure (km, sqrt)')+
  geom_ribbon(data = predicts.infra, aes(x = DistNearestInfra_sqrt,y = response, ymax= response + se.fit, ymin =
                                            response - se.fit, xmin = 0, xmax = 6, group = Month,
                                            fill = Month, colour = Month),alpha = 0.2, show.legend = T) +
  geom_point(data = predicts.infra, aes(x = DistNearestInfra_sqrt, y = response, colour = Month),  alpha =
               0.75, size=2, show.legend = F) +
  geom_line(data=predicts.br,aes(x = DistNearestInfra_sqrt,y = response, group = Month, color = Month), alpha = 0.75, show.legend = F) +
  # scale_fill_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  # scale_color_manual(labels = c("2015", "2016", "2020", "2021"), values = c("palevioletred1","lightskyblue", "lightsalmon","lightgreen"), name = "Year") +
  theme_classic()
ggmod.gam.infra

DistNearestInfra_sqrt
```

```{r}
write.csv(all.var.imp, paste(dpGAM, "all.var.imp.csv", sep= "/"))

dat.var.imp <-read.csv(paste(dpGAM, "all.var.imp.csv", sep= "/"))%>% #from local copy
  rename(resp.var=X)%>%
  gather(key=predictor,value=importance,2:ncol(.))%>%
  mutate(Model = "Depredation") %>% 
  glimpse()

dat.var.label<-dat.var.imp%>%
  mutate(label=NA)%>%
  mutate(label=ifelse(predictor=="facYear", "X",label))%>%
  mutate(label=ifelse(predictor=="Depth_logxplusone", "X",label))%>%
  mutate(label=ifelse(predictor=="DistNearestBRkm_logxplusone", "X",label))%>%
  glimpse()

legend_title <- "Importance"

re <- colorRampPalette(c("#97c1a9","#cce2bc"))(200)

# Plot gg.importance.scores
gg.importance.scores <- ggplot(dat.var.label, aes(x=predictor, y=Model,fill=importance)) +
  geom_tile(show.legend=T) +
  scale_fill_gradientn(legend_title, colours=c("white", re), na.value = "grey98",
                       limits = c(0, max(dat.var.label$importance))) +
  scale_x_discrete(limits=c("nTimesLast24m","exDecMedianTime_sqrt","Depth_logxplusone","DistNearestBRkm_logxplusone","DecFishingHr_sqrt","MaxHook_sqrt", "BoatLength",
                            "b_kdens_sqrt", "Month", "Side", "facYear", "Habitat",  "FishingType"),
                   labels=c("Visits in last 24 months","Time of Day Fishing (sqrt)","Depth (m, log + 1 )","Distance to boat ramp (km, log + 1)","Hours fishing (sqrt)","Maximum hook depth (sqrt)", "Boat Length (m)",
                            "Boat Kernal Density (sqrt)", "Month", "Side", "Year", "Habitat",  "Fishing Type"), guide = guide_axis(angle = 45)) +
  xlab(NULL) +
  ylab(NULL) +
  theme_classic() +
  geom_text(aes(label=label))
  
gg.importance.scores
```

#### without nDOP>30
```{r}
# Making all character vars factors
dpdat2 <- dpdat2 %>%
  mutate_if(is.character, as.factor) %>% 
  mutate(facYear = as.factor(facYear))
```

# GAMMs
## Assign 
## Check for correlations
```{r corrolations}
# Need to put this package in the libs
# ASk charlotte if you do this with just continuous variables of if you do factors to
# If you put them all in then stick to teh format below 
# Need to check for correlations above 0.95, any higher should be ruled out.
# Figure out which one gets ruled out

#  cor <-
#   round(cor(dpdat[,c("numYear","nTimesLast24m","exDecMedianTime","Depth", "DistNearestBRkm", "DistNearestInfra", "UseLat","UseLong", "DecFishingHr","MaxHook","BoatLength",
#                     "b_kdens", "b_5km")], use='complete.obs'), 2)
# 
# round(cor(dpdat[,c("nDP", "nHooked")], use='complete.obs'), 2)
# 
# col2 <- colorRampPalette(c("#97c1a9","#cce2bc", "white", "#cce2cb","#81caca"))
# 
# png(height=1800, width=1800, "corplot.png", type = "cairo")
# corrplot(cor, method = 'color', col = col2(10), tl.cex = 2, tl.col = "#595959", number.cex=1.5, addCoef.col="#595959", cl.cex = 2)
# dev.off()
```

## Assign variables
```{r assign cont}
DPcont <- c("nTimesLast24m","exDecMedianTime",
           "Depth","DistNearestBRkm", "DistNearestInfra", "DecFishingHr","MaxHook","BoatLength", "b_kdens", "b_5km") # No coords in this for transformations - will be added later 

DPfac <- c("Month", "Side", "facYear", "Resident", "Habitat", "LunarPhase", "FishingType") # factors
```

## Checking data
### Removing Nas from data set
```{r distrobution}

dpdat3 <- dpdat2 %>% 
  mutate(nTimesLast24m = ifelse(is.na(nTimesLast24m), 0, nTimesLast24m)) %>% 
  filter(!(ID %in% c(502, 548, 565))) %>% 
  mutate(MaxHook = ifelse(!is.na(Depth) & Depth > 0 & FishingType %in% "Demersal", Depth, MaxHook)) %>% 
  drop_na(MaxHook) %>% 
  mutate(BoatLength = ifelse(is.na(BoatLength), round(mean(BoatLength, na.rm = T)), BoatLength)) %>% 
  mutate(DecFishingHr = ifelse(is.na(DecFishingHr), round(mean(DecFishingHr, na.rm = T)), DecFishingHr)) %>% 
# turn to decimal subtract /add then turn back  into right format 
  mutate(temp_Start = StartTime) %>% 
  mutate(temp_Start = gsub("[:]", ".", temp_Start)) %>%
  mutate(temp_Start = as.numeric(temp_Start)) %>% 
  mutate(temp_Stop = StopTime) %>% 
  mutate(temp_Stop = gsub("[:]", ".", temp_Stop)) %>% # will substitute all : with . within temp_start
  mutate(temp_Stop = as.numeric(temp_Stop)) %>% 
  mutate(temp_Start = ifelse(is.na(temp_Start), temp_Stop - DecFishingHr, temp_Start)) %>% 
  mutate(temp_Stop = ifelse(is.na(temp_Stop), temp_Start + DecFishingHr, temp_Stop)) %>% 
  rowwise() %>%
  mutate(temp_Med = median(c(temp_Start, temp_Stop))) %>%
  ungroup() %>% 
  mutate(temp_Med = gsub("[.]", ":", temp_Med)) %>% 
  mutate(Resident = as.factor(Resident)) %>% 
  mutate(temp_Med = ifelse(str_length(temp_Med)==2, str_c(temp_Med, ":00"), temp_Med)) %>% # if string is 2 digits long it will ass ":00" to end of colum temp_Med
  mutate(temp_Med = ifelse(str_length(temp_Med)==3, str_c("0", temp_Med, "0"), temp_Med)) %>% # if string is 3 digits long it will add a 0 to the strat and end of whats in temp_Med
  mutate(temp_Med = ifelse(str_length(temp_Med)==4, str_c(temp_Med, "0"), temp_Med)) %>% 
  mutate(exDecMedianTime = ifelse(!is.na(temp_Med), hhmm2dec(temp_Med), exDecMedianTime)) %>% 
  mutate(exDecMedianTime = round(exDecMedianTime, digits = 2)) %>% # rounding to 2 digits after decimal 
  mutate(exDecMedianTime = ifelse(is.na(exDecMedianTime), 12, exDecMedianTime)) %>% 
  mutate(exDecMedianTime = as.numeric(exDecMedianTime)) %>% 
  select(PersonID, nDP, nHooked, nTimesLast24m, exDecMedianTime,
           Depth,DistNearestBRkm, DistNearestInfra, UseLat,
           UseLong, DecFishingHr,MaxHook, BoatLength, b_kdens, b_5km, Month, Side, facYear, Resident, Habitat, LunarPhase, FishingType) %>% 
  mutate(Depth = Depth*-1) %>% 
  mutate(Depth = ifelse(PersonID %in% 190, 0, Depth)) %>% 
  as.data.frame()

# nTimesLast24m_sqrt - assuming in non entries are due to people not entering 0s if people haven't been to Exmouth before - hence filled with 0 
# Depth_logxplusone - removed 3 rows 
# MaxHook - if fishing was demersal made max hook equal to depth
# BoatLength - filled in NAs with average boat length
# DecFishingHr - filled nas with average fishing hours (4)
# exDecMedianTime - if there was a start time i added the average fishing time (4) to get and end time and caculated the median from those times - and vice versa for when theres an end time and no start time. cases where there were neither a start or end time were filled with the average median time (12)
```

### Distribution
Looking at data through various transformations to see which transformation provides the most normal distribution and is homoscedastic. 
```{r distrobution}
# need to run in console
par(mfrow=c(3,4))
for(i in DPcont) {
  x <- dpdat2[,i]
  x = as.numeric(unlist(x))
  hist((x))
  plot((x), main=paste(i))
  hist(sqrt(x))
  plot(sqrt(x))
  hist(log(x+1))
  plot(log(x+1))
  hist((x))
  plot((x), main=paste(i))
  hist((x)^2)
  plot((x)^2)
  hist((x)^3)
  plot((x)^3)
}

# nTimesLast24m: x
# exDecMedianTime: log(x+1)
# Depth: log(x+1)
# DistNearestBR:log(x+1)
# DistNearestInfra: sqrt
# UseLat: x
# UseLong:x
# DecFishingHr:sqrt
# MaxHook: sqrt
# BoatLength: x
# b_kdens: sqrt
# b_5km: log(x+1)


# Transform variables into dataset
dpdat2$exDecMedianTime_sqrt <- sqrt(dpdat2$exDecMedianTime)
dpdat2$Depth_logxplusone <- log(dpdat2$Depth +1)
dpdat2$DistNearestBRkm_logxplusone <- log(dpdat2$DistNearestBRkm +1)
dpdat2$DistNearestInfra_sqrt <- sqrt(dpdat2$DistNearestInfra)
dpdat2$b_kdens_sqrt <- sqrt(dpdat2$b_kdens)
dpdat2$b_5km_logxplusone <- log(dpdat2$b_5km +1)
dpdat2$MaxHook_sqrt <- sqrt(dpdat2$MaxHook)
dpdat2$DecFishingHr_sqrt <- sqrt(dpdat2$DecFishingHr)


# Reset continuous predictor names and add coords
DPcont <- c("nTimesLast24m","exDecMedianTime_sqrt",
           "Depth_logxplusone","DistNearestBRkm_logxplusone", "DistNearestInfra_sqrt", "UseLat",
           "UseLong", "DecFishingHr_sqrt","MaxHook_sqrt",
           "BoatLength", "b_kdens_sqrt", "b_5km_logxplusone")

which(is.na(dpdat2)) # make sure no nas
which(is.nan(dpdat2$DPcont)) # make sure no nans

dpdat2 %>% 
  select(-UseLat) %>% 
    filter(if_any(where(is.numeric), ~ .x <0)) # make sure there are no negatives

# resenting to see if negatives in use lat are problem 
DPcont <- c("nTimesLast24m","exDecMedianTime_sqrt",
           "Depth_logxplusone","DistNearestBRkm_logxplusone","DecFishingHr_sqrt","MaxHook_sqrt",
           "BoatLength", "b_kdens_sqrt")

DPfac <- c("Month", "Side", "facYear", "Habitat",  "FishingType") # factors
```

### Zeros
Aiming to have less than 80% zeros in resposne
```{r zeros}
unique.vars=unique(as.character(dpdat2$nDP))
unique.vars.use=character()
for(i in 1:length(unique.vars)){
  temp.dat=dpdat2[which(dpdat2$nDP==unique.vars[i]),]
  if(length(which(temp.dat$response==0))/nrow(temp.dat)<0.8){
    unique.vars.use=c(unique.vars.use,unique.vars[i])}
}
unique.vars.use  

# 60% zeros - all good