---
title: "RUM Analysis"
author: "Nicole Hamre"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook:
    df_print: paged
    theme: cosmo
    toc: yes
    toc_float: yes  
  html_document:
    df_print: paged
    theme: cosmo
    toc: yes
  word_document:
    number_section: yes
    toc: yes
---

```{r libraryload, include = FALSE}
library(ezknitr)
library(here)
library(fs)
library(knitr)
```


```{r setglobal, include = FALSE}

getwd()

```

```{r plotCars}
plot(cars)
```


`..` means the parent directory or directory above the current one
`.` means the current directory
A leading `/' at the beginning of a path tells the computer to start at the root directory - should be the **absolute path**.


Attempt 

ezknit(file = "./Ch2-RUM/rumCode/rumMother.Rmd", out_dir = "./Ch2-RUM/rumOut/rumReport", fig_dir = "rumPlot",chunk_opts = list(fig.align = "center", dev = 'png', fig.width = 11, fig.height = 8.5)) 
getwd()
fs::dir_tree()
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("./Ch2-RUM"))
knitr::opts_knit$get("root.dir")

attempt
knitr::opts_knit$set(root.dir = normalizePath("./Ch2-RUM"))
knitr::opts_knit$get("root.dir") 

ezknit(file = "Ch2-RUM/rumCode/rumMother.Rmd", out_dir = "./rumOut", fig_dir = "./rumOut/rumPlot", chunk_opts = list(fig.align = "center", dev = 'png', fig.width = 11, fig.height = 8.5))



