Check habitat layers align

```{r habitat}
ggplot() +
  # geom_sf(data = Coast, fill = "grey", lwd = 0.1) +
  geom_sf(data = reef, fill = "orange",  lwd = 0.1) +
  geom_sf(data = rubble, fill = "aquamarine",  lwd = 0.1) +
  geom_sf(data = macroalgae, fill = "green",  lwd = 0.1) +
  geom_sf(data = sand, fill = "yellow",  lwd = 0.1) +
  geom_sf(data = silt, fill = "brown",  lwd = 0.1) +
  geom_sf(data = seagrass, fill = "light green",  lwd = 0.1) +
  geom_sf(data = pelagic, fill = "light blue",  lwd = 0.1)
```

## Creating Habitat Grids
If you need to change size of grids adapt below code. 
When you make these they will be large sfc - when you write, and then read them again they will become an sf object ready to append habitat data.

```{r Creating habitat grids}
# Pelagic
# PelagicGrid <- pelagic %>%
#   st_make_grid(cellsize = 0.05, square = FALSE, crs=4283) %>%
#   st_intersection(pelagic)
#  st_write(PelagicGrid, paste(gpkg.dir, "Esp_PelagicGrid.gpkg", sep = '/'), append = FALSE)
# 
# # Reef
# ReefGrid <- reef %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(reef)
# st_write(ReefGrid, paste(gpkg.dir, "Esp_ReefGrid.gpkg", sep = '/'), append = FALSE)
# 
# # Rubble
#  RubbleGrid <- rubble %>%
#    st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#    st_intersection(rubble)
#  st_write(RubbleGrid, paste(gpkg.dir, "Esp_RubbleGrid.gpkg", sep = '/'), append = FALSE)
# 
# Sand
SandGrid <- sand %>%
  st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
  st_intersection(sand)
st_write(SandGrid, paste(gpkg.dir, "Esp_SandGrid.gpkg", sep = '/'), append = FALSE)
# 
# # Silt
# SiltGrid <- silt %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(silt)
# st_write(SiltGrid, paste(gpkg.dir, "Esp_SiltGrid.gpkg", sep = '/'), append = FALSE)
# 
# # Seagrass
# SeagrassGrid <- seagrass %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(seagrass)
# st_write(SeagrassGrid, paste(gpkg.dir, "Esp_SeagrassGrid.gpkg", sep = '/'), append = FALSE)
# 
# # Macroalgae
# MacroalgaeGrid <- macroalgae %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(macroalgae)
# st_write(MacroalgaeGrid, paste(gpkg.dir, "Esp_MacroalgaeGrid.gpkg", sep = '/'), append = FALSE)
```

### Making unified habitat grid
If manipulating the gridd sizes in chunk above you need to read in those layers after writing them to turn them from sfc to sf - which will then allow you to append habitat information. 
```{r Grid}
PelagicGrid <- st_read(paste(gpkg.dir, "Esp_PelagicGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

ReefGrid <- st_read(paste(gpkg.dir, "Esp_ReefGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

RubbleGrid <- st_read(paste(gpkg.dir, "Esp_RubbleGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

MacroalgaeGrid <- st_read(paste(gpkg.dir, "Esp_MacroalgaeGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

SandGrid <- st_read(paste(gpkg.dir, "Esp_SandGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

SiltGrid <- st_read(paste(gpkg.dir, "Esp_SiltGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

SeagrassGrid <- st_read(paste(gpkg.dir, "Esp_SeagrassGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
# 
# ## Adding habitat variable to Grids
# PelagicGrid$Habitat <- c("Pelagic")
# ReefGrid$Habitat <- c("Reef")
# RubbleGrid$Habitat <- c("Rubble")
# MacroalgaeGrid$Habitat <- c("Macroalgae")
# SandGrid$Habitat <- c("Sand")
# SiltGrid$Habitat <- c("Silt")
# SeagrassGrid$Habitat <- c("Seagrass")
# 
# ## Joining Grids
Grid <- rbind(PelagicGrid, ReefGrid, RubbleGrid, MacroalgaeGrid, SandGrid, SiltGrid, SeagrassGrid) # all not work
Grid2 <- rbind(PelagicGrid, ReefGrid, RubbleGrid, MacroalgaeGrid, SiltGrid, SeagrassGrid) # multipolygon and polygon dont work
Grid3 <- rbind(PelagicGrid, ReefGrid, MacroalgaeGrid, SeagrassGrid)
# 
# ## Giveing each grid individual ID
# Grid <- Grid %>%
#   mutate(GridID = row_number())
# 
# st_write(Grid, paste(gpkg.dir, "Esp_Grid.gpkg", sep = '/'), append = FALSE)
# 
Grid2
st_geometry_type(Grid2)
```