# Clean
## Rename
Rename questions to make them R friendly
```{r select and filter}
dat <- dat %>% 
dplyr::rename(ID = "...1",
              PersonID = mObjID,
              Interviewer = "# Interviewer",
              Site = "# Site",
              Screen18 = "Screen: 18+",
              PrevInter = "Screen: Have you been surveyed in the last 2 weeks?", 
              Agreement = "Agreement to participate",
              Resident = "Are you/have you been a resident, got a holiday home here or been here over 3 months?",
              
              nDaysInArea = "If here for less than 3 months, how many DAYS will you be in the area in total on this trip?",
              BoatAccess = "While you are on your trip, do you have access to a motorised boat/jetski (private/rental)?",
              nBoatDays = "If YES, how many DAYS out of your trip will you be going out on a boat in total?",
              nShoreDays = "How many DAYS of your trip will you be doing coastal shore based activities in total?",

              nTimesLast12m = "Excluding this trip, how many times have you been to the area in the last 12 month?",
              Date = "created_date",
              exSiteType = "# Is this a boat based or shore based activity?",
              exRecall = "If RECALL, what date did they go fishing?",
              FishingType = "What type of fishing did you do?",
              BaitLure = "Did you use bait or lure?",
              exStart = "What time did you START fishing at this site?",
              exStop = "What time did you STOP fishing at this site?",
              MaxHook = "What was your maximum hook depth (m) at this site?",
              KeptUndam = "Number of fish kept undamaged at this site?",
              RelUndam = "How many fish did you release at this site?",
              nDP = "How many shark bite offs did you get at this site?",
              Species = "What species of fish did you keep?", 
              exWhyLeave = "Why did you leave this site?",
              exComments = "Comments.x",
              nexSiteType = "# Is this a boat-based or shore-based activity?",
              nexRecall = "If RECALL, what date did they participant do the activity?",
              Activity = "What activity did you do?",
              nexStart = "What time did you START the activity?",
              nexStop = "What time did you STOP the activity?",
              DiveMaxDepth = "If DIVING, what was your maximum depth (m)?",
              WhyChoose = "Why did you CHOOSE this site?",
              nexWhyLeave = "Why did you LEAVE this site?",
              nexTimes12m = "How many DAYS have you done this activity in the last 12 months?",
              nexYrs = "How many YEARS have you been doing this activity?",
              nexComments = "Comments.y",
              exTimes12m = "How many days have you fished in the last 12 months?",
              exYrs = "How many years have you been fishing for?",
              DiveCert = "Are you a scuba diver? If so, what is your certification?",
              nDives = "If SCUBA DIVER, how many dives have you done?",
              YrBorn = "What year were you born?",
              Postcode = "What is your postcode?",
              Accom = "If VISITOR, where are you staying?",
              Sex = "# Female or Male",
              nFemales = "# Number of females in party (excluding participant)",
              nMales = "# Number of males in party (excluding participant)",
              nGirls = "# Number of girls in party",
              nBoys = "# Number of boys in party",
              BoatLength = "# Boat Length (m)",
              BoatID = "# Boat number and/or name",
              aComments = "Comments",
              exDetract = "Did anything detract from the experience you had at this site?",
              nexDetract = "Did anything detract from your experience at this site?",
              Native = "Are you of Aboriginal or Torres Strait Islander origin?",
              TO = "Do you identify as a TO in this region?",
              exCustom = "If of Aboriginal or Torres Strait Islander origin, was this activity undertaken for a customary purpose?.x",
              nexCustom = "If of Aboriginal or Torres Strait Islander origin, was this activity undertaken for a customary purpose?.y")
```

## Temporal variables
  - The date column is automated so there should be no NA's unless the agreement is no. 
  - Adhere start and stop times for extractive and non-extractive activities into one column,
    there should only be blanks where agreement is no. 
  - Due to recall trips need to make a trip date column which may be different to survey date.
  - Any data out with field trip dates is a dummy/test survey
  
```{r temporal variables}
# Date (Date) and Year (num/fac)
dat <- dat %>%
  mutate(Date = parse_date_time(Date, c("mdY IMS p"))) %>%  # making POSIX
  mutate(Date = as.Date(as.character(
    as.Date(substr(Date, 1, 10), "%Y-%m-%d")))) %>% # extracting date
  mutate(numYear = as.numeric(substr(Date, 1, 4))) %>% # extracting year (numeric)
  mutate(facYear = as.factor(numYear))  # year (factor)

# TripDate (Date), TripMonth (chr), TripJulianDay (num), binRecall (fac)
dat <- dat %>%
  mutate(exRecall = parse_date_time(exRecall, c("mdY IMS p"))) %>% # making POSIX
  mutate(exRecall = as.Date(substr(exRecall, 1, 10), "%Y-%m-%d")) %>% # extracting date
  mutate(nexRecall = parse_date_time(nexRecall, c("mdY IMS p"))) %>% # making POSIX
  mutate(nexRecall = as.Date(substr(nexRecall, 1, 10), "%Y-%m-%d")) %>% # extracting date
  mutate(TripDate = as.Date(ifelse(!is.na(exRecall), as.character(as.Date(exRecall)),
                                   as.character(as.Date(nexRecall))))) %>% # adhere recalls
  mutate(TripDate = as.Date(ifelse(is.na(TripDate),as.character(as.Date(Date)),
                                       as.character(as.Date(TripDate))))) %>% # adhere date
  mutate(TripMonth = as.factor(format(TripDate,"%B"))) %>% # month (chr)
  mutate(TripJulianDay = as.numeric(julian(TripDate))) %>% # julian date (num)
  mutate(binRecall = as.factor(ifelse(TripDate==Date, 0, 1))) # binary recall (fac)
  
# Start (chr), Stop (chr), MedianTime (chr), decMedianTime (num), decDuration (num)
dat <- dat %>%
  mutate(Start = substr(as.character(ifelse(!is.na(exStart), as.character(exStart),
                                     as.character(nexStart))), 1, 5)) %>% # start time
  mutate(Start = sub("[.]", ":", Start)) %>%  # change every bit of punctuation to a colon
  mutate(Start = ifelse(str_length(sub(":.*", "", Start)) != 2, str_c("0", Start),Start)) %>% 
  # if str_length before colon !=2, add a zero to the start
  mutate(Start = ifelse(str_length(sub(".*:", "", Start)) != 2, str_c(Start, "0"), Start)) %>% # if str_length after colon !=2, add a zero to the end
  mutate(Start = ifelse(Start %in% c("011am0", "11:90"), "11:00", Start)) %>% 
  mutate(Stop = substr(as.character(ifelse(!is.na(exStop), as.character(exStop),
                                     as.character(nexStop))), 1, 5)) %>%  # stop time
  mutate(Start = sub("[.]", ":", Start)) %>% # change every bit of punctuation to a colon
  mutate(Stop = ifelse(str_length(sub(":.*", "", Stop)) != 2, str_c("0", Stop),
                         Stop)) %>% # if str_length before colon !=2, add a zero to the start
  mutate(Stop = ifelse(str_length(sub(".*:", "", Stop)) != 2, str_c(Stop, "0"),
                         Stop)) # if str_length after colon !=2, add a zero to the end

# Need to check that all times are 4 digit, 5 character, 24 hour format separated with colon
# eg. 07:30, 19:30 NOT 7:30, 7pm, 19.30
# unique(dat$Start)
# unique(dat$Stop)

dat <- dat %>%
  mutate(temp_Start = parse_date_time(Start, c("HM"))) %>% # making temp_start POSIX for calcs
  mutate(temp_Stop = parse_date_time(Stop, c("HM"))) %>% # making temp_stop POSIX for calcs
  rowwise() %>%
  mutate(MedianTime = substr(median(c(temp_Start, temp_Stop)), 12, 16)) %>% # median time (chr)
  ungroup() %>% 
  mutate(decMedianTime = as.numeric(round(hhmm2dec(MedianTime), 
                                          digits = 2))) %>% #decimal median time (num)
  mutate(decDuration = as.numeric(round(difftime(
    temp_Stop, temp_Start, units = "hours"), digits = 2))) # decimal duration of activity (num)

# FieldTrip (num)
# Trip 1: 2022-01-16 - 2022-01-25
# Trip 2: 2022-02-14 - 2022-03-01
dat <- dat %>%
  mutate(FieldTrip = as.factor(ifelse(Date %within% interval(
    ymd("2022-01-16"), ymd("2022-01-25")), "1",NA))) %>% # assigning dates to field trip 1
  mutate(FieldTrip = as.factor(ifelse(Date %within% interval(
    ymd("2022-02-14"), ymd("2022-02-28")), "2", FieldTrip))) %>%
  mutate(FieldTrip = as.factor(ifelse(Date %within% interval(
    ymd("2022-03-07"), ymd("2022-03-20")), "2", FieldTrip))) %>%
  dplyr::filter(!is.na(FieldTrip), Interviewer != "Dummy") # filtering out submissions out with field trip dates (dummys)
```

## IDs
  -TripNum: all activities on one boat launch (individual date) are collectively a trip, only       applicable is data from recalled trips was collected (if TripNum >1 then binRecall should        equal 1). TripNum is chronological. 
  -SiteNum: the chronological order a site was visited within one trip
  
  * Mutate Person ID to reflect that people with same boat ID should have the same associated person ID. 

```{r IDs}
# Need to fix boat ID so everything is just 2 letter and 3 numbers, anything with 1 in position 2 should be I, move in boat names into boat name column
# dat1 <- dat %>%
#   # dplyr::mutate(BoatID = ifelse(str_detect(BoatID, "[H, h]ire|[R,r]ental|M&H"), "Rental",
#   # BoatID)) %>%
#   mutate(BoatID = gsub("[ ]", "", BoatID)) %>%
#   mutate(BoatID = toupper(BoatID)) %>% 
#   # mutate(BoatID = ifelse(str_detect(BoatID, "^.*[a-zA-Z]{1}[0-9]{4}.*$") & 
#   #                          str_sub(BoatID, 2,2)=="1", str_sub(BoatID, 2,2) <-"CHANGE", BoatID))
#  mutate(BoatName = ifelse(str_detect(BoatID, "^.*[a-zA-Z]{1}[0-9]{4}.*$") & str_sub(BoatID,2,2)=="1", str_sub(BoatID, 2,2) <- "I", NA))
# 
# 
#   mutate(BoatName = ifelse(!str_detect(BoatID, "^.*[a-zA-Z]{2}[0-9]{3}.*$|^.*[0-9]{5}.*$"), BoatID, NA))
#   mutate(BoatID = ifelse(BoatID == "WETO", NA, BoatID)) %>% 
#   mutate(BoatID = ifelse(BoatID == "49621/PEGASUS", 49621, BoatID)) 
#   mutate(PersonID = ifelse(duplicated(BoatID), ))
#   
#   group_by(BoatID) %>% 
#   
#   tidyr::fill(PersonID) %>% # NOT WORKING - boat id is not empty so wont fill. Fill fills NAs but no person ID is NA
#   ungroup() 
# 
#    dplyr::group_by(BoatName) %>% 
#   dplyr::mutate(PersonID = ifelse(anyDuplicated(BoatName) & BoatName != "RENTAL" & !is.na(BoatName), min(PersonID), PersonID)) %>%
#   dplyr::ungroup() %>% 
#   dplyr::group_by(BoatID) %>%
#   dplyr::mutate(PersonID = ifelse(anyDuplicated(BoatID) & BoatID != "RENTAL" & !is.na(BoatID), min(PersonID), PersonID)) %>%
#   dplyr::ungroup()
# 
# Ning <- Ning %>%
#   dplyr::group_by(PersonID) %>%
#   dplyr::arrange(TripDate, StartTime) %>%
#   dplyr::mutate(TripNum = as.numeric(as.factor(TripDate))) %>%
#   dplyr::mutate(SiteNum = sequence(rle(TripNum)$lengths)) %>%
#   dplyr::arrange(Date, Time, PersonID) %>%
#   dplyr::ungroup()
# 
# table(dat1$BoatID)
# dat1[b, c("PersonID", "BoatID", "PrevInter")]

dat <- dat %>%
  group_by(PersonID) %>%
  arrange(Start) %>% 
  mutate(TripNum = as.numeric(as.factor(TripDate))) %>% # number trips done by same person across different days
  mutate(SiteNum = sequence(rle(TripNum)$lengths)) %>% # chronological order of which a participant has conducted uses in a single trip
  arrange(ID) %>%
  dplyr::ungroup()
```

## Meta cleaning
```{r meta}
dat <- dat %>% 
  mutate(nBoatDays = as.numeric(ifelse(nBoatDays %in% "All", nBoatDays, nBoatDays))) %>% 
  mutate(nShoreDays = as.numeric(ifelse(nShoreDays %in% "All", nDaysInArea, nShoreDays))) %>% 
  mutate_each_(funs(factor(.)), c("Interviewer", "Screen18", "PrevInter", "Agreement",
                                  "BoatAccess", "Resident", "facYear"))
```

## Use cleaning
```{r use surveys}
dat <- dat %>% 
  mutate(SiteType = as.factor(ifelse(!is.na(exSiteType), exSiteType, nexSiteType))) %>% 
  mutate(ActivityType = as.factor(ifelse(!is.na(exActivityType), exActivityType, nexActivityType))) %>% 
  mutate(WhyLeave = as.factor(ifelse(!is.na(exWhyLeave), exWhyLeave, nexWhyLeave))) %>% 
  mutate(Detract = as.factor(ifelse(!is.na(exDetract), exDetract, nexDetract))) %>% 
  mutate(UseLat = ifelse(!is.na(exLat), exLat, nexLat)) %>% 
  mutate(UseLong = ifelse(!is.na(exLong), exLong, nexLong)) %>% 
  mutate(Activity = as.factor(ifelse(ActivityType == "Extractive", "Fishing", Activity))) %>% 
  mutate(MaxHook = ifelse(FishingType %in% "Trolling", 0, MaxHook)) %>% 
  mutate_each_(funs(factor(.)), c("FishingType", "WhyChoose", "BaitLure", "Site")) %>% 
  mutate(Comments = paste(exComments, nexComments, aComments, sep = "|"))
```

### Catch data
```{r Catch data}
dat <- dat %>%
  mutate(CaughtUndam = as.numeric(RelUndam + KeptUndam)) %>% 
  mutate(nHooked = as.numeric(CaughtUndam + nDP)) %>% 
  mutate(binDP = ifelse(nDP > 0, "1","0")) %>% 
  mutate(binDP = as.factor(binDP)) %>% 
  mutate(perDP = ifelse(nHooked %in% "0", 0, as.numeric(round((nDP/nHooked)*100, 2)))) %>% 
  mutate(proDP = as.numeric(round(perDP/100, 2))) %>% 
  mutate(nUndamHr = as.numeric(round(CaughtUndam/decDuration, 2))) %>% 
  mutate(nDPHr = as.numeric(round(nDP/decDuration, 2))) %>% 
  mutate(nHookedHr = as.numeric(round(nHooked/decDuration, 2))) %>% 
  mutate(DPrateHr = as.numeric(round(nDPHr/nHookedHr, 2))) %>% 
  mutate(DPrateHr = ifelse(is.nan(DPrateHr), 0, DPrateHr))
```

## Avidity and demographics cleaning
```{r avidity clean}
dat <- dat %>% 
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 0, 4), "0-4", NA)) %>% 
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 5, 9), "5-9", bin_exTimes12m)) %>%
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 10, 14), "10-14", bin_exTimes12m)) %>%
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 15, 19), "15-19", bin_exTimes12m)) %>%
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 20, 29), "20-29", bin_exTimes12m)) %>%
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 30, 39), "30-39", bin_exTimes12m)) %>%
  mutate(bin_exTimes12m = ifelse(between(exTimes12m, 40, 59), "40-59", bin_exTimes12m)) %>%
  mutate(bin_exTimes12m = ifelse(exTimes12m > 59, "60+", bin_exTimes12m)) 
# Age 
dat <- dat %>% 
  mutate(Age = 2021-YrBorn) %>% 
  mutate(binAge = ifelse(between(Age, 18, 25), "18-24", NA)) %>% 
  mutate(binAge = ifelse(between(Age, 25, 34), "25-34", binAge)) %>%
  mutate(binAge = ifelse(between(Age, 35, 44), "35-44", binAge)) %>%
  mutate(binAge = ifelse(between(Age, 45, 54), "45-54", binAge)) %>%
  mutate(binAge = ifelse(between(Age, 55, 64), "55-64", binAge)) %>%
  mutate(binAge = ifelse(between(Age, 65, 74), "65-74", binAge)) %>%
  mutate(binAge = ifelse(Age>74, "75+", binAge))

# Postcode
dat <- dat %>% 
  mutate(Postcode = as.factor(ifelse(Resident == "Yes", "6450", Postcode))) %>% 
  mutate(Postcode = as.factor(ifelse(Postcode ==" 0", NA, Postcode))) 

# Party
dat <- dat %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(nFemales = ifelse(is.na(nFemales), 0, nFemales)) %>%
  mutate(nMales = ifelse(is.na(nMales), 0, nMales)) %>%
  mutate(nGirls = ifelse(is.na(nGirls), 0, nGirls)) %>%
  mutate(nBoys = ifelse(is.na(nBoys), 0, nBoys)) %>%
  mutate(nFemales = ifelse(Sex %in% "Female", nFemales + 1, nFemales)) %>%
  mutate(nMales = ifelse(Sex %in% "Male", nMales + 1, nMales)) %>% 
  mutate(nAdults = nFemales + nMales) %>% 
  mutate(nChild = nGirls + nBoys) %>% 
  mutate(PartySize = nAdults + nChild) %>% 
  mutate_each_(funs(factor(.)), c("Accom", "DiveCert"))
```

## Lunar Phase
```{r lunar phase}
dat <- dat %>%
  mutate(LunarPhase = as.factor(lunar.phase(Date, name = TRUE)))
```

```{r select}
dat <- dat %>% 
  dplyr::select(
    ID, Date, numYear, facYear, Interviewer, FieldTrip, mLat, mLong, PersonID, Screen18,
         PrevInter, Agreement, binRecall, TripDate, TripMonth, TripJulianDay, TripNum, SiteNum,
         Site, SiteType, Resident, BoatAccess, nDaysInArea, nBoatDays, nShoreDays, nTimesLast12m,
         UseLat, UseLong, ActivityType, Activity, Start, Stop, MedianTime, decMedianTime,
         decDuration, FishingType, BaitLure, MaxHook, KeptUndam, RelUndam, nDP, CaughtUndam,
         nHooked, binDP, perDP, proDP, CaughtUndam, nHooked, nUndamHr, nDPHr, DPrateHr, Species,
         DiveMaxDepth, WhyChoose, WhyLeave, Detract, bin_exTimes12m, exYrs, nexTimes12m, nexYrs,
         DiveCert, nDives, binAge, Postcode, Accom, Sex, nFemales, nMales, nGirls, nBoys, nAdults,
         nChild, PartySize, BoatLength, BoatID, Comments, LunarPhase)
```