# Set Up
```{r libs, include = FALSE}
# install.packages('tidyverse')
# install.packages('dplyr')
# install.packages('ggplot2')
# install.packages('sp')
# install.packages('raster')
# install.packages('leaflet')
# install.packages('rgeos')
# install.packages('rgdal')
# install.packages('sf') # select no
# install.packages('ggsn')
# install.packages('nngeo')
# install.packages('spatstat')

library(tidyverse)
library(dplyr)
library(ggplot2)
library(sp)
library(raster)
library(leaflet)
library(rgeos)
library(rgdal)
library(sf)
library(ggsn)
library(nngeo)
library(spatstat)
```

```{r dir, echo = FALSE}
w.dir <- "/Users/23088313/Documents/git_repos/Analysis-Hamre-Bioeconomic"
d.dir <- paste(w.dir, "Ch2-RUM/rumIgnore", sep='/')
s.dir <- paste(w.dir, "spIgnore/shp", sep='/')
r.dir <- paste(w.dir, "spIgnore/raster", sep='/')
f.dir <- paste(w.dir, "Ch2-RUM/rumFunc", sep = '/')
rumPlots <- paste(w.dir, "Ch2-RUM/rumPlots", sep='/')
```

# Spatial variables

Main unprojected crs = 4283 (GDA94), units = m 
Main projected crs = 3112 (Australian Lambert) 
Raster extraction crs = 4326 (WGS84)

## Creating layers
Everything in the following chunk only needs to be done once - then only needs read. 
```{r Creating layers}
# Adding coast
WA <- st_read(paste(s.dir, "waCoast.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

# Making bbox
StudySite = st_as_sfc(st_bbox(c(xmin = 112.5, ymin = -24.03353, xmax = 115.00, ymax = -20.5), crs = 4283))
 st_write(StudySite, paste(s.dir, "StudySite.gpkg", sep = '/'), append = F)

# Cropping coast to study site
NWS <- st_crop(WA, StudySite)
NWS <- st_union(NWS)
st_write(NWS, paste(s.dir, "NWScoast.gpkg", sep = '/'), append = F)

# Cropping commonwealth mp areas
ncmp1 <- st_read(paste(s.dir, "ausStateCommonSplit.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
ncmp1.1 <- st_crop(ncmp1, site) 
st_write(ncmp1.1, paste(s.dir, "ncmpSplit.gpkg", sep = '/'), append = F)


ncmp2 <- st_read(paste(s.dir, "ausStateSanc.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

ncmp2.1 <- st_make_valid(ncmp2)
ncmp2.1 <- st_crop(ncmp2.1, site)

st_write(ncmp2.1, paste(s.dir, "ncmpSZ.gpkg", sep = '/'), append = F)

## Adding habitat layers
Pelagic <- st_read(paste(s.dir, "PelagicHabitat.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

Rockyreef <- st_read(paste(s.dir, "RockyreefHabitat.gpkg", sep = '/')) %>%
   st_transform(crs = 4283)
Rockyreef <- st_make_valid(Rockyreef)

Reef <- st_read(paste(s.dir, "ReefHabitat.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

Lagoon <- st_read(paste(s.dir, "LagoonHabitat.gpkg", sep = '/')) %>%
  st_transform(crs = 4283) %>%
  st_crop(StudySite)

# Check layers all line up.
ggplot() +
  geom_sf(data = Rockyreef, fill = "pink", lwd = 0.1) +
  geom_sf(data = Reef, fill = "orange",  lwd = 0.1) +
  geom_sf(data = Lagoon, fill = "aquamarine",  lwd = 0.1) +
  geom_sf(data = Pelagic, fill = "light blue",  lwd = 0.1)
```

## Creating Habitat Grids
If you need to change size of grids adapt below code. 
When you make these they will be large sfc - when you write, and thenm read them again theyw ill become an sf object ready to append habitat data.

```{r Creating habitat grids}

PelagicGrid <- Pelagic %>%
  st_make_grid(cellsize = 0.05, square = FALSE, crs=4283) %>%
  st_intersection(Pelagic)
 st_write(PelagicGrid, paste(s.dir, "PelagicGrid.gpkg", sep = '/'), append = FALSE)

RockyreefGrid <- Rockyreef %>%
  st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
  st_intersection(Rockyreef)
st_write(RockyreefGrid, paste(s.dir, "RockyreefGrid.gpkg", sep = '/'), append = FALSE)

 ReefGrid <- Reef %>%
   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
   st_intersection(Reef)
 st_write(ReefGrid, paste(s.dir, "ReefGrid.gpkg", sep = '/'), append = FALSE)

LagoonGrid <- Lagoon %>%
  st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
  st_intersection(Lagoon)
st_write(LagoonGrid, paste(s.dir, "LagoonGrid.gpkg", sep = '/'), append = FALSE)

```

### Making unified habitat grid
If manipulating the gridd sizes in chunk above you need to read in those layers after writing them to turn them from sfc to sf - which will then allow you to append habitat information. 
```{r Grid}
PelagicGrid <- st_read(paste(s.dir, "PelagicGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

RockyreefGrid <- st_read(paste(s.dir, "RockyreefGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

ReefGrid <- st_read(paste(s.dir, "ReefGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

LagoonGrid <- st_read(paste(s.dir, "LagoonGrid.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

## Adding habitat variable to Grids
PelagicGrid$Habitat <- c("Pelagic")
RockyreefGrid$Habitat <- c("Rocky reef")
ReefGrid$Habitat <- c(" Coral Reef")
LagoonGrid$Habitat <- c("Lagoon")

## Joining Grids
Grid <- rbind(PelagicGrid, RockyreefGrid, ReefGrid, LagoonGrid)

## Giveing each grid individual ID
Grid <- Grid %>%
  mutate(GridID = row_number())

st_write(Grid, paste(s.dir, "Grid.gpkg", sep = '/'), append = FALSE)
```