```{r response stats, echo = FALSE}
Approached <- dat %>%
  distinct(PersonID, .keep_all = TRUE) 
nApproached <- nrow(Approached) # the number of people approached

nNE <- length(which(Approached$Agreement == "NE")) # number of people approached that were ineligible to participate in the survey
nAgree <- length(which(Approached$Agreement == "Yes")) # number of people approached who agreed to participate in the survey, including repeats 
nRefusals <- length(which(Approached$Agreement == "No")) # number of people approached who refused to take part in the survey, including repeats
nIndPeople <- length(which(Approached$Agreement %in% "Yes" & Approached$PrevInter %in% "No")) # the number of individual people surveyed
nMultiAgree <- nAgree - nIndPeople # number of multi-agreement surveys (total number of surveys repeated by previous participants - excluding their first survey)

nSpPoints <- nrow(dat) - nRefusals - nNE # the total number of spatial data points
nMultiUse <- nSpPoints - nAgree # the number of spatial point that were part of multi-use trips (excluding first instance)

RefusalRate <- round((nRefusals/(nApproached - nNE))*100) # response rate
ResponseRate <- round((nAgree/(nApproached - nNE))*100) # refusal rate
```
<br><br>

> MEG Mission statement: **Optimising human use of the ocean**

***
# Data {.tabset}
Data for recreational marine use in Esperance was collected through face-to-face surveys at boat launches.  The surveys facilitated the collection of multi-use data and multi-greement data.The survey had a **`r ResponseRate`% response rate** with **`r nAgree` participants**, of those **`r nIndPeople` survey were first instance or single agreement surveys** and **`r nMultiAgree` were multi-agreement surveys** (survey participant surveyed for a second or more times). This produced a **total of `r nSpPoints` spatial use data points**, **`r nMultiUse` of which were part of multi-use trips** (excluding first instance).

```{r data prep}
yes <- dat %>% 
  filter(Agreement == "Yes") # agreement data (filtering out refusals and non-eligibles)
bact <- yes %>% filter(SiteType == "Boat") # boat based (extractive and non-extractive)
sact <- yes %>% filter(SiteType == "Shore") # shore based (extractive and non-extractive)

ex <- yes %>% filter(Activity == "Fishing") # extractive - boat and shore
bex <- yes %>% filter(SiteType == "Boat" & ActivityType %in% c("Extractive", "Both")) # extractive - boat
sex <- yes %>% filter(SiteType == "Shore" & ActivityType %in% c("Extractive", "Both")) # extractive - shore

nex <- yes %>% filter(Activity != "Fishing") # non-extractive - boat and shore
bnex <- yes %>% filter(SiteType == "Boat" & ActivityType %in% c("Non-Extractive")) # non-extractive - boat
snex <- yes %>% filter(SiteType == "Shore" & ActivityType %in% c("Non-Extractive")) # non-extractive - shore
```

## Response 
The plot below shows the number of agreements, refusals and non-eligible people that were approached in the sampling process. 
```{r bar_response}
dat %>% 
  distinct(PersonID, .keep_all = TRUE) %>% 
  ggplot() +
   aes(x = fct_infreq(Agreement)) + 
   geom_bar(fill = col1) +
   labs(x = "Agreement to participate", y = paste("Frequency (n = ", nrow(Approached), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 
ggsave(paste(rumExplore_plots, "agree.png", sep='/'), width = 8, height = 4)
```

## Site type
The plot below shows the number of boat and shore based activities that were captured in the data.
```{r bar_sitetype}
yes %>% 
  ggplot() +
   aes(x = fct_infreq(SiteType)) +
   geom_bar(fill = col1) +
   labs(x = "Site Type", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 
 ggsave(paste(rumExplore_plots, "bar_sitetype.png", sep='/'), width = 8, height = 4)
```

## Year
The plot below shows the how the data is distributed across years. 
```{r bar_year}
yes %>% 
  ggplot()+
  aes(x = as.factor(facYear)) +
  geom_bar(fill = col1) +
  labs(x = "Year", y = paste("Number of Participants (n = ", nrow(yes), ")", sep = "")) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme

 ggsave(paste(rumExplore_plots, "bar_year.png", sep='/'), width = 8, height = 4)
```


# Demographics {.tabset}
***

## Locality
The plot below shows how many of the participants were residents and visitors.
```{r bar_locality}
ggplot(yes) + # get decending to work, add number of activities
   aes(x = reorder(Resident, desc(Resident))) +
   geom_bar(fill = col1) +
   labs(x = "Resident", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 
 ggsave(paste(rumExplore_plots, "bar_locality.png", sep='/'), width = 8, height = 4)
```

## Age
The plot below shows the ages of participants. This only includes my data, Jon did not record age. 
```{r bar_age}
yes %>% 
  filter(numYear>2018) %>% 
ggplot() +
   aes(x = binAge)+
   geom_bar(fill = col1) +
   labs(x = "Age of Participant", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 
 ggsave(paste(rumExplore_plots, "bar_age.png", sep='/'), width = 8, height = 4)
```

## Gender
The plot below shows the gender of participants.
```{r bar_gender}
ggplot(yes) + 
   aes(x = fct_infreq(Sex)) +
   geom_bar(fill = col1) +
   labs(x = "Sex of Participant", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 
 ggsave(paste(rumExplore_plots, "bar_gender.png", sep='/'), width = 8, height = 4)
```

## Fishing Avidity
The plot below shows the fishing avidity of participants.
```{r bar_exavid}
ggplot(yes) + 
   aes(x = bin_exnTimes12m) +
   geom_bar(fill = col1) +
   labs(x = "Number of times fished in last 12 months", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 
 ggsave(paste(rumExplore_plots, "bar_exavid.png", sep='/'), width = 8, height = 4)
```

# Use
***


## Overall {.tabset}
This map shows the kernal density of all boat based and shore based recreational uses. 
```{r kdens_all}
# boat based
a <- bact %>% 
  ggplot() +
  stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
  scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
  scale_alpha(guide = 'none') +
  geom_sf(data = coast, lwd = 0.07) +
  map_theme +
  labs(title = "Boat based activities")

ggsave(paste(rumExplore_plots, "kdens_bact.png", sep='/'), width = 4, height = 6)

# shore based
b <- sact %>% 
  ggplot() +
  stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
  scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
  scale_alpha(guide = 'none') +
  geom_sf(data = coast, lwd = 0.07) +
  map_theme
  labs(title = "Shore based activities")

ggsave(paste(rumExplore_plots, "kdens_sact.png", sep='/'), width = 4, height = 6)

a + b
ggsave(paste(rumExplore_plots, "kdens_bsact.png", sep='/'), width = 4, height = 6)
```

### Activity Type {.tabset}
The plots below shows how many surveys are relevant to extractive and non-extractive activities and there spatial distribution.
```{r bar_sp_acttype}
a <- ggplot(yes) +
   aes(x = fct_infreq(ActivityType)) +
   geom_bar(fill = col1) +
   labs(x = "Activity Type", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave(paste(rumExplore_plots, "bar_acttype.png", sep='/'), width = 8, height = 4)
 
b <- basemap +
  geom_sf(data = spdat, aes(color = ActivityType), size = 0.5) +
  scale_color_brewer(palette = "YlGnBu", name = " Activity Type") +
  map_theme
ggsave(paste(rumExplore_plots, "sp_acttype.png", sep='/'), width = 8, height = 4)

a + b
ggsave(paste(rumExplore_plots, "bar_sp_acttype.png", sep='/'), width = 8, height = 4)
```


### Activities
The plots below shows what activities participants conducted and there spatial distribution.
```{r bar_actfreq}
a <- yes %>% 
  ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
  geom_bar(stat="count", fill = col1) +
  labs(x = paste("Frequency (n = ", nrow(yes), ")", sep = ""), y = paste("Activity (n = ", length(na.omit(unique(yes$Activity))), ")", sep="")) +
  geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5)+
  my_theme
ggsave(paste(rumExplore_plots, "bar_actfreq.png", sep='/'), width = 8, height = 4)

b <- basemap +
  geom_sf(data = spdat, aes(color = Activity), size = 0.7) +
  scale_color_brewer(palette = "Set3") +
  map_theme
ggsave(paste(rumExplore_plots, "sp_actfreq.png", sep='/'), width = 4, height = 6)

a + b
ggsave(paste(rumExplore_plots, "bar_sp_actfreq.png", sep='/'), width = 4, height = 6)
```


## Extractive use 
***


### Boat based {.tabset}
#### Bar: Fishing type
```{r bar_bex}
bex %>%
 ggplot(aes(x = fct_infreq(FishingType))) +
  geom_bar(stat='count', position = position_dodge2(reverse = TRUE), fill = col1) +
  labs(x = paste("Fishing Type (n = ", length(na.omit(unique(bex$FishingType))), ")", sep=""), y = paste("Frequency (n = ", nrow(bex), ")", sep = "")) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme

ggsave(paste(rumExplore_plots, "bar_bex.png", sep='/'), width = 8, height = 4)
```

#### Spatial: fishing type and kernal density
```{r sp_bex}
sp_bex <- bex %>% 
  st_as_sf(coords = c("UseLong", "UseLat"), crs = 4283) %>% # making sf of boat based fishing
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2]) # need to keep Lat and Long independent of geometry

a <- basemap +
  geom_sf(data = sp_bex, aes(colour = FishingType), size = 0.5) +
  scale_color_brewer(palette = "Set3") +
  labs(color='Boat based fishing') +
  map_theme
ggsave(paste(rumExplore_plots, "sp_bex.png", sep='/'), width = 4, height = 6)

b <- sp_bex %>% 
  ggplot() +
  stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
  scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
  scale_alpha(guide = 'none') +
  geom_sf(data = coast, lwd = 0.07) +
  map_theme
ggsave(paste(rumExplore_plots, "kdens_bex.png", sep='/'), width = 4, height = 6)

a + b
ggsave(paste(rumExplore_plots, "sp_kdens_bex.png", sep='/'), width = 4, height = 6)
```


### Shore based {.tabset}
#### Bar: Fishing type
```{r bar_sex}
sex %>%
  ggplot(aes(x = fct_infreq(FishingType))) +
  geom_bar(stat='count', position = position_dodge2(reverse = TRUE), fill = col1) +
  labs(x = paste("Fishing Type (n = ", length(na.omit(unique(sex$FishingType))), ")", sep=""), y = paste("Frequency (n = ", nrow(sex), ")", sep = "")) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme

ggsave(paste(rumExplore_plots, "bar_sex.png", sep='/'), width = 8, height = 4)
```

#### Spatial: fishing type and kernal density
```{r sp_sex}
sp_sex <- st_as_sf(sex, coords = c("UseLong", "UseLat"), crs = 4283) %>% # making sf of boat based fishing
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2]) # need to keep Lat and Long independent of geometry

a <- basemap + 
  geom_sf(data = sp_sex, aes(color = FishingType), size = 1) +
  scale_color_brewer(palette = "Set3") +
   labs(color='Shore based fishing') +
  map_theme
ggsave(paste(rumExplore_plots, "sp_sex.png", sep='/'), width = 4, height = 6)

b <- sp_sex %>% 
  ggplot() +
  stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
  scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
  scale_alpha(guide = 'none') +
  geom_sf(data = coast, lwd = 0.07) +
  map_theme
ggsave(paste(rumExplore_plots, "kdens_sex.png", sep='/'), width = 4, height = 6)

a+b
ggsave(paste(rumExplore_plots, "sp_kdens_sex.png", sep='/'), width = 4, height = 6)
```


## Non-extractive use
***


### Boat Based {.tabset}
#### Bar: Actvity 
```{r bar_bnex}
bnex %>% 
  ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
  geom_bar(stat="count", fill = col1) +
  labs(x = paste("Frequency (n = ", nrow(bnex), ")", sep = ""), y = paste("Activity (n =", length(na.omit(unique(bnex$Activity))), ")", sep="")) +
  geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5) +
  my_theme

ggsave(paste(rumExplore_plots, "bar_bnex.png", sep='/'), width = 8, height = 4)
```

#### Spatial: Activity and kernal density
```{r sp_bnex}
sp_bnex <- st_as_sf(bnex, coords = c("UseLong", "UseLat"), crs = 4283) %>% # making sf of boat based non-extractive activities
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2]) # need to keep Lat and Long independent from geometry

a <- basemap +
  geom_sf(data = sp_bnex, aes(color = Activity), size = 1) + 
  scale_color_brewer(palette = "Set3") +
  map_theme
ggsave(paste(rumExplore_plots, "sp_bnex.png", sep='/'), width = 4, height = 6)

b <- sp_bnex %>% 
  ggplot() +
  stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
  scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
  scale_alpha(guide = 'none') +
  geom_sf(data = coast, lwd = 0.07) +
  map_theme
ggsave(paste(rumExplore_plots, "kdens_bnex.png", sep='/'), width = 4, height = 6)

a + b
ggsave(paste(rumExplore_plots, "sp_kdens_bnex.png", sep='/'), width = 4, height = 6)
```


### Shore Based {.tabset}
#### Bar: Activity 
```{r bar_snex}
snex %>% 
  ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
  geom_bar(stat="count", fill = col1) +
  labs(x = paste("Frequency (n = ", nrow(snex), ")", sep = ""), y = paste("Activity (n =", length(na.omit(unique(snex$Activity))), ")", sep="")) +
  geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5) +
  my_theme

ggsave(paste(rumExplore_plots, "bar_snex.png", sep='/'), width = 8, height = 4)
```

#### Spatial: Activity and kernal density
```{r sp_snex}
sp_snex <- st_as_sf(snex, coords = c("UseLong", "UseLat"), crs = 4283) %>% 
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2])

a <- basemap +
  geom_sf(data = sp_snex, aes(color = Activity), size = 1) + 
  scale_color_brewer(palette = "Set3") +
  map_theme
ggsave(paste(rumExplore_plots, "sp_snex.png", sep='/'), width = 4, height = 6)

b <- snex %>% 
  ggplot() +
  stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
  scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
  scale_alpha(guide = 'none') +
  geom_sf(data = coast, lwd = 0.07) +
  map_theme
ggsave(paste(rumExplore_plots, "kdens_snex.png", sep='/'), width = 4, height = 6)

a + b
ggsave(paste(rumExplore_plots, "sp_kdens_snex.png", sep='/'), width = 4, height = 6)
```