This script is filled with test pieces of code, drafts of code and redundant coee I know long need by might be useful for future reference, 

## Making maps Cape and CB - attempted insets
```{r spDataPrep}
## Setting up spatial layers
## crs1 <- 4283 is it useful to give a crs a name ?? GDA94
# Data 
# spNing <- st_as_sf(N1, coords = c("UseLong", "UseLat"), ## Won't work if there are blanks in df so  filtered
#                   crs = 4283)
# 
# # Coast
# WA <- st_read(paste(s.dir, "waCoast.shp", sep = '/')) %>%
#   st_transform(crs = 4283)
# 
# NWS <- st_crop(WA, xmin = 113.00, ymin = -24.00, xmax = 115.00, ymax = -20.5) # what you want to crop, masked onto what

# # 
# # # cropping sNing to CB/cape extent
# # CBning <- st_crop(sNing, CB)
# # #Capening <- st_crop(sNing, Cape)
# # Capening <- st_crop(sNing, Cape)
# 
# ## BR
# ausBR <- st_read(paste(s.dir, "ausRamps.shp", sep = '/')) %>%
#   st_transform(crs = 4283)
# 
# BR <- st_crop(ausBR, sNing)
# # CBR <- st_crop(BR, CB)
# # CapeBR <- st_crop(BR, Cape)
# 
# 
#  #### Bathy
# bathy <- st_read(paste(s.dir, "waDepths.shp", sep = '/')) %>%
#   st_transform(crs = 4283) %>%
#   st_make_valid()
# 
# bathy
# 
# NWSbathy <- st_crop(bathy, NWS)
# CBbathy <- st_crop(NWSbathy, CB)
# Capebathy <- st_crop(NWSbathy, Cape)

# ggplot() +
#   geom_sf(data = NWSbathy, lwd = 0.1, fill = "white") +
#   geom_sf(data = NWS, lwd = 0.1) +
#   geom_sf(data = sNing)
# 
# # summary(Ning$UseLat)
# # max(Ning$UseLat, na.rm = T)
# # which(Ning$UseLat == "-21.06476")
# # Ning[567,]
# # max(N1$UseLat, na.rm = T)
# 
# NWS xmin: 113.4615 ymin: -24.0027 xmax: 115 ymax: -21.4219
# NWSbathy xmin: 113.4615 ymin: -24.0027 xmax: 115 ymax: -21.4219
# sNing xmin: 113.2797 ymin: -23.46818 xmax: 114.7503 ymax: -21.06476
# 
# 113, -24, 115, -21
# 
# NWS
```

```{r}
CB <- st_crop(WA, xmin = 113.6271, xmax = 114.000, ymin = -23.46973, ymax = -22.562)
Cape <- st_crop(WA, xmin = 113.6271, ymin = -22.562, xmax = 114.7503, ymax = -21.56369)

Making bbox for CB
CBbbox = st_as_sfc(st_bbox(CB))

# Making CB inset
CBinset <- ggplot() +
  geom_sf(data = NWS, lwd = 0.1) +
  geom_sf(data = BR, color = "mediumaquamarine") +
  #geom_sf_text(data = BR, aes(label = boat_ramp), size = 2, nudge_x = 0.15) +
  geom_sf_label(data = BR, aes(label = boat_ramp), size = 1, nudge_x = 0.15, nudge_y = 0.02, label.padding =
                  unit(0.5, "mm"),label.size = 0) +
  labs(x = "", y = "") +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title=element_blank()) +
  geom_sf(data = CBbbox, fill = NA, color = "mediumaquamarine", size = 1)

CBinset

 ggsave(paste(rumPlots, "spBR.png", sep='/'), width = 8, height = 4)
 
##### Spatial plot for all activties at CB 
CBspAct <- ggplot() +
  geom_sf(data = CB, lwd = 0.1) +
  geom_sf(data = CBning, aes(color = Activity, shape = ActivityType), size = 1) + # colour for discrete, fill for continuous
  scale_color_brewer(palette = "Set3", direction = -1) + # colour for discrete, fill for continuous
  labs(caption = "Spatial distribution of marine recreational use", title = "Coral Bay") +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.5, "cm"), pad_y = unit(1, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm")) +
  theme(axis.text.x = element_text(angle = 90, size = 7),
        axis.text.y = element_text(size = 7),
        legend.key = element_rect(fill = "White", color = "White"),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 7, vjust = 1),
        legend.text = element_text(size = 7),
        title = element_text(size = 8))

CBspAct
ggsave(paste(rumPlots, "CBspAct.png", sep='/'), width = 5, height = 6)

# 
# CBspAct_inset = ggdraw() +
#   draw_plot(CBspAct) +
#   draw_plot(CBinset, x = 0.5, y = 0.05, width = 0.3, height = 0.2)
# 
# CBspAct_inset
# 
# ggsave(paste(rumPlots, "CBspAct_inset.png", sep='/'), width = 5, height = 6)
```

```{r}
##### Spatial plot for all activties at CB 
CapespAct <- ggplot() +
  geom_sf(data = Cape, lwd = 0.1) +
  geom_sf(data = Capening, aes(color = Activity, shape = ActivityType), size = 1) + # colour for discrete, fill for continuous
  scale_color_brewer(palette = "Set3") + # colour for discrete, fill for continuous
  labs(caption = "Spatial distribution of marine recreational use") +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.5, "cm"), pad_y = unit(1, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm")) +
  theme(axis.text.x = element_text(angle = 90, size = 7),
        axis.text.y = element_text(size = 7),
        legend.key = element_rect(fill = "White", color = "White"),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 7, vjust = 1),
        legend.text = element_text(size = 7),
        title = element_text(size = 8))

CapespAct
ggsave(paste(rumPlots, "CapespAct.png", sep='/'), width = 5, height = 6)
```

## Violin/Raincloud plot
```{r}
ggviolin(
  data = dpdat,
  x = c("DP"),
  y = c("nHooked"),
  color = "#81caca",
  fill = "#ceeded",
  alpha = 1,
  xlab = "Presence of depredation",
  ylab = "Number of fish hooked",
  # draw_quantiles = c(0.25, 0.5, 0.75),
  add = "boxplot",
  add.params = list(fill = "#81caca", color = "#ceeded"),
  error.plot = "linerange",
  label = NULL,
  font.label = list(size = 8, color = "black"))

ggviolin(
  data = dpdat,
  x = c("facYear"),
  y = c("nDP", "nHooked"),
  color = "#81caca",
  fill = c("DP"),
  alpha = 1,
  xlab = "Year",
  ylab = "Nmber of fish depredated",
  # draw_quantiles = c(0.25, 0.5, 0.75),
  # facet.by = c("facYear"),
  add = "boxplot",
  add.params = list(fill = "#81caca", color = "#ceeded"),
  error.plot = "linerange",
  label = NULL,
  font.label = list(size = 8, color = "black"))

"#ceeded"

dpdat %>%
  ggplot(aes(x = factor(DP), y = nHooked, fill = factor(DP))) +
  ggdist::stat_halfeye(adjust = 0.5, justification = -0.2, .width = 0, point_colour = NA) +
  geom_boxplot(width = 0.12, outlier.colour = NA, alpha = 0.5) +
  ggdist::stat_dots(side = "left", justification = 1.1, binwidth = 0.25) +
  labs(y = "Number of fish hooked", x = "") +
  scale_x_discrete(breaks=c("1", "0"),
                      labels=c("", "")) +
  scale_fill_manual(values=c("#97c1a9", "#81caca"),
                    name="Depredation",
                    breaks=c("1", "0"),
                    labels=c("Present", "Absent")) +
  theme(axis.ticks.y = element_blank()) +
  coord_flip()
```

# %DP facetted by Year
## Grouped bar
```{r}

dpdat %>%
  group_by(Side, facYear) %>%
  summarise(TotalHooked = sum(nHooked, na.rm = TRUE),
                     TotalDP = sum(nDP, na.rm = TRUE),
            TotalperDP = (TotalDP/TotalHooked)*100) %>%
  gather(Total, Value, TotalHooked:TotalDP) %>%
  mutate(Total = as.factor(ifelse(Total %in% "TotalHooked", "Hooked", "Depredated"))) %>%
drop_na()  %>%
ggplot(aes(x = factor(1), y = Value, fill = Total)) +
  geom_bar(width = 1, stat = "identity", position = position_dodge2(reverse=TRUE)) +
  scale_fill_manual(values=c("light coral", "medium aquamarine")) +
  guides(fill = guide_legend(reverse = TRUE)) +
  geom_text(aes(label=Value), position=position_dodge2(width=1, reverse = TRUE), vjust=-0.25, size = 2) +
  labs(x = "", y = "Number of fish") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) 
  facet_grid(Side~facYear)
```
## Pie chart

## Averages
Barplot showing the Average number of fish depredated per site facetted by year. 
```{r avDPYr}
# Average number of fish depredated per year
avDPYr <- exdat %>%
  filter(!is.na(nHooked),  nHooked != "", !is.na(nDP), nDP != "") %>%
  group_by(facYear) %>%
  summarise(ave = mean(nDP)) %>% 
  # mutate(numYear = as.character(Year)) %>%
  ggplot( 
  aes(x = facYear, y = ave)) +
  geom_col(fill = "mediumaquamarine") +
  labs(x = "Year", y = "Average number of fish depredated per site")
  
avDPYr
# ggsave(paste(rumPlots, "avDPYr.png", sep='/'), width = 8, height = 4)
```

## Catch rate
Barplot showing the average catch rate per site facetted by year. 
```{r CR}
CR <- exdat1 %>%
  group_by(facYear) %>%
  summarise(avCR = mean(CaughtUndam)) 
%>% 
CR

# Hist of nDP
avCR <- ggplot(CR) +
  aes(y = avCR, x = facYear) +
  geom_col(fill = "mediumaquamarine") +
  labs(x = "Average catch rate per site", y = "Year")

avCR
# ggsave(paste(rumPlots, "avCR.png", sep='/'), width = 8, height = 4)
```

# CAoe and CB spatial plots for depredation
```{r spDP_CB}
# Make an extent for coral bay and run the above code
## st of just nDP == 0
CBsexdp0 <- st_crop(sexdp0, CB)

## st  nDP > 0
CBsexdp <- st_crop(sexdp, CB)

CBspexdp <- ggplot() +
  geom_sf(data = CB, lwd = 0.1) +
  geom_sf(data = CBsexdp0, shape = 1, size = 0.5, color= "grey") +
  geom_sf(data = CBsexdp, aes(color = perDP), size = 1) + # colour for discrete, fill for continuous
  scale_color_distiller(palette = "YlGnBu", trans = "reverse") + # colour for discrete, fill for continuous
  scale_fill_continuous(trans = 'reverse') +
  labs(color = "Percentage\nDepredated", caption = "Points indicate sites of recreational fishing coloured by percentage depredation.\nSites denoted by hollow grey circle are sites where zero depredation occured.", title = "Coral Bay") +
  theme(axis.text.x = element_text(angle = 90, size = 5),
        axis.text.y = element_text(size = 5),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 5, vjust = 1),
        legend.text = element_text(size = 5),
        legend.key.height = unit(25, "mm"),
        title = element_text(size = 8)) +
  guides(color = guide_colorbar(reverse = TRUE)) +
  facet_wrap(. ~ Year) +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm"))

CBspexdp

# ggsave(paste(rumPlots, "CBspexdp.png", sep='/'), width = 4, height = 6)
```

```{r spDP_Cape}
# Cape
## st of just nDP == 0
Capesexdp0 <- st_crop(sexdp0, Cape)

## st  nDP > 0
Capesexdp <- st_crop(sexdp, Cape)

Capespexdp <- ggplot() +
  geom_sf(data = Cape, lwd = 0.1) +
  geom_sf(data = Capesexdp0, shape = 1, size = 0.5, color= "grey") +
  geom_sf(data = Capesexdp, aes(color = perDP), size = 1) + # colour for discrete, fill for continuous
  scale_color_distiller(palette = "YlGnBu", trans = "reverse") + # colour for discrete, fill for continuous
  scale_fill_continuous(trans = 'reverse') +
  labs(color = "Percentage\nDepredated", caption = "Points indicate sites of recreational fishing coloured by percentage depredation.\nSites denoted by hollow grey circle are sites where zero depredation occured.") +
  theme(axis.text.x = element_text(angle = 90, size = 5),
        axis.text.y = element_text(size = 5),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 5, vjust = 1),
        legend.text = element_text(size = 5),
        legend.key.height = unit(25, "mm"),
        title = element_text(size = 8)) +
  guides(color = guide_colorbar(reverse = TRUE)) +
  facet_wrap(. ~ Year) +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm"))

Capespexdp
# 
# ggsave(paste(rumPlots, "Capespexdp.png", sep='/'), width = 4, height = 6)
```

```{r}
 a <- st_read(paste(s.dir, "gpkg/FADs.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "FADs.shp", sep = '/'), append = F)

 
  a <- st_read(paste(s.dir, "gpkg/ncmpSZ.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "ncmpSZ.shp", sep = '/'), append = F)
 
   a <- st_read(paste(s.dir, "gpkg/NingBR.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "NingBR.shp", sep = '/'), append = F)
 
 
    a <- st_read(paste(s.dir, "gpkg/nmpSpPurposeShore.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "nmpSpPurposeShore.shp", sep = '/'), append = F)
 
     a <- st_read(paste(s.dir, "gpkg/NWScoast.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "NWScoast.shp", sep = '/'), append = F)
 
  a <- st_read(paste(s.dir, "gpkg/NWScoast.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "NWScoast.shp", sep = '/'), append = F)
 
  
  a <- st_read(paste(s.dir, "gpkg/ReefHabitat.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "Reef.shp", sep = '/'), append = F)
 
 
  a <- st_read(paste(s.dir, "gpkg/Wells.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "shp/Wells.shp", sep = '/'), append = F)
 
   a <- st_read(paste(s.dir, "gpkg/Pipes.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)
 st_write(a, paste(s.dir, "shp/Pipes.shp", sep = '/'), append = F)
```

```{r}
FTyp <- exdat %>% 
  group_by(FishingType) %>%   
  mutate(count_name_occurr = n()) %>%
  filter(!is.na(FishingType), FishingType != "") %>% 
  ggplot(
  aes(x=reorder(FishingType,-count_name_occurr))) + #reorder(col to reorder, by what var)
  geom_bar(stat='count', fill = "#8fcaca") +
  labs(x = paste("Fishing Type (n = ", length(na.omit(unique(exdat$FishingType))), ")", sep=""), y = paste("Frequency (n = ", nrow(exdat), ")", sep = "")) +
  # facet_wrap(. ~ numYear) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5)

FTyp
ggsave(paste(rumPlots, "FTyp.png", sep='/'), width = 8, height = 4)


bf_FTyp <- bf %>% 
  group_by(FishingType) %>%   
  mutate(count_name_occurr = n()) %>%
  filter(!is.na(FishingType), FishingType != "") %>% 
  ggplot(
  aes(x=reorder(FishingType,-count_name_occurr))) + #reorder(col to reorder, by what var)
  geom_bar(stat='count', fill = "#8fcaca") +
    labs(x = paste("Boat based fishing type (n = ", length(na.omit(unique(bf$FishingType))), ")", sep=""), y = paste("Frequency (n = ", nrow(bf), ")", sep = "")) +
  # facet_wrap(. ~ numYear) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5)

bf_FTyp

ggsave(paste(rumPlots, "bf_FTyp.png", sep='/'), width = 8, height = 4)

```

```{r}
a <- dpdat %>%
  group_by(PersonID, TripNum, facYear) %>% 
  summarise(hookedTrip = sum(nHooked),
            DPTrip = sum(nDP), 
            per = DPTrip/hookedTrip*100, # percentage of depredation per trip 
            bin = ifelse(per > 0, 1, 0),
            bin = ifelse(hookedTrip == 0 & nDP == 0, 0, bin)) 

perDPTrip <- a %>% 
  group_by(facYear) %>% 
  summarise(Percentage = sum(bin==1)/n()*100) %>% 
  mutate(label = paste(round(Percentage, digits = 1), "%", sep =" "),
         data = "Trips")


perDPSite <- dpdat %>%
  group_by(facYear) %>%
  summarise(Percentage = (sum(DP == 1)/n())*100) %>% 
    mutate(label = paste(round(Percentage, digits = 1), "%", sep =" "),
           data = "Sites") 

c <- bind_rows(perDPTrip, perDPSite) %>% 
  ggplot(aes(factor(1), y = Percentage, fill = data)) +
   geom_bar(stat = "identity", position = position_dodge2(reverse = TRUE)) +
  labs(y = "Number of fish", x="") +
  guides(fill = guide_legend(reverse = TRUE)) +
  scale_fill_manual(values=c("#97c1a9", "#81caca"), name = "") +
   theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  geom_text(aes(label=label), vjust = -0.5, size = 2.5, position = position_dodge2(.9, reverse = TRUE)) +
  facet_grid(~facYear)
c
```

```{r}
var_names <- c('lat', 'lon', 'time', 'sea_surface_temperature', 'sst_dtime', 'dt_analysis', 'satellite_zenith_angle', 'l2p_flags')

sst_f1 <- nc_open(paste(cdf.dir, 'SST_nc4_f1/IMOS_aggregation_20211209T025615Z.nc', sep = "/"), write = TRUE)
sst_f2 <- nc_open(paste(cdf.dir, 'SST_nc4_f2/IMOS_aggregation_20211209T025716Z.nc', sep = "/"), write = TRUE)
sst_f3 <- nc_open(paste(cdf.dir, 'SST_nc4_f3/IMOS_aggregation_20211209T025817Z.nc', sep = "/"), write = TRUE)
sst_f4 <- nc_open(paste(cdf.dir, 'SST_nc4_f4/IMOS_aggregation_20211209T030015Z.nc', sep = "/"), write = TRUE)
sst_f5 <- nc_open(paste(cdf.dir, 'SST_nc4_f5/IMOS_aggregation_20211209T033329Z.nc', sep = "/"), write = TRUE)
sst_f6 <- nc_open(paste(cdf.dir, 'SST_nc4_f6/IMOS_aggregation_20211209T033413Z.nc', sep = "/"), write = TRUE)
# Just for one variable for now
dat_new <- cbind(
  ncvar_get(sst_f1, var_names),
  ncvar_get(sst_f2, var_names),
  ncvar_get(sst_f3, var_names),
  ncvar_get(sst_f4, var_names),
  ncvar_get(sst_f5, var_names),
  ncvar_get(sst_f6, var_names))
dim(dat_new)
var <- sst_f1$var[var_names]$varnames

# Create a new file
file_new3 <- nc_create(
  filename = 'sst_f2f6.nc',
  # We need to define the variables here
  vars = ncvar_def(
    name = 'var_names',
    units = var$units,
    dim = dim(dat_new)))

# And write to it
ncvar_put(
  nc = file_new,
  varid = 'var_names',
  vals = dat_new)

# Finally, close the file
nc_close(file_new)
# 
# 
# 
# 
# 
var_names <- c('lat', 'lon', 'time', 'sea_surface_temperature', 'sst_dtime', 'dt_analysis', 'satellite_zenith_angle', 'l2p_flags')

for (var_name in var_names) {

  # Create raster stack
  x <- stack(
    raster(sst_f1n, varname = var_name),
    raster(sst_f2n, varname = var_name),
    raster(sst_f3n, varname = var_name),
    raster(sst_f4n, varname = var_name),
    raster(sst_f5n, varname = var_name),
    raster(sst_f6n, varname = var_name))

  # Name each layer
  names(x) <- c("f1", "f2", "f3", "f4", "f5", "f6")

  writeRaster(x = x,
              filename = paste0(var_name, '_out.nc'),
              overwrite = TRUE,
              format = 'CDF')
}

s.dir <- paste(w.dir, "spIgnore/shp", sep='/')
gpkg.dir <- paste(w.dir, "spIgnore/gpkg", sep='/')

Esp_ss <- st_read(paste(gpkg.dir, "Esp_StudySite.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

st_write(Esp_ss, paste(s.dir, "Esp_StudySite.shp", sep = '/'))
```

```{r}

# 12 days
# 2 units of free time 
# 60% chamce of getting Esperance 
# 40% chance of getting Duke
# 
# 
runif(3, 1, 8)
sample(c('Esperance','Duke'), size=6, replace=T, prob=c(.8,.2))

sample(c('Esperance','Duke', 'Lucky', 'Alexander'), size=18, replace=T, prob=c(.4,.4, .1, .1))

```

