```{r librarys}
library(ggBRT)
library(gbm)
library(dismo)
library(MASS)
library(dplyr)
```

```{r dir, echo = FALSE}
w.dir <- "/Users/23088313/Documents/git_repos/Analysis-Hamre-Bioeconomic"
d.dir <- paste(w.dir, "Ch2-RUM/rumIgnore", sep='/')
f.dir <- paste(w.dir, "Ch2-RUM/rumFunc", sep = '/')
m.dir <- paste(w.dir, "Ch2-RUM/rumDP/models", sep = '/') # model directory
dpBRT <- paste(m.dir, "dpBRT", sep='/')
crBRT <- paste(m.dir, "crBRT", sep='/')
```


```{r functions}
source(paste(f.dir, "brtFunc.R", sep = '/'))
```

```{r read data}
dpdat <- read.csv(paste(d.dir, "dpDat_v1.csv", sep='/'))
```

# BRT {.tabset}
Boosted regression trees

The BRT approach differs fundamentally from traditional regression methods that produce a single ‘best’ model, instead using the technique of boosting to combine large numbers of relatively simple tree models adaptively, to optimize predictive performance. 

Learning rate (lr) determines the contribution of each tree to the growing model. Dedpeasing (slowing) lr indpeases the number of trees required, and in general a smaller lr (and larger nt) are preferable, conditional on the number of observations and time available for computation.

Tree complexity (tc) controls whether interactions are fitted. A tc of 1 (single decision stump; two terminal nodes) fits an additive model, a tc of two fits a model with up to two-way interactions, and so on. For a given lr, fitting more complex trees leads to fewer trees being required for minimum error. So, as tc is indpeased, lr must be dedpeased if sufficient trees are to be fitted.

As a general guide, lr needs to be dedpeased as tc indpeases, usually inversely: doubling tc should be matched with halving lr to give approximately the same nt. 

Number of trees (nt), the lr and tc determine the number of trees (nt) required for optimal prediction.

Bag fraction (bf), controls stochasticity that specifies the proportion of data to be selected at each step. The default bag fraction is 0·5, meaning that, at each iteration, 50% of the data are drawn at random, without replacement, from the full training set. In our experience, stochasticity improves model performance, and fractions in the range 0·5–0·75 have given best results for presence–absence responses. 

Step size, is the number of trees added at each step. 

Note that the data has to be in data.frame ONLY format - not a tibble or sf as well.

```{r}
# Making all character vars factors
dpdat <- dpdat %>%
  mutate_if(is.character, as.factor) %>% 
  mutate(facYear = as.factor(facYear))
```
## Depredation BRT 1 - prep overall

### Data Prep 
```{r data prep for BRTs}
# making a grid for step.loop
## make a res.dev variable to store in grid
res.dev<-c() 
dp_step.loop<-expand.grid(lr=c(0.01, 0.005, 0.0001), tc=c(2, 3, 5, 7), bf=c(0.5, 0.75, 0.9), ss=c(5, 20, 30, 50)) ## make grid of lr, tc and bag.fraction that gbm.step will run through
dp_step.loop$res.dev<-NA ## add res.dev to grid

# define response
nDP <- which(colnames(dpdat)=="nDP") 

# defining covariates
DPcov <- which(colnames(dpdat) %in% c("facYear", "Month", "Side", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "DecFishingHr","MaxHook",
                                   "exYrs","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst"))

# defining offset
# dp_offset = dpdat$nHooked
# Raw offset very zero inflated and doesnt run 
dp_offset_log = round(log(dpdat$nHooked +1))
# hist(log(dpdat$nHooked)+1)

#Set up CV
  k=10 #number of folds
  p<-unique(dpdat$PersonID)
  p<-data.frame(PersonID=(sample(p)))# shuffle the data
  p$k<-c(rep(1:k, floor(nrow(p)/k)), sample(1:k, size = nrow(p) -length(rep(1:k, floor(nrow(p)/k)))))
  dpdat<-left_join(dpdat, p, by="PersonID")
  table(dpdat$k) #k provides indicator for 10 fold cross validation
```

### gbm.step loop
Loop which runs through several parameters 
```{r dp.step loop, warning=FALSE}
# This first gbm.step is looping through parameters so i can sleect the optimal model
for(i in 1:nrow(dp_step.loop)) {
  dp.step<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov),
                           gbm.y=nDP, ## response - number of fish depredated
                           offset = dp_offset_log,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           lr = dp_step.loop[i,"lr"], ## ref to step.loop
                           tree.complexity = dp_step.loop[i,"tc"], ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction = dp_step.loop[i,"bf"],
                           step.size = dp_step.loop[i, "ss"],
                           plot.main = F)
  dp_step.loop[i, "res.dev"] <-dp.step$self.statistics$mean.resid ### store res.dev in step.loop
  dp_step.loop[i, "nt"] <-dp.step$n.trees} 
```

### Optimal parameters
Optimal parameters  which minimise the residual devience as defined by outputs of gbm.step loop
```{r optimal parameters, echo = FALSE}
# Full loop data set if intere
# dp_step.loop 
dp_opt.resdev <- dp_step.loop[dp_step.loop$res.dev==min(dp_step.loop$res.dev, na.rm= T),]
dp_opt.resdev <- dp_opt.resdev %>% drop_na()
dp_opt.resdev
dp_opt.nt <- dp_step.loop[dp_step.loop$nt==max(dp_step.loop$nt, na.rm= T),]
dp_opt.nt <- dp_opt.nt %>% drop_na()
dp_opt.nt

dp_opt <- full_join(dp_opt.resdev, dp_opt.nt)
dp_opt <- dp_opt %>% drop_na
dp_opt
```

Running gbm.step with optimal parameters and extracting n.trees. 
```{r dp.brt, warning=FALSE}

dp.brt<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov),
                           gbm.y=nDP,## response - number of fish depredated
                           offset=dp_offset,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = dp_opt.nt$lr, ## ref to step.loop
                           tree.complexity = dp_opt.nt$tc, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=dp_opt.nt$bf,
                           step.size = dp_opt.nt$ss)

dp.brt<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov),
                           gbm.y=nDP,## response - number of fish depredated
                           offset=dp_offset_log,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)

dp.brt_v1<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov),
                           gbm.y=nDP,## response - number of fish depredated
                           offset=dp_offset_log,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
# nt <- dp.step$n.trees # adds number of trees to step param
# dp_best$nt <- nt # defines nt for automation
# dp_best$res.dev  <- dp.step$self.statistics$mean.resid # updates residual devience
# dp_best
# 
dp.brt_v1$self.statistics$discrimination
```

### Outputs
#### Default outputs
```{r dp.brt_v1 output}

ggPerformance(dp.brt_v1)

dp.brt_influence_v1 <- ggInfluence(dp.brt_v1, signif = TRUE, main = "Significant predictors on number of fish depredated") 
ggsave(paste(dpBRT, "dp.brt_Influence_v1.png", sep='/'), width = 8, height = 4)
dp.brt_influence_v1

for(i in 1:nrow(dp.brt_influence_v1)) {
dp.brt_PD_v1 <- ggPD(dp.brt_v1, rug = T, smooth = T, predictor = i) 
ggsave(dp.brt_PD_v1, file = paste(dpBRT, paste("dp.brt_PD_v1", i, ".png"), sep = "/"), width = 8, height = 4)
}

for(i in 1:nrow(dp.brt_influence_v1)) {
dp.brt_PDFit_v1 <- ggPDfit(dp.brt_v1, smooth = T, se = T, predictor = i) 
ggsave(dp.brt_PDFit_v1, file = paste(dpBRT, paste("dp.brt_PDFit_v1", i, ".png"), sep = "/"), width = 8, height = 4)
}

 dp_interaction_v1 <- ggInteract_list(dp.brt_v1, index = F) 
 
 dp_interaction_v1
```

#### Bootstrapped cis
```{r dp.bootstrap, include=FALSE}

## This will not work if there is an offset.
# Might need something ike the code below to get it to work - but maybe not worth using this code to plot if there is an offset  
# dp.brt3$gbm.call$dataframe$nDP <- dp.brt3$gbm.call$dataframe$nDP + dp_offset[1:581]
# dp.brt3$fitted <- dp.brt3$fitted + dp_offset[1:581]

dp.brt_prerun_v1<- plot.gbm.4list(dp.brt_v1)

dp.brt_boot_v1 <- gbm.bootstrap.functions(dp.brt_v1, list.predictors=dp.brt_prerun_v1, n.reps=100)

for(i in 1:nrow(dp.brt_influence_v1)) {
dp.brt_PDboot_v1<-ggPD_boot(dp.brt_v1, list.4.preds=dp.brt_prerun_v1,
                booted.preds=dp.brt_boot_v1$function.preds, type.ci="ribbon", rug=T, smooth = T, predictor = i)
ggsave(dp.brt_PDboot_v1, file = paste(dpBRT, paste("dp.brt_PDboot_v1", i, ".png"), sep = "/"), width = 8, height = 4)
}
```

## Depredation BRT 1 - orginal 

response: nDP
family: poisson
offset: log(nHooked+1)
boat kernal density and boat within 5 km

```{r}
nDP <- which(colnames(dpdat)=="nDP") 

# defining covariates
DPcov <- which(colnames(dpdat) %in% c("facYear", "Month", "Side", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "DecFishingHr","MaxHook",
                                   "exYrs","exnTimes12m","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst", "BaitLure", "TripNum", "SiteNum"))



dp_offset_log = round(log(dpdat$nHooked +1))

dp.brt_v1<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov),
                           gbm.y=nDP,## response - number of fish depredated
                           offset=dp_offset_log,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
```

```{r}
ggPerformance(dp.brt_v1)

dp.brt_influence_v1 <- ggInfluence(dp.brt_v1, signif = TRUE, main = "Significant predictors on number of fish depredated", col.bar = "#97c1a9") 
ggsave(paste(dpBRT, "dp.brt_Influence_v1.png", sep='/'), width = 8, height = 4)
dp.brt_influence_v1


for(i in 1:nrow(dp.brt_influence_v1)) {
dp.brt_PD_v1 <- ggPD(dp.brt_v1, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_v1, file = paste(dpBRT, paste("dp.brt_PD_v1", i, ".png"), sep = "/"), width = 8, height = 4)
}

# for(i in 1:nrow(dp.brt_influence_v1)) {
# dp.brt_PDFit_v1 <- ggPDfit(dp.brt_v1, smooth = T, se = T, predictor = i) 
# ggsave(dp.brt_PDFit_v1, file = paste(dpBRT, paste("dp.brt_PDFit_v1", i, ".png"), sep = "/"), width = 8, height = 4)
# }

 dp_interaction_v1 <- ggInteract_list(dp.brt_v1, index = F) 
 
 dp_interaction_v1
```
## Depredation BRT 3 - West

```{r}
west <- dpdat %>% 
  filter(Side %in% "West")

nDP_west <- which(colnames(west)=="nDP") 

DPcov_west <- which(colnames(west) %in% c("facYear", "Month", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "DecFishingHr","MaxHook",
                                   "exYrs","exnTimes12m","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst", "BaitLure", "TripNum", "SiteNum"))

dp_offset_w = round(log(west$nHooked +1))

dp.brt_west<-dismo::gbm.step(data=west,
                           gbm.x=c(DPcov_west),
                           gbm.y=nDP_west,## response - number of fish depredated
                           offset = dp_offset_w,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
```

```{r}
ggPerformance(dp.brt_west)

dp.brt_influence_west <- ggInfluence(dp.brt_west, signif = TRUE, main = "Significant predictors on number of fish depredated", col.bar = "#97c1a9") 
ggsave(paste(dpBRT, "dp.brt_Influence_west.png", sep='/'), width = 8, height = 4)
dp.brt_influence_west

for(i in 1:nrow(dp.brt_influence_west)) {
dp.brt_PD_west <- ggPD(dp.brt_west, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_west, file = paste(dpBRT, paste("dp.brt_PD_west", i, ".png"), sep = "/"), width = 8, height = 4)
}

# for(i in 1:nrow(dp.brt_influence_west)) {
# dp.brt_PDFit_west <- ggPDfit(dp.brt_west, smooth = T, se = T, predictor = i) 
# ggsave(dp.brt_PDFit_west, file = paste(dpBRT, paste("dp.brt_PDFit_west", i, ".png"), sep = "/"), width = 8, height = 4)
# }

 dp_interaction_west <- ggInteract_list(dp.brt_west, index = F) 
 
 dp_interaction_west
```


east

```{r}

east <- dpdat %>% 
  filter(Side %in% "East")

nDP_east <- which(colnames(east)=="nDP") 
DPcov_east <- which(colnames(east) %in% c("facYear", "Month", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "DecFishingHr","MaxHook",
                                   "exYrs","exnTimes12m","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst", "BaitLure", "TripNum", "SiteNum"))

dp_offset_e = round(log(east$nHooked +1))

dp.brt_east<-dismo::gbm.step(data=east,
                           gbm.x=c(DPcov_east),
                           gbm.y=nDP_east,## response - number of fish depredated
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
```

```{r}
ggPerformance(dp.brt_east)

dp.brt_influence_east <- ggInfluence(dp.brt_east, signif = TRUE, main = "Significant predictors on number of fish depredated", col.bar = "#97c1a9") 
ggsave(paste(dpBRT, "dp.brt_Influence_east.png", sep='/'), width = 8, height = 4)
dp.brt_influence_east

for(i in 1:nrow(dp.brt_influence_east)) {
dp.brt_PD_east <- ggPD(dp.brt_east, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_east, file = paste(dpBRT, paste("dp.brt_PD_east", i, ".png"), sep = "/"), width = 8, height = 4)
}

for(i in 1:nrow(dp.brt_influence_east)) {
dp.brt_PDFit_east <- ggPDfit(dp.brt_east, smooth = T, se = T, predictor = i) 
ggsave(dp.brt_PDFit_east, file = paste(dpBRT, paste("dp.brt_PDFit_east", i, ".png"), sep = "/"), width = 8, height = 4)
}

 dp_interaction_east <- ggInteract_list(dp.brt_east, index = F) 
 
 dp_interaction_east
```

```{r}
dp_dem <- dpdat %>% 
  filter(FishingType %in% "Demersal")

nDP_dem <- which(colnames(dp_dem)=="nDP") 

# defining covariates
DPcov_dem <- which(colnames(dp_dem) %in% c("facYear", "Month", "Side", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","DecFishingHr","MaxHook",
                                   "exYrs","exnTimes12m","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst", "BaitLure", "TripNum", "SiteNum"))



dp_offset_dem = round(log(dp_dem$nHooked +1))

dp.brt_dem<-dismo::gbm.step(data=dp_dem,
                           gbm.x=c(DPcov_dem),
                           gbm.y=nDP_dem,## response - number of fish depredated
                           offset=dp_offset_dem,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
```
```{r}
ggPerformance(dp.brt_dem)

dp.brt_influence_dem <- ggInfluence(dp.brt_dem, signif = TRUE, main = "Significant predictors on number of fish depredated", col.bar = "#97c1a9") 
ggsave(paste(dpBRT, "dp.brt_Influence_dem.png", sep='/'), width = 8, height = 4)
dp.brt_influence_dem

for(i in 1:nrow(dp.brt_influence_dem)) {
dp.brt_PD_dem <- ggPD(dp.brt_dem, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_dem, file = paste(dpBRT, paste("dp.brt_PD_dem", i, ".png"), sep = "/"), width = 8, height = 4)
}

# for(i in 1:nrow(dp.brt_influence_dem)) {
# dp.brt_PDFit_dem <- ggPDfit(dp.brt_dem, smooth = T, se = T, predictor = i) 
# ggsave(dp.brt_PDFit_dem, file = paste(dpBRT, paste("dp.brt_PDFit_dem", i, ".png"), sep = "/"), width = 8, height = 4)
# }

 dp_interaction_dem <- ggInteract_list(dp.brt_dem, index = F) 
 
 dp_interaction_dem
```

```{r}
east_dem <- east %>% 
  filter(FishingType %in% "Demersal")

dp_east_dem <- which(colnames(east_dem)=="nDP") 

# defining covariates
DPcov_east_dem <- which(colnames(east_dem) %in% c("facYear", "Month", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","DecFishingHr","MaxHook",
                                   "exYrs","exnTimes12m","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst", "BaitLure", "TripNum", "SiteNum"))



dp_offset_east_dem = round(log(east_dem$nHooked +1))

dp.brt_east_dem<-dismo::gbm.step(data=east_dem,
                           gbm.x=c(DPcov_east_dem),
                           gbm.y=dp_east_dem,## response - number of fish depredated
                           offset=dp_offset_east_dem,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
```

```{r}
ggPerformance(dp.brt_east_dem)

dp.brt_influence_east_dem <- ggInfluence(dp.brt_east_dem, signif = TRUE, main = "Significant predictors on number of fish depredated", col.bar = "#97c1a9") 
ggsave(paste(dpBRT, "dp.brt_Influence_east__dem.png", sep='/'), width = 8, height = 4)
dp.brt_influence_east_dem

for(i in 1:nrow(dp.brt_influence_east_dem)) {
dp.brt_PD_east_dem <- ggPD(dp.brt_east_dem, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_east_dem, file = paste(dpBRT, paste("dp.brt_PD_east_dem", i, ".png"), sep = "/"), width = 8, height = 4)
}

# for(i in 1:nrow(dp.brt_influence_dem)) {
# dp.brt_PDFit_dem <- ggPDfit(dp.brt_dem, smooth = T, se = T, predictor = i) 
# ggsave(dp.brt_PDFit_dem, file = paste(dpBRT, paste("dp.brt_PDFit_dem", i, ".png"), sep = "/"), width = 8, height = 4)
# }

 dp_interaction_east_dem <- ggInteract_list(dp.brt_east_dem, index = F) 
 
 dp_interaction_east_dem
 
 names(dpdat)
```
```{r}

dpdat$proDP <- dpdat$nDP/100

proDP <- which(colnames(dpdat)=="proDP") 

# defining covariates
DPcov <- which(colnames(dpdat) %in% c("facYear", "Month", "Side", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "DecFishingHr","MaxHook","exnTimes12m","BoatLength", "LunarPhase", "b_kdens", "b_5km", "sst", "BaitLure", "TripNum", "SiteNum"))



dp_offset_log = round(log(dpdat$nHooked +1))

dp.brt_ber<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov),
                           gbm.y=proDP,## response - number of fish depredated
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="bernoulli", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)

ggPerformance(dp.brt_ber)

dp.brt_influence_east_dem <- ggInfluence(dp.brt_ber, signif = TRUE, main = "Significant predictors on number of fish depredated", col.bar = "#97c1a9") 
ggsave(paste(dpBRT, "dp.brt_Influence_east__dem.png", sep='/'), width = 8, height = 4)
dp.brt_influence_east_dem

for(i in 1:nrow(dp.brt_influence_east_dem)) {
dp.brt_PD_east_dem <- ggPD(dp.brt_ber, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_east_dem, file = paste(dpBRT, paste("dp.brt_PD_east_dem", i, ".png"), sep = "/"), width = 8, height = 4)
}

# for(i in 1:nrow(dp.brt_influence_dem)) {
# dp.brt_PDFit_dem <- ggPDfit(dp.brt_dem, smooth = T, se = T, predictor = i) 
# ggsave(dp.brt_PDFit_dem, file = paste(dpBRT, paste("dp.brt_PDFit_dem", i, ".png"), sep = "/"), width = 8, height = 4)
# }

 dp_interaction_east_dem <- ggInteract_list(dp.brt_ber, index = F) 
 
 dp_interaction_east_dem
```

## Depredation BRT 2 - fishing vessels vs vessels
response: nDP
family: poisson
offset: log(nHooked+1)
fishing boat kernal density and boat within 5 km
```{r}
DPcov_v2 <- which(colnames(dpdat) %in% c("facYear", "Month", "Side", "Resident",
                                   "nTimesLast24m","exDecMedianTime",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "DecFishingHr","MaxHook",
                                   "exYrs","BoatLength", "LunarPhase", "bf_kdens", "bf_5km", "sst"))

dp.brt_v2<-dismo::gbm.step(data=dpdat,
                           gbm.x=c(DPcov_v2),
                           gbm.y=nDP,## response - number of fish depredated
                           offset=dp_offset_log,
                           fold.vector = dpdat$k,
                           n.folds = k,
                           learning.rate = 0.001, ## ref to step.loop
                           tree.complexity = 4, ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction=0.5,
                           step.size = 50)
```

```{r}
ggPerformance(dp.brt_v2)

dp.brt_influence_v2 <- ggInfluence(dp.brt_v2, signif = TRUE, main = "Significant predictors on number of fish depredated") 
ggsave(paste(dpBRT, "dp.brt_Influence_v2.png", sep='/'), width = 8, height = 4)
dp.brt_influence_v2

for(i in 1:nrow(dp.brt_influence_v2)) {
dp.brt_PD_v2 <- ggPD(dp.brt_v2, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_v2, file = paste(dpBRT, paste("dp.brt_PD_v2", i, ".png"), sep = "/"), width = 8, height = 4)
}

for(i in 1:nrow(dp.brt_influence_v2)) {
dp.brt_PDFit_v2 <- ggPDfit(dp.brt_v2, smooth = T, se = T, predictor = i) 
ggsave(dp.brt_PDFit_v2, file = paste(dpBRT, paste("dp.brt_PDFit_v2", i, ".png"), sep = "/"), width = 8, height = 4)
}

 dp_interaction_v2 <- ggInteract_list(dp.brt_v2, index = F) 
 
 dp_interaction_v2
```


```{r}
ggPerformance(dp.brt_east)

dp.brt_influence_east <- ggInfluence(dp.brt_east, signif = TRUE, main = "Significant predictors on number of fish depredated") 
ggsave(paste(dpBRT, "dp.brt_Influence_east.png", sep='/'), width = 8, height = 4)
dp.brt_influence_east

for(i in 1:nrow(dp.brt_influence_east)) {
dp.brt_PD_east <- ggPD(dp.brt_east, rug = T, smooth = T, predictor = i, col.line = "#8fcaca", col.smooth = "#97c1a9") 
ggsave(dp.brt_PD_east, file = paste(dpBRT, paste("dp.brt_PD_east", i, ".png"), sep = "/"), width = 8, height = 4)
}

for(i in 1:nrow(dp.brt_influence_east)) {
dp.brt_PDFit_east <- ggPDfit(dp.brt_east, smooth = T, se = T, predictor = i) 
ggsave(dp.brt_PDFit_east, file = paste(dpBRT, paste("dp.brt_PDFit_east", i, ".png"), sep = "/"), width = 8, height = 4)
}

 dp_interaction_east <- ggInteract_list(dp.brt_east, index = F) 
 
 dp_interaction_east
```


## Catch rate BRT
### Data Prep 
```{r data prep for BRTs}
# making a grid for step.loop
## make a res.dev variable to store in grid
res.dev<-c() 
cr_step.loop<-expand.grid(lr=c(0.1, 0.01, 0.001), tc=c(2, 3, 5), bf=c(0.5, 0.75, 0.9), ss=c(20, 30, 50)) ## make grid of lr, tc and bag.fraction that gbm.step will run through
cr_step.loop$res.dev<-NA ## add res.dev to grid
cr_step.loop$nt <- NA

# define response
cr <- which(colnames(dpdat)=="CaughtUndam") 

# defining covariates
crcov <- which(colnames(dpdat) %in% c("numYear", "Month", "Side", "Resident",
                                  "exDecMedianTime", "exnTimes12m",
                                   "Depth", "Habitat","DistNearestBRkm", "DistNearestInfra", "UseLat",
                                   "UseLong","FishingType", "BaitLure", "DecFishingHr","MaxHook",
                                   "exYrs","BoatLength", "LunarPhase", "b_kdens", "b_5km"))

# add number of times 12 months 

```

### gbm.step loop
Loop which runs through several parameters 
```{r cr.step loop, warning=FALSE}
# This first gbm.step is looping through parameters so i can select the optimal model
for(i in 1:nrow(cr_step.loop)) {
  cr.step<-dismo::gbm.step(data=crdat,
                           gbm.x=c(crcov),
                           gbm.y=cr, ## response - number of fish caught undamaged
                           lr = cr_step.loop[i,"lr"], ## ref to step.loop
                           tree.complexity = cr_step.loop[i,"tc"], ## ref to step.loop
                           family="poisson", ## distribution family
                           bag.fraction = cr_step.loop[i,"bf"],
                           step.size = cr_step.loop[i, "ss"],
                           plot.main = F)
  cr_step.loop[i, "res.dev"] <-cr.step$self.statistics$mean.resid ### store res.dev in step.loop
  cr_step.loop[i, "nt"]<- cr.step$n.trees} ### store n.trees in step.loop
```

### Optimal parameters
Optimal parameters  which minimise the residual devience as defined by outputs of gbm.step loop
```{r step parameters, echo = FALSE}
cr_step.loop # all gbm outputs
cr_best1 <- cr_step.loop[cr_step.loop$res.dev==min(cr_step.loop$res.dev, na.rm= T),] # output with lowest residual devience
cr_best2 <- cr_step.loop[cr_step.loop$nt==max(cr_step.loop$nt, na.rm = T),] # output with highets number of trees

cr_best <-cr_best %>% drop_na
cr_best
```

### Running catch rate BRT
Running gbm.step with optimal parameters and extracting n.trees. 
```{r cr.brt, warning=FALSE}

cr.brt<-dismo::gbm.step(data=dpdat,
                        gbm.x=c(crcov),
                        gbm.y=cr,## response - number of fish caught undamaged
                        learning.rate = 0.01, ## ref to step.loop
                        tree.complexity = 2, ## ref to step.loop
                        family="poisson", ## distribution family
                        bag.fraction=0.9,
                        step.size = 30)
```
### Outputs
#### Default outputs
```{r cr.step output}

ggPerformance(cr.brt)

cr.brt_influence <- ggInfluence(cr.brt, signif = TRUE, main = "Significant predictors for catch rate") 
ggsave(paste(crBRT_plots, "cr.brt_Influence.png", sep='/'), width = 8, height = 4)
cr.brt_influence

for(i in 1:nrow(cr.brt_influence)) {
cr.brt_PD <- ggPD(cr.brt, rug = T, smooth = T, predictor = i) 
ggsave(cr.brt_PD, file = paste(crBRT_plots, paste("cr.brt_PD", i, ".png"), sep = "/"), width = 8, height = 4)
}

for(i in 1:nrow(cr.brt_influence)) {
cr.brt_PDFit <- ggPDfit(cr.brt, smooth = T, se = T, predictor = i) 
ggsave(cr.brt_PDFit, file = paste(crBRT_plots, paste("cr.brt_PDFit", i, ".png"), sep = "/"), width = 8, height = 4)
}

ggInteract_list(cr.brt, index = F) 
```
#### Boostrapped ci
```{r cr.bootstrap, include=FALSE}

cr.brt_prerun<- plot.gbm.4list(cr.brt)

cr.brt_boot <- gbm.bootstrap.functions(cr.brt, list.predictors=cr.brt_prerun, n.reps=100)

for(i in 1:nrow(cr.brt_influence)) {
cr.brt_PDboot<-ggPD_boot(cr.brt, list.4.preds=cr.brt_prerun,
                booted.preds=cr.brt_boot$function.preds, type.ci="ribbon", rug=T, smooth = T, col.smooth = "red", predictor = i)
ggsave(cr.brt_PDboot, file = paste(crBRT_plots, paste("cr.brt_PDboot", i, ".png"), sep = "/"), width = 8, height = 4)
}



# The above code doesnt produce plots for factor variables? 
```