
```{r libs, include = FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(chron)
```

# Data Prep {.tabset} 
After downloading the data from arcCollector, you end up three individual .csv docs for each survey layer; meta_21.csv, use_21.csv, avidity_21.csv (this includes data from 2020 field trip to). Merge the three sets of data, linking them by their GlobalID and GUID. 

1. Read in documents
2. Merge meta.csv and use.csv to make MetaUse.csv 
3. Merge MetaUse.csv and avidity.csv to make Ning0421

```{r  readData and Merge}
meta <- read.csv("rumIgnore/Data-042021/meta_21.csv")
use <- read.csv("rumIgnore/Data-042021/use_21.csv")
avidity <- read.csv("rumIgnore/Data-042021/avidity_21.csv")
jon <- read.csv("rumIgnore/Data-jon/ALL3TRIPSMASTERDATASETFEB22.csv")

is.data.frame(meta) # checking object is df -> TRUE
is.data.frame(use) # checking object is df -> TRUE
is.data.frame(avidity) # checking object is df -> TRUE
is.data.frame(jon)

MetaUse<- merge(meta, use, by.x = "GlobalID", 
                by.y = "Guid", all.x = TRUE, all.y = TRUE) # merge Meta.csv (x) and Use.csv (y) dfs by GlobalID and GUID column
Ning0421 <- merge(MetaUse, avidity, by.x = "GlobalID.y", 
                   by.y = "Guid", all.x = TRUE, all.y = TRUE) # merge MetaUse0421.csv (x) and Avidity.csv (y) dfs by GlobalUse and GUIDavid column

write.csv(Ning0421, 'rumIgnore/Ning0421.csv') # save merged df as .csv, Path specified for rumIgnore folder.

tbl_df(jon)
glimpse(jon)
tbl_df(Ning0421) # converting to tibble 
glimpse(Ning0421) # kind of like str for tibble 


```
```{r Clean, cache = TRUE, echo = FALSE}
#### Bind ####
# changing name of date variable so it doesn't not append to Jon's Date variable which is in different format
Ning0421 <- Ning0421 %>% 
  rename(nDate = Date) 

Ning <- bind_rows(Ning0421, jon)  %>% # Appending Jon's data
  
#### Select ####
# The below variables have been deselected because they are unnecessary, ID vars will be replaced with a new ID col, Creator and Editor is always Harrison, only date variable I need is date which was inputted by interviewer and CreationDate.x incase it has been missed

  dplyr::select(-c(GlobalID.y, GlobalID.x, GlobalID.y.y, Guid,
                   OBJECTID.x, OBJECTID.y, OBJECTID,
                   Creator.x, Creator.y, Creator, EditDate.x, EditDate.y, EditDate, Editor.x,
                   Editor.y,Editor, CreationDate.y, CreationDate))  %>%
  
# Below are Canarvan Q's
  
  dplyr::select(-c(Boat_ramp_location, Interviewer_name, Day_type,Competition_noncompetition, 
            Number_boats_interviewed_that_day,What_time_did_lines_enter_the_water, 
            What_time_did_lines_leave_the_water, What_fishing_method_was_used, 
            What_type_of_bait_or_lure_was_used, Was_berleying_used, 
            What_was_the_maximum_depth_of_hooks, What_was_the_minimum_depth_of_hooks,
            Why_did_you_choose_to_fish_at_that_site, 
            How_many_fish_did_you_catch_including_those_kept_and_those_returned,
            What_species_of_fish_did_you_catch, Did_you_experience_shark_bite_offs,
            If_yes_how_many_fish_did_you_lose, 
            Were_these_fish_lost_completely_or_was_part_of_the_fish_.e.g.head._retrieved,
            Boat_name_or_number, Boat_length, Time_of_interview, 
            Full_completion_or_partial_or_refusal, If_partial_or_refusal_why,
            Comments, Have_you_been_interviewed_about_shark_bite_offs_before,
            How_many_times_have_you_fished_from_this_boat_ramp_before,
            How_many_days_have_you_fished_from_a_boat_this_year,
            How_many_years_have_you_been_fishing_for,
            Home_postcode, What_is_your_income_before_tax, How_many_sharks_were_hooked,
            If_you_suffered_depredation_what_was_the_cost_of_the_gear_that_you_lost_today))  %>%
  
# Below are Qs from Jon's data that will not be used in this analysis
  
  dplyr::select(-c("Day.number", "Julian.day", "Month.Year","Lat.bins", "X5nm.grid..block.code",
                   "Maidenhead..Grid.Square",
                   "Cost.of.gear.lost.to.depredation....", "Burley.used..Y.N", "Min.hook..depth")) %>%

#### Rename ####
# format: rename(newname = currentname)
  dplyr:: rename(Tripn = "Trip.number",
                 Interviewer = "REQUIRED..Interviewer",
                 BR = "REQUIRED..Boat.ramp.transect.name",
                 Screen18 = "SCREEN..18.", 
                 PrevInter = "SCREEN..have.we.interviewed.you.in.the.last.2.weeks.", 
                 Agreement = "REQUIRED..Agreement.to.Participate",
                 nDaysInArea = "How.many.days.are.you.in.the.area.",
                 nDaysInArea_other = "If.OTHER....days.in.area",
                 BoatAccess = "Do.you.have.access.to.a.private.boat.while.your.here.",
                 nBoatDays = "If.YES....days.you.intend.to.spend.on.boat.",
                 nBoatDays_other = "If.OTHER....days.on.private.boat.",
                 nShoreDays = "X..days.you.intend.to.be.engaged.in.shore.based.activities.",
                 nShoreDays_other = "If.OTHER....days.on.shore.",
                 nTimesLast12m = "X..separate.times.in.area.in.last.12.months.",
                 nTimesLast12m_other = "If.OTHER....times.in.last.12.months",
                 nTimes19 = "X..separate.times.in.area.in.2019.",
                 nTimes19_other = "If.OTHER....times.in.area.in.2019",
                 Covid = "Would.you.be.elsewhere.if.it.weren.t.for.travel.restrictions.",
                 SurveyLong = "x.x",
                 SurveyLat = "y.x",
                 EXTRACTIVE = "EXTRACTIVE.SURVEY",
                 exRecall = "If.RECALL..date",
                 FishingType = "What.type.of.fishing.did.you.do.at.this.site.",
                 FishingType_other = "If.OTHER..fishing.type",
                 BaitLure = "Did.you.use.a.bait.or.lure.",
                 exStart = "What.time.did.you..if.spearfishing...your.lines.get.in.the.water.",
                 exStop = "What.time.did.you..if.spearfishing..your.lines.get.out.the.water..",
                 MaxHook = "Max.depth.of.hook.at.this.site.",
                 KeptUndam = "X..fish.you.kept.undamaged.at.this.site.",
                 KeptUdam_other = "If.OTHER....kept.undamaged",
                 RelUndam = "X..released.undamaged.at.this.site.",
                 RelUndam_other = "If.OTHER....released.undamaged",
                 nDP = "X..depredated",
                 nDP_other = "If.OTHER....depredated",
                 Species = "What.species.did.you.catch.",
                 exWhyLeave = "Why.did.you.leave.this.site.",
                 exWhyLeave_other = "If.OTHER..why.did.you.leave.this.site.",
                 NONEXTRACTIVE = "NON_EXTRACTIVE.SURVEY",
                 nexRecall = "If.RECALL..date.1",
                 Activity = "What.activity.did.you.do.",
                 Activity_other = "If.OTHER..activity",
                 nexStart = "What.time.did.you.start.the.activity.",
                 nexStop = "What.time.did.you.end.the.activity.",
                 DiveMaxDepth = "If.DIVING..max.depth.",
                 WhyChoose = "Why.did.you.choose.this.site.",
                 WhyChoose_other = "If.OTHER..why.this.site",
                 nexWhyLeave = "Why.did.you.leave.this.site..1",
                 nexWhyLeave_other = "If.OTHER..why.did.you.leave.this.site..1",
                 UseLong = "x.y",
                 UseLat = "y.y",
                 exnTimes12m = "X..times.fished.from.boat.in.last.12.months.",
                 nexnTimes12m = "X..times.doing.NEX.last.12.months..",
                 exYrs = "X..years.fishing.",
                 nexYrs = "X..years.doing.NEX.",
                 DiveCert = "Are.you.a.diver..If.so..what.is.your.certification.",
                 nDives = "If.DIVING....dives.",
                 YrBorn = "What.year.were.you.born.",
                 Postcode = "Home.postcode",
                 Accom = "If.VISITOR..where.are.you.staying.",
                 Accom_other = "If.OTHER..where.are.you.staying",
                 Sex = "Male.or.Female",
                 Party = "Describe.party",
                 BoatLength = "Boat.length",
                 BoatType = "Boat.Type",
                 BoatID = "Boat.name.or.number",
                 jDate = "Date",
                 jBR = "Boat.ramp",
                 BRLat = "Boat.ramp.lat",
                 BRLong = "Boat.ramp.long",
                 nBoatsFishThatDay = "No..boats.fishing.that.day",
                 jUseLat = "Lat",
                 jUseLong = "Long",
                 KernDens = "Kernel.density",
                 SST = "SST..NOAA.data.",
                 DistBRm = "Distance.of..fishing.location..from.boat.ramp..miles.",
                 DistBRnm = "Distance.of..fishing.location..from.boat.ramp..nautical.miles.",
                 DistBRkm = "Distance.of..fishing.location..from.boat.ramp..km.",
                 jCaughtUndam = "No..fish.caught.undamaged",
                 jnDP = "No..fish.lost.to.sharks",
                 nHooked = "Total.fish.hooked",
                 perDP = "Percentage.depredation",
                 proDP = "Proportional.depredation...between.0.1.",
                 binDP = "Binary.depredation.presence.absence",
                 nUndamHr = "No..fish.caught..undamaged.per.hour",
                 nDPHr = "No..fish.lost.per.hour",
                 nHookedHr = "Total.fish.hooked.per.hour",
                 DPRateHr = "Depredation.rate.....per.hour",
                 jexStart = "Time.lines.in",
                 jexStop = "Time.lines.out",
                 MedianTime = "Median.time",
                 DecMedianTime = "Decimal.Median.time",
                 FishingHr = "Fishing.hours",
                 DecFishingHr = "Decimal.fishing.hours",
                 FishTime = "Fishing.time..rounded.to.nearest.whole.number",
                 jFishingType = "Fishing.method",
                 jBaitLure = "Bait.type",
                 jMaxHook = "Max.hook..depth",
                 jBoatID = "Boat.name.number",
                 jBoatLength = "Boat.Length",
                 jTime = "Time.of..interview",
                 nTimesFished = "How_many_times_have_you_fished_",
                 jexYrs = "How_many_years_have_you_been_fi",
                 jDaysBoatFished = "How_many_days_have_you_boat._fished_f",
                 jDaysBoatFishedCat = "How.many.days.have.you.boat.fished.in.the.last.year.categorical..1...0.25..2...26.50..3...51...75..4...76...100..5.",
                 jBRAvid = "How.many.times.fished.from.this.boat.ramp.categorical..1...0.25.times..2...26.50..3...51.75..4...76.100..5.....100.",
                 jPrevInter = "Have_you_been_interviewed_about") %>%
          
#### Filter ####
  dplyr::filter_all(all_vars(!grepl('ummy', .))) %>% # getting rid of Dummys/dummys
  dplyr::filter_all(all_vars(!grepl("lah", .))) %>% # getting rid of blah blahs
  dplyr::filter(is.na(BR) | BR != "BeadonCreek") %>% # Getting rid of Onslow data
# how to get rid of blank observations
# use string detect to make this one line of code
  
##### Date  and Year ####
##### ### Date and Time
# Date and jDate need to be a standard format, appended then formatted as.Date
# Add year column and filter our 2018
  dplyr::mutate(Date = ifelse(nDate %in% "", CreationDate.x, nDate)) %>% # Appending Creation date to blanks
  tidyr::separate(Date, into = c("Date", "Time", "ampm"), sep = " ") %>% # Separate datetime var for later
  dplyr::mutate(Date = parse_date_time(nDate, c("mdY IMS p")), # making POSIX
               Date = substr(Date, 1, 10), # separating out Date
               Date = as.Date(Date, "%Y-%m-%d"), # formatting as.Date to standardise format
               jDate = as.Date(jDate, "%d/%m/%Y"), # formatting as.Date to standardise format
               Year = substr(Date, 1, 4), # taking substring to make year col (1,4) chacrter in col date
               jYear = substr(jDate, 1, 4)) %>% 
  dplyr::filter(is.na(Year) | Year != 2018) %>% # keep NAs and years that dont equal 2018
  dplyr::mutate(Date = as.character(Date, "%d/%m/%Y"), # format as chacrter to be able to append 
                jDate = as.character(jDate, "%d/%m/%Y"), # without changing to character comes back numeric nonsense
               Date = ifelse(is.na(Date), jDate, Date), # Append date cols
               Date = as.Date(Date, "%d/%m/%Y")) %>% # format as date
  dplyr::mutate(Year = ifelse(is.na(Year), jYear, Year)) %>% # Appending year data
  dplyr::relocate(Date, Time, ampm, jTime, Year) %>% # Putting temporal data at front
  dplyr::select(-c(CreationDate.x, nDate, jYear, jDate)) %>% # de selecting redundant cols
  dplyr::mutate(facYear = as.factor(Year),
                Year = as.numeric(Year)) %>%

##### Time ####
##  Format as 24h
# Ning <- Ning %>% 
#  unite("nTime", Time, ampm, sep=" ") %>%
#  mutate(Time = hms::as.hms(parse_date_time(nTime, "IMS p")))
#  mutate(Time = format(strptime(nTime, "%I:%M:%S %p"), format="%H:%M"))
##  Add 8 h 
##  Append jTime
# mutate(Time = ifelse(Time %in% "NA", jTime, Time))

##### ID ####
# Add in ID col arranged by Date and Time
  dplyr::arrange(Date, Time) %>% 
  dplyr::mutate(ID = row_number()) %>%
  dplyr::relocate(ID) %>%
  
##### Interviewer ####
# If Year = 2015|2016, interviewer = Jon 
  dplyr::mutate(Interviewer = ifelse(Year < "2020", "Jon", Interviewer)) %>%

#### Site Data ####
# BR
  dplyr::mutate(BR = ifelse(is.na(BR), jBR, BR), # Appending jBR and BR
               BR = ifelse(str_detect(BR, "Coral"), "Coral Bay", # Standardising format
                           ifelse(str_detect(BR, "Exmouth"), "Exmouth", BR)),# Standardising format
               BR = ifelse(!is.na(Site), "", BR),
      # Site         
               Site = ifelse(str_detect(Site, "Turquoise"), "Turquoise", 
                             ifelse(str_detect(Site, "Bills"), "BillsBay", 
                                    ifelse(str_detect(Site, "Pilg"), "Pilgramunna", 
                                           ifelse(str_detect(Site, "Mildura"), "MilduraWreck", Site))))) %>% # remove last bracket if adding transect
# Transect         
#        Transect = ifelse(str_detect(Site,"Turquoise|Sandy|Oys|Osp|Lake|Kori|Tris|Mes|Yar|Tul|Neds|Pilg|SKurr"),
#                           "Tantabiddi",
#                           ifelse(str_detect(Site, "Yacht|Golf|Town|BundegiSouth"), "ExmouthN",
#                                 ifelse(str_detect(Site, "Pebble|Lear|ExmouthS|McL"), "ExmouthS",
#                                        ifelse(str_detect(Site, "Maud|Bill"), "CoralBay",
#                                               ifelse(str_detect(Site,
#                                                                 "LH|Wob|Dunes|Jan|BundegiBeach|Hunt|NTant|Mil|Bau|"), "Bundegi", BR))))) %>%
#  relocate(Transect, .before = BR)) <- works by itself just not after transect

#### PrevInter ####
# Appending to Jon make sure any duplicate boat names or numbers within the same trip = Yes
#### Agreement ####
  dplyr::mutate(Agreement = ifelse(Year < "2020", "Yes", Agreement)) %>%
# note there is a weird gap between 1357 - 1377
# Figure out what to do with all the blank observations

##### Resident vs local  #####
# Adding Binary Resident column
  dplyr::mutate(Resident = ifelse(str_detect(nDaysInArea, "Resident"), 1, 0)) %>%
  dplyr::relocate(Resident, .before = nDaysInArea) %>%

##### nDaysinArea ####
# Standardise nDaysinAreaOther  
# Turning times into days
  dplyr::mutate(nDaysInArea_other = ifelse(str_detect(nDaysInArea_other, "Three weeks|3W|three weeks"), "21",
                                    ifelse(str_detect(nDaysInArea_other, "Month|4 week"), "28",
                                           ifelse(str_detect(nDaysInArea_other, "4 months"), "112",
                                                  ifelse(str_detect(nDaysInArea_other, "6 months"), "168",
                                                         ifelse(str_detect(nDaysInArea_other, "8 weeks - semi resident"), "56",
                                                                ifelse(str_detect(nDaysInArea_other, "3 months|3mm"), "84",
                                                                       ifelse(str_detect(nDaysInArea_other, "5 days"), 5,
                                                                              ifelse(str_detect(nDaysInArea_other, "7 Exmouth|7 days in Exmouth "), 7,
                                                                                     ifelse(str_detect(nDaysInArea_other, "4 in ex"), 4,
                                                                                            ifelse(str_detect(nDaysInArea_other, "Maybe|Traveling|visiting"), NA, # NAing misc comments that arnt useful
                                                                                              ifelse(str_detect(nDaysInArea_other, "7 month"), "196", nDaysInArea_other)))))))))))) %>%
  # Dealing with rougue comments 
  mutate(Refusal = ifelse(str_detect(nDaysInArea_other, "Did|Language|Change|dog"), nDaysInArea_other, NA), 
         Party = ifelse(str_detect(nDaysInArea_other, "Family"), "2a2c", Party), 
         PrevInter = ifelse(str_detect(nDaysInArea_other, "Family"), "Yes", PrevInter),
         Site = ifelse(str_detect(nDaysInArea_other, "Mildura"), "MilduraWreck", Site),
         nDaysInArea_other = ifelse(str_detect(nDaysInArea_other, "Family|Mil|DUNES|Did|Language|Change|dog"), NA, nDaysInArea_other),
         nDaysInArea = ifelse(str_detect(nDaysInArea_other, "112|168|56|84|21|25|28|5|7|4|196|30|90|23"), nDaysInArea_other, nDaysInArea)) %>%
  dplyr::select(-c(nDaysInArea_other)) %>%
# Each of the ifelse statements were checked after each run, this was done by checking which row number in nDaysInArea_other associated with comment changed, and was checked to make sure it matched the corresponing row in new col eg. which(Ning$nDaysInArea_other == "2a2c") == which(Ning$Party == "2a2c")

#### BoatAccess #### 
  dplyr::mutate(BoatAccess = ifelse(Year < "2020", "Yes", BoatAccess)) %>%

#### nBoatDays ####
  dplyr::mutate(nBoatDays = ifelse(nBoatDays == "All", nDaysInArea, nBoatDays)) %>%

#### nBoatDays_other ####
# Explore
# Ning %>%
#  select(ID, Resident, nBoatDays_other, nBoatDays, nDaysInArea) %>%
#  mutate(nBoatDays_other = ifelse(nBoatDays_other == "", NA, nBoatDays_other)) %>%
#  filter(nBoatDays_other == ""|!is.na(nBoatDays_other)) %>% 
#  group_by(nBoatDays_other) %>%
#  arrange(Resident)

# Add ResBoatAvid - got to be applicable only when resident = 1 make NA
# Deal with visitor nBoatDays
 #  mutate(ResBoatAvidYr = ifelse(str_detect(nBoatDays_other, "1 formiye|Every couple of weeks|For|"), "26",
    #                             ifelse(str_detect(nBoatDays_other, ""), T, F)))
#### nTimesLast12m ####
  dplyr::mutate(nTimesLast12m_other = ifelse(str_detect(nTimesLast12m_other, "1|July"), "1",
                                      ifelse(str_detect(nTimesLast12m_other, "0|Zero|5-6|o"), NA, nTimesLast12m_other)),
         Comments = ifelse(str_detect(nTimesLast12m_other, "dog|house"), nTimesLast12m_other, NA),
         nTimesLast12m_other = ifelse(str_detect(nTimesLast12m_other, "dog|house"), NA, nTimesLast12m_other),
         nTimesLast12m = ifelse(nTimesLast12m_other == "1", "1", nTimesLast12m)) %>%
  dplyr::select(-nTimesLast12m_other) %>%
#### nTimes19 #### 
  dplyr::mutate(nTimes19_other = ifelse(str_detect(nTimes19_other, "1|Once"), "1",
                                 ifelse(str_detect(nTimes19_other, "Can’t|H|She’d|For|ould|Might"), NA,
                                        ifelse(str_detect(nTimes19_other, "Was living here"), "Resident", nTimes19_other))),
         Resident = ifelse(nTimes19_other == "Resident", "1", Resident),
         nTimes19_other = ifelse(str_detect(nTimes19_other, "Resident"), NA, nTimes19_other),
         nTimes19 = ifelse(nTimes19_other == "1", "1", nTimes19)) %>%
  dplyr::select(-nTimes19_other) %>%
  #### Append useLat/Long ####
 # mutate(UseLong = ifelse(Year == "2015", jUseLong, UseLong),
  #       UseLong = ifelse(Year == "2016", jUseLong, UseLong),
   #      UseLat = ifelse(Year == "2015", jUseLat, UseLat),
    #     UseLat = ifelse(Year == "2016", jUseLat, UseLat)) %>%
  # select(-c(jUseLong, jUseLat)) %>%
#### EXTRACTIVE ####
# Every entry which had a location name had coordinates checked to se if they matched or not, if not new coord added
  dplyr::mutate(UseLong = ifelse(str_detect(EXTRACTIVE, "ooper|squidding"), "114.2212", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Artificial reef|King"), "114.187083", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Marion"),"114.359287", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Exmouth reef"), "114.3667", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Five finger reef"), "113.751633", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "10k of bay"), "114.1126", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "South tanatabiddid"), "113.901809", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Ned’s"), "113.915051", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Fishing off Exmouth boat ramp|harbour"), "114.139613", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "trolled"), "113.919484", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "100m line"), "113.923947", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Helby bank|west side?"), "114.027586", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Long Island"), "114.674036", UseLong),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Peak island"), "114.509023", UseLat)) %>% 
    
  dplyr::mutate(UseLat = ifelse(str_detect(EXTRACTIVE, "ooper|squidding"), "-22.0555", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Artificial reef|King"), "-21.9155", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Marion"), "-21.6583", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Exmouth reef"), "-21.8500", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Five finger reef"), "-23.1820", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "10k of bay"), "-22.4413", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "South tanatabiddid"), "-21.9977", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Ned’s"), "-21.9956", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Fishing off Exmouth boat ramp|harbour"), "-21.9559", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "trolled"), "-21.9193", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "100m line"), "-21.8559", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Helby bank|west side?"), "-21.8100", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Long Island"), "-21.6144", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Peak island"), "-21.6039", UseLat)) %>%
  
  dplyr::mutate(EXTRACTIVE = ifelse(str_detect(EXTRACTIVE, "ooper|squidding|Artificial reef|King|Marion|Exmouth reef|Five finger reef|10k of bay|South tanatabiddid|Ned’s|Fishing off Exmouth boat ramp|harbour|trolled|100m line|Helby bank|west side?|Long Island|Peak island|Trolling|trawled|Don’t fish|interviewed|school|Janz|inlet|Bro|I|22k|Larked|tang"), NA, EXTRACTIVE)) %>%
  dplyr::mutate(YrBorn = ifelse(str_detect(EXTRACTIVE, "1982"), "1982", YrBorn),
         Postcode = ifelse(str_detect(EXTRACTIVE, "6166"), "6166", Postcode),
         Comments = ifelse(str_detect(EXTRACTIVE, "70%depredated"), "0%depredated - gone up, less fish", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Doesn’t fish deeper than 16m because of sharks"), "Doesn’t fish deeper than 16m because of sharks", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "6 spots"), "6 spots between tantabiddi br to light house outside reef", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Targeting marlin"), "Targeting marlin", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Oil"), "Oil rigs", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "They tag and release sharks"), "They tag and release sharks", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Couldn’t be blthard telling me about all the sites but they went to 5 sites and left every time because of sharks"), "Couldn’t be blthard telling me about all the sites but they went to 5 sites and left every time because of sharks", Comments),
         BaitLure = ifelse(str_detect(EXTRACTIVE, "bait"), "Both", BaitLure),
         BoatID = ifelse(str_detect(EXTRACTIVE, "BO 343"), "BO 343", BoatID)) %>%
  dplyr::select(-EXTRACTIVE) %>%
#### exRecall ####
# DPrum <- DPrum %>% 
#  separate(exRecall, into = c("exDate", "b", "c"), sep = " ") %>%# Separate datetime var for later
#  mutate(exDate = parse_date_time(exDate, c("mdY IMS p")), # making POSIX
#         exDate = substr(exDate, 1, 10), # separating out Date
#         exDate = as.Date(exDate, "%Y-%m-%d"), # formatting as.Date to standardise format
#         exDate = as.character(exDate, "%d/%m/%Y"), # format as chacrter to be able to append 
#         Date = as.character(Date, "%Y/%m/%d"), # without changing to character comes back numeric nonsense
#         Date = ifelse(str_detect(exDate, "201"), exDate, Date), # Append date cols
#         Date = as.Date(Date, "%d/%m/%Y")) # format as date%>% # Appending year data

#### Fishing Type ####
  dplyr::mutate(FishingType_other = ifelse(str_detect(FishingType_other, "we’re just squidding|Going for squid|Squidding"), "Squidding", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "ottom|Drift|Demersal"), "Demersal", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Free|Spearing"), "Spearing", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Casting|Lining"), "Casting", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Trawling|Trolling|For. Arlin|Marlin|game"), "Trolling", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Prawning"), "Prawning", FishingType_other),
         BaitLure = ifelse(str_detect(FishingType_other, "Both|Bith"), "Both", BaitLure),
         BaitLure = ifelse(str_detect(FishingType_other, "Bait"), "Bait", BaitLure),
         MaxHook = ifelse(str_detect(FishingType_other, "80"), "80", FishingType_other),
         UseLong = ifelse(str_detect(FishingType_other, "fly"), "114.555110", UseLong),
         UseLat = ifelse(str_detect(FishingType_other, "fly"), "-21.809301", UseLat),
         exnTimes12m = ifelse(str_detect(FishingType_other, "25 times a year"), "25", exnTimes12m),
         FishingType_other = ifelse(str_detect(FishingType_other, "Both|Bith|Bait|80|fly|25|0ff|From"), NA, FishingType_other),
         FishingType = ifelse(str_detect(FishingType_other, "Squidding"), "Squidding", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Demersal"), "Demersal", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Spearing"), "Spearing", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Casting"), "Casting", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Trolling"), "Trolling", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Prawning"), "Prawning", FishingType),
         FishingType_other = ifelse(str_detect(FishingType_other, "Squidding|Demersal|Spear|Casting|Trolling|Prawning"), NA, FishingType_other),
         FishingType = ifelse(Year < "2020", jFishingType, FishingType)) %>%
#### exStart ####
# Standardise and append to jexStart and nexStart -> activity start
#### exStop ####
# Standardise and append to jexStop and nexStop -> sctivity stop
### Activity ####
# This is being done quickly for the purposes of plot - double check no comments misinterpreted
# eg. Tant sz should be check for coordinates/scenery and waiting should be considered, are Sup and surfing clumpable? Making assumong that Kid/Family means beachgoing? Swimming but went for mega fauna?
  dplyr::mutate(Activity_other= ifelse(str_detect(Activity_other,"Swim|swim|play|Relax|castles|Chilling|Drinking|Kid|scenery|Waiting|waiting|Beach|Family|relax|whale"), "Beachgoing", Activity_other),
         Activity_other= ifelse(str_detect(Activity_other, "Dog|dog"), "DogWalking",Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Kayak|kayak"), "RowSport", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "walk|Walk"), "Walking", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Whale|turtle|Mega"), "Megafauna", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Sand|glass|Rockpooling|Quad|Bird|shells"), "Other", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Wing"), "WindSport", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Skurfing|Sup|Paddle|surf"), "Surfing", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Snork|snork|Tant|trout"), "Snorkeling", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Explore|explore|weather|look"), "Explore", Activity_other),
         Activity_other = ifelse(str_detect(Activity_other, "Just briefly |With group of fishers"), "", Activity_other),
         Activity = ifelse(str_detect(Activity, "Relaxing"), "Beachgoing", Activity),
         Activity = ifelse(str_detect(Activity, "Kite surfing "), "WindSport", Activity),
         Activity = ifelse(str_detect(Activity, "Snork"), "Snorkeling", Activity),
         Activity = ifelse(str_detect(Activity_other,"Beachgoing|DogWalking|RowSport|Walking|Mega|Other|Wind|Surfing|Snork|Explore"), Activity_other, Activity),
         Activity = ifelse(Year < "2020", "Fishing", Activity),
         Activity = ifelse(Activity == "", "Fishing", Activity),
         FishingType = ifelse(str_detect(FishingType, "Demersal|Bottom"), "Demersal", FishingType),
         UseLong = as.numeric(UseLong),
         UseLat = as.numeric(UseLat)) %>%
  dplyr::mutate(ActivityType = ifelse(str_detect(Activity, "Fishing"), "Extractive", "Non-Extractive")) %>%
#### MaxHook ####
# Standardise and append to jMaxHook
#### KeptUndam ####
# Standardise and append other
  dplyr::mutate(Species = ifelse(str_detect(KeptUdam_other, "uid"), "Squid", Species),
         Species = ifelse(str_detect(KeptUdam_other, "1 spangled emperor; 1 threadfin "), "Spangled1/thredfin1", Species),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "Caught|catch"), NA, KeptUdam_other),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "2|spangled"), "2", KeptUdam_other),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "4"), "4", KeptUdam_other),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "1 squid|Flatty"), "1", KeptUdam_other),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "litres"), "Ten litres between two people", KeptUdam_other),
         KeptUndam = ifelse(str_detect(KeptUdam_other, "2|5|4|6|1|3"), KeptUdam_other, KeptUndam),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "2|5|4|6|1|3"), NA, KeptUdam_other),
         KeptUndam = ifelse(Year > "2016" & is.na(KeptUndam) & Activity == "Fishing", "0", KeptUndam),
         KeptUndam = as.numeric(KeptUndam)) %>%
#### RelUndam ####
# Standardise and append other
 dplyr::mutate(RelUndam_other = ifelse(str_detect(RelUndam_other, "Rel|Too"), "1", RelUndam_other),
        RelUndam = ifelse(str_detect(RelUndam_other, "0|1|8|4|15"), RelUndam_other, RelUndam),
        RelUndam_other = ifelse(str_detect(RelUndam_other, "0|1|8|4|15"), NA, RelUndam_other),
        RelUndam = ifelse(Year > "2016" & is.na(RelUndam), "0", RelUndam),
        RelUndam = ifelse(Year > "2016" & is.na(RelUndam) & Activity == "Fishing", "0", RelUndam),
        RelUndam = as.numeric(RelUndam)) %>%
#### nDP ####
# standardise and append to other and jnDP
  dplyr::mutate(nDP_other = ifelse(str_detect(nDP_other, "mate"), "1", nDP_other),
         nDP = ifelse(str_detect(nDP_other, "1"), nDP_other, nDP),
         nDP = ifelse(Year < "2020", jnDP, nDP),
         nDP = ifelse(is.na(nDP)|is.nan(nDP), "0", nDP),
         nDP = as.numeric(nDP)) %>%
  dplyr::select(-jnDP) %>%
#### jCaughtUndam ####
# Append Rel Undam + Kept Undam to 2020/21 values
  dplyr::mutate(jCaughtUndam = ifelse(Year > "2016", RelUndam + KeptUndam, jCaughtUndam)) %>%
  dplyr::rename(CaughtUndam = jCaughtUndam) %>%
#### nHooked ####
# If Year = 2020/21, then = RelUndam + nDP + KeptUndam
 dplyr::mutate(nHooked = ifelse(Year > "2016", KeptUndam + RelUndam + nDP, nHooked)) %>%
#### perDP ####
# if Year = 2020/21, then = nDP/nHooked
 dplyr::mutate(perDP = ifelse(Year > "2016", (nDP/nHooked)*100, perDP)) %>%
  dplyr::relocate(CaughtUndam, nHooked, perDP, proDP, binDP, .before = Species) %>%

#### Append useLat/Long ####
  dplyr::mutate(UseLong = ifelse(Year < "2020", jUseLong, UseLong),
                UseLat = ifelse(Year < "2020", jUseLat, UseLat)) %>%
  dplyr::select(-c(jUseLong, jUseLat))

```

```{r csv}
write.csv(Ning, "rumIgnore/Ning_v2.csv")
```
