
```{r libraries, include = FALSE}
# library(tidyverse)
# library(lubridate)
# library(chron)
# library(sp)
# library(raster)
# library(ggplot2)
# library(leaflet)
# library(rgeos)
# library(rgdal)
# library(sf)
# library(tmap)
# library(RColorBrewer)
# library(ggspatial)
# library(ggsn)
# library(cowplot)
# library(rcartocolor)
```

```{r DataPrep}
### All data 
 N1 <- Ning %>%
  dplyr::filter(!is.na(UseLong), UseLong != "",
                !is.na(UseLat), UseLat != "")
               # UseLat != "-21.06476")

# Extractive data 
exdat <- N1 %>% 
  dplyr::filter(Activity == "Fishing") 

unique(exdat$Activity)

# Non-extractive data
nexdat <- N1 %>% 
  dplyr::filter(Activity != "Fishing")

unique(nexdat$Activity)
```

```{r spDataPrep}
## Setting up spatial layers
## crs1 <- 4283 is it useful to give a crs a name ?? GDA94
# Data 
sNing <- st_as_sf(N1, coords = c("UseLong", "UseLat"), ## Won't work if there are blanks in df so  filtered
                  crs = 4283)

# Coast
WA <- st_read(paste(s.dir, "waCoast.shp", sep = '/')) %>%
  st_transform(crs = 4283)

NWS <- st_crop(WA, xmin = 113.00, ymin = -24.00, xmax = 115.00, ymax = -20.5) # what you want to crop, masked onto what
# CB <- st_crop(WA, xmin = 113.6271, xmax = 114.000, ymin = -23.46973, ymax = -22.562)
# # Cape <- st_crop(WA, xmin = 113.6271, ymin = -22.562, xmax = 114.7503, ymax = -21.56369)
# 
# # cropping sNing to CB/cape extent
# CBning <- st_crop(sNing, CB)
# #Capening <- st_crop(sNing, Cape)
# Capening <- st_crop(sNing, Cape)

## BR
ausBR <- st_read(paste(s.dir, "ausRamps.shp", sep = '/')) %>%
  st_transform(crs = 4283)

BR <- st_crop(ausBR, sNing)
# CBR <- st_crop(BR, CB)
# CapeBR <- st_crop(BR, Cape)


 #### Bathy
bathy <- st_read(paste(s.dir, "waDepths.shp", sep = '/')) %>%
  st_transform(crs = 4283) %>%
  st_make_valid()

bathy

NWSbathy <- st_crop(bathy, NWS)
# CBbathy <- st_crop(NWSbathy, CB)
# Capebathy <- st_crop(NWSbathy, Cape)

# ggplot() +
#   geom_sf(data = NWSbathy, lwd = 0.1, fill = "white") +
#   geom_sf(data = NWS, lwd = 0.1) +
#   geom_sf(data = sNing)
# 
# # summary(Ning$UseLat)
# # max(Ning$UseLat, na.rm = T)
# # which(Ning$UseLat == "-21.06476")
# # Ning[567,]
# # max(N1$UseLat, na.rm = T)
# 
# NWS xmin: 113.4615 ymin: -24.0027 xmax: 115 ymax: -21.4219
# NWSbathy xmin: 113.4615 ymin: -24.0027 xmax: 115 ymax: -21.4219
# sNing xmin: 113.2797 ymin: -23.46818 xmax: 114.7503 ymax: -21.06476
# 
# 113, -24, 115, -21
# 
# NWS
```

```{r Insets}
# Making bbox for CB 
CBbbox = st_as_sfc(st_bbox(CB))

# Making CB inset
CBinset <- ggplot() +
  geom_sf(data = NWS, lwd = 0.1) +
  geom_sf(data = BR, color = "mediumaquamarine") +
  #geom_sf_text(data = BR, aes(label = boat_ramp), size = 2, nudge_x = 0.15) +
  geom_sf_label(data = BR, aes(label = boat_ramp), size = 1, nudge_x = 0.15, nudge_y = 0.02, label.padding =
                  unit(0.5, "mm"),label.size = 0) +
  labs(x = "", y = "") +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title=element_blank()) +
  geom_sf(data = CBbbox, fill = NA, color = "mediumaquamarine", size = 1)
  
CBinset

 ggsave(paste(rumPlots, "spBR.png", sep='/'), width = 8, height = 4)
 
```

# Overview of recreational marine use {.tabset}
## Activities
Histogram showing of data collected across all activities captured. 
```{r ActFreq, fig.dim = c(7, 7)}
ActFreq <- N1 %>% 
  group_by(Activity) %>%   
  mutate(count_name_occurr = n()) %>% 
  ggplot(aes(y=reorder(Activity,count_name_occurr))) +
  geom_bar(stat="count", fill = "mediumaquamarine") +
  labs(x = "Activity (n = 12)", y = "Frequency (n = 1217)") +
  geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5)
  
ActFreq
ggsave(paste(rumPlots, "ActFreq.png", sep='/'), width = 8, height = 4)
```

## Activity Type
Histogram showing data collected of extractive and non-extractive activities. 
```{r ActTyp}
 ActTyp <- ggplot(N1) +
   aes(x = ActivityType) +
   geom_bar(fill = "mediumaquamarine") +
   labs(x = "Activity Type", y = "Frequency") +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5)
 
 ActTyp
 ggsave(paste(rumPlots, "ActTyp.png", sep='/'), width = 8, height = 4)
```

## Spatial distribution of activities
Spatial distribution of marine reacrretional use. This included extarctive and non-extractive activities, boat and shore based. 
```{r spAct}
#### Spatial plot for all eactivities across whole study site
spAct <- ggplot() +
  geom_sf(data = NWSbathy, lwd = 0.1, fill = "white") +
   #geom_sf_text(data = NWSbathy, aes(label = LABEL), size = 2) +
  geom_sf(data = NWS, lwd = 0.1) +
  geom_sf(data = sNing, aes(color = Activity, shape = ActivityType), size = 1) + # colour for discrete, fill for continuous
  scale_color_brewer(palette = "Set3") + # colour for discrete, fill for continuous
  labs(caption = "Spatial distribution of marine recreational use") +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.5, "cm"), pad_y = unit(1, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm")) +
  theme(axis.text.x = element_text(angle = 90, size = 7),
        axis.text.y = element_text(size = 7),
        legend.key = element_rect(fill = "White", color = "White"),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 7, vjust = 1),
        legend.text = element_text(size = 7)) 

spAct
ggsave(paste(rumPlots, "spAct.png", sep='/'), width = 5, height = 6)
```

```{r CBActFreq}
##### Spatial plot for all activties at CB 
CBspAct <- ggplot() +
  geom_sf(data = CB, lwd = 0.1) +
  geom_sf(data = CBning, aes(color = Activity, shape = ActivityType), size = 1) + # colour for discrete, fill for continuous
  scale_color_brewer(palette = "Set3", direction = -1) + # colour for discrete, fill for continuous
  labs(caption = "Spatial distribution of marine recreational use", title = "Coral Bay") +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.5, "cm"), pad_y = unit(1, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm")) +
  theme(axis.text.x = element_text(angle = 90, size = 7),
        axis.text.y = element_text(size = 7),
        legend.key = element_rect(fill = "White", color = "White"),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 7, vjust = 1),
        legend.text = element_text(size = 7),
        title = element_text(size = 8))

CBspAct
ggsave(paste(rumPlots, "CBspAct.png", sep='/'), width = 5, height = 6)

# 
# CBspAct_inset = ggdraw() +
#   draw_plot(CBspAct) +
#   draw_plot(CBinset, x = 0.5, y = 0.05, width = 0.3, height = 0.2)
# 
# CBspAct_inset
# 
# ggsave(paste(rumPlots, "CBspAct_inset.png", sep='/'), width = 5, height = 6)
```

```{r CapeActFreq}
##### Spatial plot for all activties at CB 
CapespAct <- ggplot() +
  geom_sf(data = Cape, lwd = 0.1) +
  geom_sf(data = Capening, aes(color = Activity, shape = ActivityType), size = 1) + # colour for discrete, fill for continuous
  scale_color_brewer(palette = "Set3") + # colour for discrete, fill for continuous
  labs(caption = "Spatial distribution of marine recreational use") +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.5, "cm"), pad_y = unit(1, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm")) +
  theme(axis.text.x = element_text(angle = 90, size = 7),
        axis.text.y = element_text(size = 7),
        legend.key = element_rect(fill = "White", color = "White"),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 7, vjust = 1),
        legend.text = element_text(size = 7),
        title = element_text(size = 8))

CapespAct
ggsave(paste(rumPlots, "CapespAct.png", sep='/'), width = 5, height = 6)

```

# Extractive Activities {.tabset}
## Fishing Type
Histogram showing data collected across all fishing types facetted by year. 
```{r FTyp}
FTyp <- exdat %>% 
  group_by(FishingType) %>%   
  mutate(count_name_occurr = n()) %>%
  filter(!is.na(FishingType), FishingType != "") %>% 
  ggplot(
  aes(x=reorder(FishingType,-count_name_occurr))) + #reorder(col to reorder, by what var)
  geom_bar(stat='count', fill = "mediumaquamarine") +
  labs(x = "Fishing Type (n = 6)", y = "Frequency (n = 665)") +
  facet_wrap(. ~ Year) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5)

FTyp

ggsave(paste(rumPlots, "FTyp.png", sep='/'), width = 8, height = 4)
```

## Catch rate
Barplot showing the average catch rate per site facetted by year. 
```{r CR}
CR <- exdat1 %>%
  group_by(facYear) %>%
  summarise(avCR = mean(CaughtUndam)) 

CR

# Hist of nDP
avCR <- ggplot(CR) +
  aes(y = avCR, x = facYear) +
  geom_col(fill = "mediumaquamarine") +
  labs(x = "Average catch rate per site", y = "Year")

avCR
ggsave(paste(rumPlots, "avCR.png", sep='/'), width = 8, height = 4)
```
```{r}
nDP <- ggplot(exdat1) +
  aes(x = nDP) +
  geom_histogram(fill = "mediumaquamarine") +
  labs(x = "Number of fish depredated")

nDP
```

## Average depredation
Barplot showing the Average number of fish depredated per site facetted by year. 
```{r avDPYr}
# Average number of fish depredated per year
avDPYr <- exdat %>%
  filter(!is.na(nHooked),  nHooked != "", !is.na(nDP), nDP != "") %>%
  group_by(facYear) %>%
  summarise(ave = mean(nDP)) %>%
  #mutate(numYear = as.character(Year)) %>%
  ggplot( 
  aes(x = facYear, y = ave)) +
  geom_col(fill = "mediumaquamarine") +
  labs(x = "Year", y = "Average number of fish depredated per site")
  
avDPYr
ggsave(paste(rumPlots, "avDPYr.png", sep='/'), width = 8, height = 4)

```

```{r DPFTyp}
# dpFT <- exdat %>%
#   group_by(FishingType) %>%
#   summarise(tHooked = sum(nHooked),
#             tDP = sum(nDP),
#             tperDP = tDP/tHooked*100) 
# 
# dpFT
# 
# pdpFT <- ggplot(dpFT) +
#   aes(x = FishingType, y = tperDP) +
#    geom_col(fill = "mediumaquamarine") +
#   labs(x = "Fishing Type", y = "Percentage depredation")
# 
# pdpFT

```

## Percentage depredation
Barplot showing the percentage of total yearly catch lost to depredation
```{r perDP}
# Making data
exdp <- exdat %>%
  filter(!is.na(nHooked),  nHooked != "", !is.na(nDP), nDP != "") %>%
  group_by(Year) %>%
  summarise(tHooked = sum(nHooked),
            tDP = sum(nDP),
            tperDP = tDP/tHooked*100) %>%
  mutate(Year = as.character(Year)) 

exdp

# Bar plot of percentage depredation per year
 dpYr <- ggplot(exdp) +
  aes(x = Year, y = tperDP) +
  geom_col(colour = "mediumaquamarine", fill = "mediumaquamarine") +
  labs(y = "Percentage of catch depredated", x = "Year")
 
 dpYr
 ggsave(paste(rumPlots, "dpYr.png", sep='/'), width = 8, height = 4)
```

## Spatial distribution of shark depredation
Spatial distribution of recreational fishing coloured by percentage of depredation at each site. 
```{r spDP}
#### Spatial plot  of fishing ####
## st of just nDP == 0
sexdp0 <- exdat %>%
  filter(nDP == "0") %>%
  st_as_sf(coords = c("UseLong", "UseLat"), ## This wont worj if there are blanks in data set so  filtered
                  crs = 4283)
sexdp0
unique(sexdp0$nDP)

## st  nDP > 0
sexdp <- exdat %>%
  filter(nDP != "0") %>%
  st_as_sf(coords = c("UseLong", "UseLat"), ## This wont worj if there are blanks in data set so  filtered
                  crs = 4283)
sexdp
unique(sexdp$nDP)

## making spatial plot for whole study site
spexdp <- ggplot() +
  geom_sf(data = NWS, lwd = 0.1) +
  geom_sf(data = sexdp0, shape = 1, size = 0.5, color= "grey") +
  geom_sf(data = sexdp, aes(color = perDP), size = 1) + # colour for discrete, fill for continuous
  scale_color_distiller(palette = "YlGnBu", trans = "reverse") + # colour for discrete, fill for continuous
  scale_fill_continuous(trans = 'reverse') +
  labs(color = "Percentage\nDepredated", caption = "Points indicate sites of recreational fishing coloured by percentage depredation.\nSites denoted by hollow grey circle are sites where zero depredation occured.") +
  theme(axis.text.x = element_text(angle = 90, size = 5),
        axis.text.y = element_text(size = 5),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 5, vjust = 1),
        legend.text = element_text(size = 5),
        legend.key.height = unit(25, "mm")) +
  guides(color = guide_colorbar(reverse = TRUE)) +
  facet_wrap(. ~ Year) +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm"))

spexdp
ggsave(paste(rumPlots, "spexdp.png", sep='/'), width = 4, height = 6)
```

```{r spDP_CB}
# Make an extent for coral bay and run the above code
## st of just nDP == 0
CBsexdp0 <- st_crop(sexdp0, CB)

## st  nDP > 0
CBsexdp <- st_crop(sexdp, CB)

CBspexdp <- ggplot() +
  geom_sf(data = CB, lwd = 0.1) +
  geom_sf(data = CBsexdp0, shape = 1, size = 0.5, color= "grey") +
  geom_sf(data = CBsexdp, aes(color = perDP), size = 1) + # colour for discrete, fill for continuous
  scale_color_distiller(palette = "YlGnBu", trans = "reverse") + # colour for discrete, fill for continuous
  scale_fill_continuous(trans = 'reverse') +
  labs(color = "Percentage\nDepredated", caption = "Points indicate sites of recreational fishing coloured by percentage depredation.\nSites denoted by hollow grey circle are sites where zero depredation occured.", title = "Coral Bay") +
  theme(axis.text.x = element_text(angle = 90, size = 5),
        axis.text.y = element_text(size = 5),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 5, vjust = 1),
        legend.text = element_text(size = 5),
        legend.key.height = unit(25, "mm"),
        title = element_text(size = 8)) +
  guides(color = guide_colorbar(reverse = TRUE)) +
  facet_wrap(. ~ Year) +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm"))

CBspexdp

ggsave(paste(rumPlots, "CBspexdp.png", sep='/'), width = 4, height = 6)
```

```{r spDP_Cape}
# Cape
## st of just nDP == 0
Capesexdp0 <- st_crop(sexdp0, Cape)

## st  nDP > 0
Capesexdp <- st_crop(sexdp, Cape)

Capespexdp <- ggplot() +
  geom_sf(data = Cape, lwd = 0.1) +
  geom_sf(data = Capesexdp0, shape = 1, size = 0.5, color= "grey") +
  geom_sf(data = Capesexdp, aes(color = perDP), size = 1) + # colour for discrete, fill for continuous
  scale_color_distiller(palette = "YlGnBu", trans = "reverse") + # colour for discrete, fill for continuous
  scale_fill_continuous(trans = 'reverse') +
  labs(color = "Percentage\nDepredated", caption = "Points indicate sites of recreational fishing coloured by percentage depredation.\nSites denoted by hollow grey circle are sites where zero depredation occured.") +
  theme(axis.text.x = element_text(angle = 90, size = 5),
        axis.text.y = element_text(size = 5),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 5, vjust = 1),
        legend.text = element_text(size = 5),
        legend.key.height = unit(25, "mm"),
        title = element_text(size = 8)) +
  guides(color = guide_colorbar(reverse = TRUE)) +
  facet_wrap(. ~ Year) +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.5, "mm"), text_cex = 0.3, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.25, "cm"), pad_y = unit(0.5, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm"))

Capespexdp

ggsave(paste(rumPlots, "Capespexdp.png", sep='/'), width = 4, height = 6)
```

# Non-extractive
## Spatial distribution
Spatial plot showing the spatial distribution of non-extractive activities. 
```{r NonextractiveExplore}

snexdat <- st_as_sf(nexdat, coords = c("UseLong", "UseLat"), ## This wont worj if there are blanks in data set so  filtered
                  crs = 4283)

snexdat

spnex <- ggplot() +
  geom_sf(data = NWS, lwd = 0.1) +
  geom_sf(data = snexdat, aes(color = Activity), size = 1) + # colour for discrete, fill for continuous
  scale_color_brewer(palette = "Set3") + # colour for discrete, fill for continuous
  labs(caption = "Spatial distribution of non-extractive marine recreational use") +
  annotation_scale(location = "br", pad_y = unit(0.1, "cm"), height = unit(0.1, "cm"), text_cex = 0.5, bar_cols = c("grey", "white"), line_width = 0.1) +
    annotation_north_arrow(location = "br", which_north = "true", 
        pad_x = unit(0.5, "cm"), pad_y = unit(1, "cm"),
        style = north_arrow_fancy_orienteering(line_width = 1, text_size = 3, line_col = "Grey",
                                               fill = c("white", "Grey"), text_col = "grey"),
        height = unit(0.5, "cm"), width = unit(0.5, "cm")) +
  theme(axis.text.x = element_text(angle = 90, size = 7),
        axis.text.y = element_text(size = 7),
        legend.key = element_rect(fill = "White", color = "White"),
        plot.caption = element_text(hjust = 0, size = 5),
        legend.title = element_text(size = 7, vjust = 1),
        legend.text = element_text(size = 7))

spnex
ggsave(paste(rumPlots, "spnex.png", sep='/'), width = 4, height = 6)
```




```{r}
grid

## res.dev=3.05, lr=0.01, tc=1, nt=540
```
```{r}
  dp.stepW_dem<-dismo::gbm.step(data=exdat1,
                           gbm.x=c("numYear", "Resident", "nTimesLast12m", "UseLat", "UseLong", "DecFishingHr", "MaxHook",
                                   "exYrs", "jBRAvid", "BoatLength", "nHooked"), ## explanatory
                           gbm.y=c(40),## response
                           #offset=c("nHooked"), ## adds offset 
                           ##fold.vector=dat$folds, 
                           lr=0.01, ## ref to grid 
                           tc=3, ## ref to grid 
                           step.size=5,
                           family="poisson", ## distribution family
                           bag.fraction=0.75)
  grid[i, "res.dev"] <-dp.stepW_dem$self.statistics$mean.resid ### store res.dev in grid 
#sink()
#
#790

```

```{r Creating layers}
## Everything in this chunk only needs to be done once
## crs =  4283 (GDA94), units = m

# Adding coast
# WA <- st_read(paste(s.dir, "waCoast.shp", sep = '/')) %>%
#   st_transform(crs = 4283)

# Making bbox
# StudySite = st_as_sfc(st_bbox(c(xmin = 112.5, ymin = -24.03353, xmax = 115.00, ymax = -20.5), crs = 4283))
# st_write(StudySite, paste(s.dir, "StudySite.shp", sep = '/'))

# Cropping coast to study site
# NWS <- st_crop(WA, StudySite)
# NWS <- st_union(NWS)
# st_write(NWS, paste(s.dir, "NWScoast.shp", sep = '/'))

NWS <- st_read(paste(s.dir, "NWScoast.shp", sep = '/')) %>%
  st_transform(crs = 4283)

StudySite <- st_read(paste(s.dir, "StudySite.shp", sep = '/')) %>%
  st_transform(crs = 4283)

GridExtent <- st_read(paste(s.dir, "GridExtent.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

# Adding habitat layers
# Pelagic <- st_read(paste(s.dir, "Pelagic.gpkg", sep = '/')) %>%
#   st_transform(crs = 4283) 
# 
# Backreef <- st_read(paste(s.dir, "Backreef.shp", sep = '/')) %>%
#   st_transform(crs = 4283)
# 
# Reef <- st_read(paste(s.dir, "Reef.shp", sep = '/')) %>%
#   st_transform(crs = 4283)
# 
# Lagoon <- st_read(paste(s.dir, "Lagoon.shp", sep = '/')) %>%
#   st_transform(crs = 4283)

# Making grids - use this when you need to manipulate size of grids
# PelagicGrid <- Pelagic %>%
#   st_make_grid(cellsize = 0.05, square = FALSE, crs=4283) %>%
#   st_intersection(Pelagic)
# st_write(PelagicGrid, paste(s.dir, "PelagicGrid.shp", sep = '/'))
# 
# BackreefGrid <- Backreef %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(Backreef)
# st_write(BackreefGrid, paste(s.dir, "BackreefGrid.shp", sep = '/'))
# 
# ReefGrid <- Reef %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(Reef)
# st_write(ReefGrid, paste(s.dir, "ReefGrid.shp", sep = '/'))
# 
# LagoonGrid <- Lagoon %>%
#   st_make_grid(cellsize = 0.03, square = FALSE, crs=4283) %>%
#   st_intersection(Lagoon)
# st_write(LagoonGrid, paste(s.dir, "LagoonGrid.shp", sep = '/'))

## Reading grids
PelagicGrid <- st_read(paste(s.dir, "PelagicGrid.shp", sep = '/')) %>%
  st_transform(crs = 4283)

BackreefGrid <- st_read(paste(s.dir, "BackreefGrid.shp", sep = '/')) %>%
  st_transform(crs = 4283)

ReefGrid <- st_read(paste(s.dir, "ReefGrid.shp", sep = '/')) %>%
  st_transform(crs = 4283)

LagoonGrid <- st_read(paste(s.dir, "LagoonGrid.shp", sep = '/')) %>%
  st_transform(crs = 4283)

## Adding habitat variable to Grids
PelagicGrid$Habitat <- c("Pelagic")
BackreefGrid$Habitat <- c("Backreef")
ReefGrid$Habitat <- c("Reef")
LagoonGrid$Habitat <- c("Lagoon")

## Joining Grids
Grid <- rbind(PelagicGrid, BackreefGrid)
Grid <- rbind(Grid, ReefGrid)
Grid <- rbind(Grid, LagoonGrid)

## Adding depth 
Bathy <- raster(paste(r.dir, "NWSbathy.tif", sep = '/'))
crs(Bathy)  # 4362

Grid_wgs84<-st_transform(Grid, 4326)

# Extracting depth   
Grid_wgs84<-extract_mean_from_raster(Bathy, Grid_wgs84)
Grid_wgs84$Depth<-Grid_wgs84$mean

# convert back to 4283
Grid <- st_transform(Grid_wgs84, 4283)

## Add Grid ID
Grid <- Grid %>%
  mutate(GridID = row_number())

ggplot() +
  theme_void() +
  geom_sf(data = PelagicGrid, lwd = 0.1, fill = "light blue") +
  geom_sf(data = BackreefGrid, lwd = 0.1, fill = "pink") +
  #geom_sf(data = Grid, lwd = 0.1, aes(fill = Habitat))
  geom_sf(data = ReefGrid, lwd = 0.1, fill = "coral") +
  geom_sf(data = LagoonGrid, lwd = 0.1, fill = "aquamarine") +
  geom_sf(data = spNing, size = 0.05) 
```

```{r Allocating sites}
spNing <- Ning %>%
  dplyr::filter(!is.na(UseLong), UseLong != "",
                !is.na(UseLat), UseLat != "")

spNing <- spNing %>%
  filter(!is.na(UseLat), !is.na(UseLong), UseLat < -21.25) # defo a mistake in the survey

spNing <- st_as_sf(spNing, coords = c("UseLong", "UseLat"), crs = 4283)

spNing %>% 
  st_join(Grid[ , c("GridID")], left = T, join = st_intersects) %>%
  st_join(Grid[ , c("Depth")], left = T, join = st_intersects) %>%
  st_join(Grid[ , c("Habitat")], left = T, join = st_intersects) 
```

Proximity to artificial infrastructure 
```{r}
# Add artificial infrastructure
Pipes <- st_read(paste(s.dir, "Pipes.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

Wells <- st_read(paste(s.dir, "Wells.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

ArtReef <- st_read(paste(s.dir, "ArtReef.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

## note that some distance  that there is no argument to prevent distance from intersecting with land - however as we are just taking the proximity to nearest it doesn't matter

n <- st_nn(spNing, Wells, returnDist = TRUE, sparse = FALSE) # Calculate distance to nearest well
DistNearestWellm = sapply(n[[2]], "[", 1) # Name distance
DistNearestWellkm = DistNearestWell/1000 # Convert to km
spNing$DistNearestWellkm = DistNearestWellkm # Bind distances

n <- st_nn(spNing, Pipes, returnDist = TRUE, sparse = FALSE) # Calculate distance to nearest well
DistNearestPipem = sapply(n[[2]], "[", 1) # Name distance
DistNearestPipekm = DistNearestPipem/1000 # Convert to km
spNing$DistNearestPipekm = DistNearestPipekm # Bind distances

DistArtReefm = st_distance(spNing, ArtReef)
DistArtReefkm = DistArtReefkm/1000
spNing$DistArtReefkm = DistArtReefkm

spNing <- spNing %>%
  rowwise() %>%
  mutate(DistNearestInfra = min(DistNearestWellkm, DistNearestPipekm, DistArtReefkm)) %>%
  relocate(DistNearestWellkm, DistNearestPipekm, DistArtReefkm, DistNearestInfra, .before =
             geometry) %>%
  ungroup()
```
```{r spNing and shp}
st_write(spNing, paste(s.dir, "spNing.shp", sep = '/'))
```
```{r Distance to BR}
BR <- st_read(paste(s.dir, "NingBR.gpkg", sep = '/')) %>%
  st_transform(crs = 4283)

n <- st_nn(spNing, BR, returnDist = TRUE, sparse = FALSE) # Calculate distance to nearest well
DistNearestBRm = sapply(n[[2]], "[", 1) # Name distance
DistNearestBRkm = DistNearestBRm/1000 # Convert to km
spNing$DistNearestBRkm = DistNearestBRkm # Bind distances
```


```{r Kernal denisty}
# Ning1 <- Ning %>%
#   dplyr::filter(!is.na(UseLong),!is.na(UseLat))
# n <- kde2d(Ning1$UseLong, Ning1$UseLat)
# KernalDensity = sapply(n[[3]], "[", 1)
# 
# Ning1$KernalDensity = KernalDensity
# 
# ggplot() +
#   theme_void() +
#   geom_sf(data = StudySite, lwd = 0.1, fill = "white") +
#   geom_sf(data = NWS, lwd = 0.1) +
#   stat_density2d(aes(fill = ..level..), h = .02, n = 300,
#                  geom = "polygon", data = Ning1) +
#   scale_fill_viridis_c()
#   
# rm(Ning1)
# 
# ## Using denisty from spatstat
# spNing_Lam <- st_read(paste(s.dir, "spNing.shp", sep = '/')) %>%
#   st_transform(crs = 3112)
# 
# spNing_Lam <- as.ppp(spNing_Lam)
# spNingkm_Lam <- rescale(spNing_Lam, 1000, "km")
# 
# spNingkm_Lam$KernalDensity <- density(spNingkm_Lam, sigma = 5) # Using the default bandwidth
# kdens <- plot(spNingkm_Lam$KernalDensity, main=NULL, las=1)
# contour(spNingkm_Lam$KernalDensity, add=TRUE)
```

```{r}
ggplot() +
  theme_void() +
  geom_sf(data = StudySite, lwd = 0.1, fill = "white") +
  geom_sf(data = NWS, lwd = 0.1) +
  geom_sf(data = Pelagic, lwd = 0.1, colour = "purple") +
  geom_sf(data = Backreef, lwd = 0.1, aes(colour = "light blue")) +
  geom_sf(data = Reef, lwd = 0.1, aes(colour = "coral")) +
  geom_sf(data = Lagoon, lwd = 0.1, aes(colour = "aquamarine")) +
  geom_sf(data = Pipes, lwd = 0.1) +
  geom_sf(data = ArtInf, lwd = 0.1) 
```

