
```{r libraries, include = FALSE}
library(tidyverse)
library(lubridate)
library(chron)

library(sp)
library(rgdal)
library(raster)
library(ggplot2)
library(leaflet)
library(sf)
library(tmap)
```

# Data Prep {.tabset} 

## Collating Data 
After downloading the data from arcCollector, you end up three individual .csv docs for each survey layer; meta_21.csv, use_21.csv, avidity_21.csv (this includes data from 2020 field trip to). Merge the three sets of data, linking them by their GlobalID and GUID. 

1. Read in documents
2. Merge meta.csv and use.csv to make MetaUse.csv 
3. Merge MetaUse.csv and avidity.csv to make Ning0421

```{r  readData and Merge}
meta <- read.csv("rumIgnore/Data-042021/meta_21.csv")
use <- read.csv("rumIgnore/Data-042021/use_21.csv")
avidity <- read.csv("rumIgnore/Data-042021/avidity_21.csv")
jon <- read.csv("rumIgnore/Data-jon/ALL3TRIPSMASTERDATASETFEB22.csv")

is.data.frame(meta) # checking object is df -> TRUE
is.data.frame(use) # checking object is df -> TRUE
is.data.frame(avidity) # checking object is df -> TRUE
is.data.frame(jon)

MetaUse<- merge(meta, use, by.x = "GlobalID", 
                by.y = "Guid", all.x = TRUE, all.y = TRUE) # merge Meta.csv (x) and Use.csv (y) dfs by GlobalID and GUID column
Ning0421 <- merge(MetaUse, avidity, by.x = "GlobalID.y", 
                   by.y = "Guid", all.x = TRUE, all.y = TRUE) # merge MetaUse0421.csv (x) and Avidity.csv (y) dfs by GlobalUse and GUIDavid column

write.csv(Ning0421, 'rumIgnore/Ning0421.csv') # save merged df as .csv, Path specified for rumIgnore folder.

tbl_df(jon)
glimpse(jon)
tbl_df(Ning0421) # converting to tibble 
glimpse(Ning0421) # kind of like str for tibble 
```

## Cleaning  
To start I am going to select relevant variables and rename them into a R friendly format. 

```{r General}
#### Bind ####
# changing name of date variable so it doesn't not append to Jon's Date variable which is in different format

Ning0421 <- Ning0421 %>% 
  rename(nDate = Date) 

Ning <- bind_rows(Ning0421, jon)  %>% # Adding on Jons data
  
#### Select ####
# The below variables have been deselected because they are unnecessary, ID vars will be replaced with a new ID col, Creator and Editor is always Harrison, only date variable I need is date which was inputted by interviewer and CreationDate.x incase it has been missed

  select(-c(GlobalID.y, GlobalID.x, GlobalID.y.y, Guid, OBJECTID.x, OBJECTID.y, OBJECTID,
            Creator.x, Creator.y, Creator, EditDate.x, EditDate.y, EditDate, Editor.x, Editor.y,
            Editor, CreationDate.y, CreationDate))  %>%
  
# Below are Canarvan Q's
  
  select(-c(Boat_ramp_location, Interviewer_name, Day_type,Competition_noncompetition, 
            Number_boats_interviewed_that_day,What_time_did_lines_enter_the_water, 
            What_time_did_lines_leave_the_water, What_fishing_method_was_used, 
            What_type_of_bait_or_lure_was_used, Was_berleying_used, 
            What_was_the_maximum_depth_of_hooks, What_was_the_minimum_depth_of_hooks,
            Why_did_you_choose_to_fish_at_that_site, 
            How_many_fish_did_you_catch_including_those_kept_and_those_returned,
            What_species_of_fish_did_you_catch, Did_you_experience_shark_bite_offs,
            If_yes_how_many_fish_did_you_lose, 
            Were_these_fish_lost_completely_or_was_part_of_the_fish_.e.g.head._retrieved,
            Boat_name_or_number, Boat_length, Time_of_interview, 
            Full_completion_or_partial_or_refusal, If_partial_or_refusal_why,
            Comments, Have_you_been_interviewed_about_shark_bite_offs_before,
            How_many_times_have_you_fished_from_this_boat_ramp_before,
            How_many_days_have_you_fished_from_a_boat_this_year,
            How_many_years_have_you_been_fishing_for,
            Home_postcode, What_is_your_income_before_tax, How_many_sharks_were_hooked,
            If_you_suffered_depredation_what_was_the_cost_of_the_gear_that_you_lost_today))  %>%
  
  # Below are Qs from Jon's data that will not be used in this analysis
  
  select(-c("Day.number", "Julian.day", "Trip.number", "Month.Year","Lat.bins", "X5nm.grid..block.code", "Maidenhead..Grid.Square",
            "Cost.of.gear.lost.to.depredation....", "Burley.used..Y.N", "Min.hook..depth", "Have_you_been_interviewed_about",
            "How.many.times.fished.from.this.boat.ramp.categorical..1...0.25.times..2...26.50..3...51.75..4...76.100..5.....100.",
            )) %>%

#### Rename ####
# format: rename(newname = currentname)
  rename(Interviewer = "REQUIRED..Interviewer",
         BR = "REQUIRED..Boat.ramp.transect.name",
         Screen18 = "SCREEN..18.", 
         PrevInter = "SCREEN..have.we.interviewed.you.in.the.last.2.weeks.", 
         Agreement = "REQUIRED..Agreement.to.Participate",
         nDaysInArea = "How.many.days.are.you.in.the.area.",
         nDaysInArea_other = "If.OTHER....days.in.area",
         BoatAccess = "Do.you.have.access.to.a.private.boat.while.your.here.",
         nBoatDays = "If.YES....days.you.intend.to.spend.on.boat.",
         nBoatDays_other = "If.OTHER....days.on.private.boat.",
         nShoreDays = "X..days.you.intend.to.be.engaged.in.shore.based.activities.",
         nShoreDays_other = "If.OTHER....days.on.shore.",
         nTimesLast12m = "X..separate.times.in.area.in.last.12.months.",
         nTimesLast12m_other = "If.OTHER....times.in.last.12.months",
         nTimes19 = "X..separate.times.in.area.in.2019.",
         nTimes19_other = "If.OTHER....times.in.area.in.2019",
         Covid = "Would.you.be.elsewhere.if.it.weren.t.for.travel.restrictions.",
         SurveyLong = "x.x",
         SurveyLat = "y.x",
         EXTRACTIVE = "EXTRACTIVE.SURVEY",
         exRecall = "If.RECALL..date",
         FishingType = "What.type.of.fishing.did.you.do.at.this.site.",
         FishingType_other = "If.OTHER..fishing.type",
         BaitLure = "Did.you.use.a.bait.or.lure.",
         exStart = "What.time.did.you..if.spearfishing...your.lines.get.in.the.water.",
         exStop = "What.time.did.you..if.spearfishing..your.lines.get.out.the.water..",
         MaxHook = "Max.depth.of.hook.at.this.site.",
         KeptUndam = "X..fish.you.kept.undamaged.at.this.site.",
         KeptUdam_other = "If.OTHER....kept.undamaged",
         RelUndam = "X..released.undamaged.at.this.site.",
         RelUndam_other = "If.OTHER....released.undamaged",
         nDP = "X..depredated",
         nDP_other = "If.OTHER....depredated",
         Species = "What.species.did.you.catch.",
         exWhyLeave = "Why.did.you.leave.this.site.",
         exWhyLeave_other = "If.OTHER..why.did.you.leave.this.site.",
         NONEXTRACTIVE = "NON_EXTRACTIVE.SURVEY",
         nexRecall = "If.RECALL..date.1",
         Activity = "What.activity.did.you.do.",
         Activity_other = "If.OTHER..activity",
         nexStart = "What.time.did.you.start.the.activity.",
         nexStop = "What.time.did.you.end.the.activity.",
         DiveMaxDepth = "If.DIVING..max.depth.",
         WhyChoose = "Why.did.you.choose.this.site.",
         WhyChoose_other = "If.OTHER..why.this.site",
         nexWhyLeave = "Why.did.you.leave.this.site..1",
         nexWhyLeave_other = "If.OTHER..why.did.you.leave.this.site..1",
         UseLong = "x.y",
         UseLat = "y.y",
         exnTimes12m = "X..times.fished.from.boat.in.last.12.months.",
         nexnTimes12m = "X..times.doing.NEX.last.12.months..",
         exYrs = "X..years.fishing.",
         nexYrs = "X..years.doing.NEX.",
         DiveCert = "Are.you.a.diver..If.so..what.is.your.certification.",
         nDives = "If.DIVING....dives.",
         YrBorn = "What.year.were.you.born.",
         Postcode = "Home.postcode",
         Accom = "If.VISITOR..where.are.you.staying.",
         Accom_other = "If.OTHER..where.are.you.staying",
         Sex = "Male.or.Female",
         Party = "Describe.party",
         BoatLength = "Boat.length",
         BoatType = "Boat.Type",
         BoatID = "Boat.name.or.number",
         jDate = "Date",
         jBR = "Boat.ramp",
         BRLat = "Boat.ramp.lat",
         BRLong = "Boat.ramp.long",
         nBoatsFishThatDay = "No..boats.fishing.that.day",
         jUseLat = "Lat",
         jUseLong = "Long",
         KernDens = "Kernel.density",
         SST = "SST..NOAA.data.",
         DistBRm = "Distance.of..fishing.location..from.boat.ramp..miles.",
         DistBRnm = "Distance.of..fishing.location..from.boat.ramp..nautical.miles.",
         DistBRkm = "Distance.of..fishing.location..from.boat.ramp..km.",
         jCaughtUndam = "No..fish.caught.undamaged",
         jnDP = "No..fish.lost.to.sharks",
         nHooked = "Total.fish.hooked",
         perDP = "Percentage.depredation",
         proDP = "Proportional.depredation...between.0.1.",
         binDP = "Binary.depredation.presence.absence",
         nUndamHr = "No..fish.caught..undamaged.per.hour",
         nDPHr = "No..fish.lost.per.hour",
         nHookedHr = "Total.fish.hooked.per.hour",
         DPRateHr = "Depredation.rate.....per.hour",
         jexStart = "Time.lines.in",
         jexStop = "Time.lines.out",
         MedianTime = "Median.time",
         DecMedianTime = "Decimal.Median.time",
         FishingHr = "Fishing.hours",
         DecFishingHr = "Decimal.fishing.hours",
         FishTime = "Fishing.time..rounded.to.nearest.whole.number",
         jFishingType = "Fishing.method",
         jBaitLure = "Bait.type",
         jMaxHook = "Max.hook..depth",
         jBoatID = "Boat.name.number",
         jBoatLength = "Boat.Length",
         jTime = "Time.of..interview",
         nTimesFished = "How_many_times_have_you_fished_",
         jexYrs = "How_many_years_have_you_been_fi",
         jDaysBoatFished = "How_many_days_have_you_boat._fished_f",
         jDaysBoatFishedCat = "How.many.days.have.you.boat.fished.in.the.last.year.categorical..1...0.25..2...26.50..3...51...75..4...76...100..5.") %>% 
  
 #### Filter ####
  filter_all(all_vars(!grepl('ummy', .))) %>% # getting rid of Dummys/dummys
  filter_all(all_vars(!grepl("lah", .))) %>% # getting rid of blah blahs
  filter(is.na(BR) | BR != "BeadonCreek")  # Getting rid of Onslow data
# how to get rid of blank observations

#### Checkpoint ####
# Checking df is as expected
tbl_df(Ning)

```
Data Frame is now is less unweildy and contains only data that I need for my analysis. I am now going to clean each column starting with the leftmost. 
 
### Date and Time
* Date and jDate need to be a standard format, appended then formatted as.Date
* Add year column and filter our 2018
* Time needs to be changed from UTC to AWST and put into 24h format and appended to jTime
```{r DateClean}
##### Date  and Year ####

Ning <- Ning %>%
  mutate(Date = ifelse(nDate %in% "", CreationDate.x, nDate)) %>% # Appending Creation date to blanks
  separate(Date, into = c("Date", "Time", "ampm"), sep = " ") %>% # Separate datetime var for later
  mutate(Date = parse_date_time(nDate, c("mdY IMS p")), # making POSIX
         Date = substr(Date, 1, 10), # separating out Date
         Date = as.Date(Date, "%Y-%m-%d"), # formatting as.Date to standardise format
         jDate = as.Date(jDate, "%d/%m/%Y"), # formatting as.Date to standardise format
         Year = substr(Date, 1, 4), # taking substring to make year col (1,4) chacrter in col date
         jYear = substr(jDate, 1, 4)) %>% 
  filter(is.na(Year) | Year != 2018) %>% # keep NAs and years that dont equal 2018
  mutate(Date = as.character(Date, "%d/%m/%Y"), # format as chacrter to be able to append 
         jDate = as.character(jDate, "%d/%m/%Y"), # without changing to character comes back numeric nonsense
         Date = ifelse(is.na(Date), jDate, Date), # Append date cols
         Date = as.Date(Date, "%d/%m/%Y")) %>% # format as date
  mutate(Year = ifelse(is.na(Year), jYear, Year)) %>% # Appending year data
  relocate(Date, Time, ampm, jTime, Year) %>% # Putting temporal data at front
  select(-c(CreationDate.x, nDate, jYear, jDate)) # de selecting redundant cols

##### Time ####
##  Format as 24h
# Ning <- Ning %>% 
#  unite("nTime", Time, ampm, sep=" ") %>%
#  mutate(Time = hms::as.hms(parse_date_time(nTime, "IMS p")))
#  mutate(Time = format(strptime(nTime, "%I:%M:%S %p"), format="%H:%M"))
##  Add 8 h 
##  Append jTime
# mutate(Time = ifelse(Time %in% "NA", jTime, Time))
```

### Survey Meta data 
* Add ID column in chonological order - Done
* Add Jon as interviewer - Done
* Append jBR to BR and standardise format - Done
* Standardise Site names - Done
* Add Transect column which = corresponding shore transect/BR
* If boat nume/number is duplicated within one trip = Yes
* 
```{r SurveyMetaClean}
##### ID ####
# Add in ID col arranged by Date and Time

Ning <- Ning %>%
  arrange(Date, Time) %>% 
  mutate(ID = row_number()) %>%
  relocate(ID) %>%
  
##### Interviewer ####
# If Year = 2015|2016, interviewer = Jon 
# 
  mutate(Interviewer = ifelse(Year == "2015", "Jon", Interviewer),
         Interviewer = ifelse(Year == "2016", "Jon", Interviewer))

#### Site Data ####
#### 
Ning <- Ning %>% 
  mutate(BR = ifelse(is.na(BR), jBR, BR), # Appending jBR and BR
         BR = ifelse(str_detect(BR, "Coral"), "Coral Bay", # Standardising format
                     ifelse(str_detect(BR, "Exmouth"), "Exmouth", BR)),# Standardising format
         BR = ifelse(!is.na(Site), "", BR),
         
         Site = ifelse(str_detect(Site, "Turquoise"), "Turquoise", 
                       ifelse(str_detect(Site, "Bills"), "BillsBay", 
                              ifelse(str_detect(Site, "Pilg"), "Pilgramunna", 
                                     ifelse(str_detect(Site, "Mildura"), "MilduraWreck", Site))))) # remove last bracket if adding transect
         
#        Transect = ifelse(str_detect(Site,"Turquoise|Sandy|Oys|Osp|Lake|Kori|Tris|Mes|Yar|Tul|Neds|Pilg|SKurr"),
#                           "Tantabiddi",
#                           ifelse(str_detect(Site, "Yacht|Golf|Town|BundegiSouth"), "ExmouthN",
#                                 ifelse(str_detect(Site, "Pebble|Lear|ExmouthS|McL"), "ExmouthS",
#                                        ifelse(str_detect(Site, "Maud|Bill"), "CoralBay",
#                                               ifelse(str_detect(Site,
#                                                                 "LH|Wob|Dunes|Jan|BundegiBeach|Hunt|NTant|Mil|Bau|"), "Bundegi", BR))))) %>%
#  relocate(Transect, .before = BR)) <- works by itself just not after transect
  
names(Ning)
unique(Ning$BR)
unique(Ning$Site)
unique(Ning$Transect)
class(Ning$Transect)

#### PrevInter ####
#### Agreement ####
Ning <- Ning %>%
  mutate(Agreement = ifelse(Year == "2015", "Yes",
                            ifelse( Year == "2016", "Yes", Agreement)))
unique(Ning$Agreement)
which(is.na(Ning$Agreement))
which(Ning$Agreement == "")
# note there is a weird gap between 1357 - 1377
# Figure out what to do with all the blank observations
```

### Meta Data
```{r MetaClean}
##### Resident vs local  #####
# Adding Binary Resident column
#
Ning <- Ning %>%
  mutate(Resident = ifelse(str_detect(nDaysInArea, "Resident"), 1, 0)) %>%
  relocate(Resident, .before = nDaysInArea)

names(Ning)
class(Ning$Resident)

##### nDaysinArea ####
# Standardise nDaysinAreaOther  
Ning <- Ning %>% 
  # Turning times into days
  mutate(nDaysInArea_other = ifelse(str_detect(nDaysInArea_other, "Three weeks|3W|three weeks"), "21",
                                    ifelse(str_detect(nDaysInArea_other, "Month|4 week"), "28",
                                           ifelse(str_detect(nDaysInArea_other, "4 months"), "112",
                                                  ifelse(str_detect(nDaysInArea_other, "6 months"), "168",
                                                         ifelse(str_detect(nDaysInArea_other, "8 weeks - semi resident"), "56",
                                                                ifelse(str_detect(nDaysInArea_other, "3 months|3mm"), "84",
                                                                       ifelse(str_detect(nDaysInArea_other, "5 days"), 5,
                                                                              ifelse(str_detect(nDaysInArea_other, "7 Exmouth|7 days in Exmouth "), 7,
                                                                                     ifelse(str_detect(nDaysInArea_other, "4 in ex"), 4,
                                                                                            ifelse(str_detect(nDaysInArea_other, "Maybe|Traveling|visiting"), NA, # NAing misc comments that arnt useful
                                                                                              ifelse(str_detect(nDaysInArea_other, "7 month"), "196", nDaysInArea_other)))))))))))) %>%
  # Dealing with rougue comments 
  mutate(Refusal = ifelse(str_detect(nDaysInArea_other, "Did|Language|Change|dog"), nDaysInArea_other, NA), 
         Party = ifelse(str_detect(nDaysInArea_other, "Family"), "2a2c", Party), 
         PrevInter = ifelse(str_detect(nDaysInArea_other, "Family"), "Yes", PrevInter),
         Site = ifelse(str_detect(nDaysInArea_other, "Mildura"), "MilduraWreck", Site),
         nDaysInArea_other = ifelse(str_detect(nDaysInArea_other, "Family|Mil|DUNES|Did|Language|Change|dog"), NA, nDaysInArea_other),
         nDaysInArea = ifelse(str_detect(nDaysInArea_other, "112|168|56|84|21|25|28|5|7|4|196|30|90|23"), nDaysInArea_other, nDaysInArea)) %>%
  select(-c(nDaysInArea_other))
# Each of the ifelse statements were checked after each run, this was done by checking which row number in nDaysInArea_other associated with comment changed, and was checked to make sure it matched the corresponing row in new col eg. which(Ning$nDaysInArea_other == "2a2c") == which(Ning$Party == "2a2c")

#### BoatAccess #### 
Ning <- Ning %>% 
  mutate(BoatAccess = ifelse(Year == "2015", "Yes", BoatAccess),
         BoatAccess = ifelse(Year == "2016", "Yes", BoatAccess))

#### nBoatDays ####
unique(Ning$nBoatDays)

Ning <- Ning %>%
  mutate(nBoatDays = ifelse(nBoatDays == "All", nDaysInArea, nBoatDays))


#### nBoatDays_other ####
unique(Ning$nBoatDays)
unique(Ning$nBoatDays_other)

# Explore
Ning %>%
  select(ID, Resident, nBoatDays_other, nBoatDays, nDaysInArea) %>%
  mutate(nBoatDays_other = ifelse(nBoatDays_other == "", NA, nBoatDays_other)) %>%
  filter(nBoatDays_other == ""|!is.na(nBoatDays_other)) %>% 
  group_by(nBoatDays_other) %>%
  arrange(Resident)

# Add ResBoatAvid - got to be applicable only when resident = 1 make NA
# Deal with visitor nBoatDays
#
# Ning <- Ning %>%
 #  mutate(ResBoatAvidYr = ifelse(str_detect(nBoatDays_other, "1 formiye|Every couple of weeks|For|"), "26",
    #                             ifelse(str_detect(nBoatDays_other, ""), T, F)))
# Manual method
# unique(Ning$nBoatDays_other)
# which(Ning$nBoatDays_other == "Couple weekly")
# which(colnames(Ning) == "Resident")

# 

```

# DP RUM Prep
```{r DP_RUM}

meta <- read.csv("rumIgnore/Data-042021/meta_21.csv")
use <- read.csv("rumIgnore/Data-042021/use_21.csv")
avidity <- read.csv("rumIgnore/Data-042021/avidity_21.csv")
jon <- read.csv("rumIgnore/Data-jon/ALL3TRIPSMASTERDATASETFEB22.csv")

is.data.frame(meta) # checking object is df -> TRUE
is.data.frame(use) # checking object is df -> TRUE
is.data.frame(avidity) # checking object is df -> TRUE
is.data.frame(jon)

MetaUse<- merge(meta, use, by.x = "GlobalID", 
                by.y = "Guid", all.x = TRUE, all.y = TRUE) # merge Meta.csv (x) and Use.csv (y) dfs by GlobalID and GUID column
Ning0421 <- merge(MetaUse, avidity, by.x = "GlobalID.y", 
                   by.y = "Guid", all.x = TRUE, all.y = TRUE) # merge MetaUse0421.csv (x) and Avidity.csv (y) dfs by GlobalUse and GUIDavid column

write.csv(Ning0421, 'rumIgnore/Ning0421.csv') 

#### Bind ####
# changing name of date variable so it doesn't not append to Jon's Date variable which is in different format

Ning0421 <- Ning0421 %>% 
  rename(nDate = Date) 

DPrum <- bind_rows(Ning0421, jon) %>%  # Adding on Jons data

#### Select ####
  select(nDate, CreationDate.x, REQUIRED..Boat.ramp.transect.name, Site,
         SCREEN..have.we.interviewed.you.in.the.last.2.weeks.,
         How.many.days.are.you.in.the.area., If.OTHER....days.in.area,
         Do.you.have.access.to.a.private.boat.while.your.here., If.YES....days.you.intend.to.spend.on.boat.,
         If.OTHER....days.on.private.boat., X..separate.times.in.area.in.last.12.months.,
         If.OTHER....times.in.last.12.months,
         X..separate.times.in.area.in.2019., If.OTHER....times.in.area.in.2019, x.y, y.y, EXTRACTIVE.SURVEY,
         If.RECALL..date, What.type.of.fishing.did.you.do.at.this.site., If.OTHER..fishing.type,
         Did.you.use.a.bait.or.lure., What.time.did.you..if.spearfishing...your.lines.get.in.the.water.,
         What.time.did.you..if.spearfishing..your.lines.get.out.the.water.., Max.depth.of.hook.at.this.site.,
         X..fish.you.kept.undamaged.at.this.site., If.OTHER....kept.undamaged, X..released.undamaged.at.this.site.,
         If.OTHER....released.undamaged, X..depredated, If.OTHER....depredated, What.species.did.you.catch.,
         Why.did.you.leave.this.site., If.OTHER..why.did.you.leave.this.site.,
         X..times.fished.from.boat.in.last.12.months., X..years.fishing., What.year.were.you.born., Home.postcode, 
         If.VISITOR..where.are.you.staying., If.OTHER..where.are.you.staying, FINISH.SURVEY, Male.or.Female,
         Describe.party, Boat.length, Boat.Type, Boat.name.or.number, Date, Boat.ramp, Boat.ramp.lat, Boat.ramp.long,
         Lat, Long, Distance.of..fishing.location..from.boat.ramp..miles.,
         Distance.of..fishing.location..from.boat.ramp..nautical.miles.,
         Distance.of..fishing.location..from.boat.ramp..km., No..fish.caught.undamaged, No..fish.lost.to.sharks, 
         Total.fish.hooked, Percentage.depredation, Proportional.depredation...between.0.1.,
         Binary.depredation.presence.absence, No..fish.caught..undamaged.per.hour, No..fish.lost.per.hour,
         Total.fish.hooked.per.hour, Depredation.rate.....per.hour, Time.lines.in, Time.lines.out, Median.time, 
         Decimal.Median.time, Fishing.hours, Decimal.fishing.hours, Fishing.time..rounded.to.nearest.whole.number, 
         Fishing.method, Bait.type, Max.hook..depth, Min.hook..depth, Boat.name.number, Boat.Length, Time.of..interview,
         Have_you_been_interviewed_about, How_many_times_have_you_fished_, 
         How.many.times.fished.from.this.boat.ramp.categorical..1...0.25.times..2...26.50..3...51.75..4...76.100..5.....100.,
         How_many_years_have_you_been_fi, How_many_days_have_you_boat._fished_f, 
         How.many.days.have.you.boat.fished.in.the.last.year.categorical..1...0.25..2...26.50..3...51...75..4...76...100..5.) %>%
         
#### Rename ####
# format: rename(newname = currentname)
  rename(BR = "REQUIRED..Boat.ramp.transect.name",
         PrevInter = "SCREEN..have.we.interviewed.you.in.the.last.2.weeks.", 
         nDaysInArea = "How.many.days.are.you.in.the.area.",
         nDaysInArea_other = "If.OTHER....days.in.area",
         BoatAccess = "Do.you.have.access.to.a.private.boat.while.your.here.",
         nBoatDays = "If.YES....days.you.intend.to.spend.on.boat.",
         nBoatDays_other = "If.OTHER....days.on.private.boat.",
         nTimesLast12m = "X..separate.times.in.area.in.last.12.months.",
         nTimesLast12m_other = "If.OTHER....times.in.last.12.months",
         nTimes19 = "X..separate.times.in.area.in.2019.",
         nTimes19_other = "If.OTHER....times.in.area.in.2019",
         EXTRACTIVE = "EXTRACTIVE.SURVEY",
         exRecall = "If.RECALL..date",
         FishingType = "What.type.of.fishing.did.you.do.at.this.site.",
         FishingType_other = "If.OTHER..fishing.type",
         BaitLure = "Did.you.use.a.bait.or.lure.",
         exStart = "What.time.did.you..if.spearfishing...your.lines.get.in.the.water.",
         exStop = "What.time.did.you..if.spearfishing..your.lines.get.out.the.water..",
         MaxHook = "Max.depth.of.hook.at.this.site.",
         KeptUndam = "X..fish.you.kept.undamaged.at.this.site.",
         KeptUdam_other = "If.OTHER....kept.undamaged",
         RelUndam = "X..released.undamaged.at.this.site.",
         RelUndam_other = "If.OTHER....released.undamaged",
         nDP = "X..depredated",
         nDP_other = "If.OTHER....depredated",
         Species = "What.species.did.you.catch.",
         exWhyLeave = "Why.did.you.leave.this.site.",
         exWhyLeave_other = "If.OTHER..why.did.you.leave.this.site.",
         UseLong = "x.y",
         UseLat = "y.y",
         exnTimes12m = "X..times.fished.from.boat.in.last.12.months.",
         exYrs = "X..years.fishing.",
         YrBorn = "What.year.were.you.born.",
         Postcode = "Home.postcode",
         Accom = "If.VISITOR..where.are.you.staying.",
         Accom_other = "If.OTHER..where.are.you.staying",
         Sex = "Male.or.Female",
         Party = "Describe.party",
         BoatLength = "Boat.length",
         BoatType = "Boat.Type",
         BoatID = "Boat.name.or.number",
         jDate = "Date",
         jBR = "Boat.ramp",
         BRLat = "Boat.ramp.lat",
         BRLong = "Boat.ramp.long",
         jUseLat = "Lat",
         jUseLong = "Long",
         DistBRm = "Distance.of..fishing.location..from.boat.ramp..miles.",
         DistBRnm = "Distance.of..fishing.location..from.boat.ramp..nautical.miles.",
         DistBRkm = "Distance.of..fishing.location..from.boat.ramp..km.",
         jCaughtUndam = "No..fish.caught.undamaged",
         jnDP = "No..fish.lost.to.sharks",
         nHooked = "Total.fish.hooked",
         perDP = "Percentage.depredation",
         proDP = "Proportional.depredation...between.0.1.",
         binDP = "Binary.depredation.presence.absence",
         nUndamHr = "No..fish.caught..undamaged.per.hour",
         nDPHr = "No..fish.lost.per.hour",
         nHookedHr = "Total.fish.hooked.per.hour",
         DPRateHr = "Depredation.rate.....per.hour",
         jexStart = "Time.lines.in",
         jexStop = "Time.lines.out",
         MedianTime = "Median.time",
         DecMedianTime = "Decimal.Median.time",
         FishingHr = "Fishing.hours",
         DecFishingHr = "Decimal.fishing.hours",
         FishTime = "Fishing.time..rounded.to.nearest.whole.number",
         jFishingType = "Fishing.method",
         jBaitLure = "Bait.type",
         jMaxHook = "Max.hook..depth",
         jBoatID = "Boat.name.number",
         jBoatLength = "Boat.Length",
         jTime = "Time.of..interview",
         nTimesFished = "How_many_times_have_you_fished_",
         jexYrs = "How_many_years_have_you_been_fi",
         jDaysBoatFished = "How_many_days_have_you_boat._fished_f",
         jDaysBoatFishedCat = "How.many.days.have.you.boat.fished.in.the.last.year.categorical..1...0.25..2...26.50..3...51...75..4...76...100..5.",
         jTimesFishThisBR = "How.many.times.fished.from.this.boat.ramp.categorical..1...0.25.times..2...26.50..3...51.75..4...76.100..5.....100.",
         jPrevInter = "Have_you_been_interviewed_about") %>% 
  
 #### Filter ####
  filter_all(all_vars(!grepl('ummy', .))) %>% # getting rid of Dummys/dummys
  filter_all(all_vars(!grepl("lah", .))) %>% # getting rid of blah blahs
  filter(is.na(BR) | BR != "BeadonCreek") %>% # Getting rid of Onslow data

# how to get rid of blank observations

#### Checkpoint ####
# Checking df is as expected
# tbl_df(DPrum)
# names(DPrum)

##### Date  and Year ####
  mutate(Date = ifelse(nDate %in% "", CreationDate.x, nDate)) %>% # Appending Creation date to blanks
  separate(Date, into = c("Date", "Time", "ampm"), sep = " ") %>% # Separate datetime var for later
  mutate(Date = parse_date_time(nDate, c("mdY IMS p")), # making POSIX
         Date = substr(Date, 1, 10), # separating out Date
         Date = as.Date(Date, "%Y-%m-%d"), # formatting as.Date to standardise format
         jDate = as.Date(jDate, "%d/%m/%Y"), # formatting as.Date to standardise format
         Year = substr(Date, 1, 4), # taking substring to make year col (1,4) chacrter in col date
         jYear = substr(jDate, 1, 4)) %>% 
  filter(is.na(Year) | Year != 2018) %>% # keep NAs and years that dont equal 2018
  mutate(Date = as.character(Date, "%d/%m/%Y"), # format as chacrter to be able to append 
         jDate = as.character(jDate, "%d/%m/%Y"), # without changing to character comes back numeric nonsense
         Date = ifelse(is.na(Date), jDate, Date), # Append date cols
         Date = as.Date(Date, "%d/%m/%Y")) %>% # format as date
  mutate(Year = ifelse(is.na(Year), jYear, Year)) %>% # Appending year data
  relocate(Date, Time, ampm, jTime, Year) %>% # Putting temporal data at front
  select(-c(CreationDate.x, nDate, jYear, jDate)) %>% # de selecting redundant cols
  arrange(Date, Time) %>% 
#### useID ####
  mutate(useID = row_number()) %>% 
  relocate(useID) %>%
#### Appending BR ####
  mutate(BR = ifelse(is.na(BR), jBR, BR), # Appending jBR and BR
         BR = ifelse(str_detect(BR, "Coral"), "Coral Bay", # Standardising format
                     ifelse(str_detect(BR, "Exmouth"), "Exmouth", BR))) %>% # Standardising format
#### Appending PrevInter ####
  mutate(PrevInter = ifelse(Year == "2015", jPrevInter, PrevInter),
         PrevInter = ifelse(Year == "2016", jPrevInter, PrevInter)) %>%
#### Adding resident col ####
  mutate(Resident = ifelse(str_detect(nDaysInArea, "Resident"), 1, 0)) %>%
  relocate(Resident, .before = nDaysInArea) %>%
##### nDaysinArea ####
# Standardise nDaysinAreaOther  
  # Turning times into days
  mutate(nDaysInArea_other = ifelse(str_detect(nDaysInArea_other, "Three weeks|3W|three weeks"), "21",
                                    ifelse(str_detect(nDaysInArea_other, "Month|4 week"), "28",
                                           ifelse(str_detect(nDaysInArea_other, "4 months"), "112",
                                                  ifelse(str_detect(nDaysInArea_other, "6 months"), "168",
                                                         ifelse(str_detect(nDaysInArea_other, "8 weeks - semi resident"), "56",
                                                                ifelse(str_detect(nDaysInArea_other, "3 months|3mm"), "84",
                                                                       ifelse(str_detect(nDaysInArea_other, "5 days"), 5,
                                                                              ifelse(str_detect(nDaysInArea_other, "7 Exmouth|7 days in Exmouth "), 7,
                                                                                     ifelse(str_detect(nDaysInArea_other, "4 in ex"), 4,
                                                                                            ifelse(str_detect(nDaysInArea_other, "Maybe|Traveling|visiting"), NA, # NAing misc comments that arnt useful
                                                                                              ifelse(str_detect(nDaysInArea_other, "7 month"), "196", nDaysInArea_other)))))))))))) %>%
  # Dealing with rougue comments 
  mutate(Refusal = ifelse(str_detect(nDaysInArea_other, "Did|Language|Change|dog"), nDaysInArea_other, NA), 
         Party = ifelse(str_detect(nDaysInArea_other, "Family"), "2a2c", Party), 
         PrevInter = ifelse(str_detect(nDaysInArea_other, "Family"), "Yes", PrevInter),
         Site = ifelse(str_detect(nDaysInArea_other, "Mildura"), "MilduraWreck", Site),
         nDaysInArea_other = ifelse(str_detect(nDaysInArea_other, "Family|Mil|DUNES|Did|Language|Change|dog"), NA, nDaysInArea_other),
         nDaysInArea = ifelse(str_detect(nDaysInArea_other, "112|168|56|84|21|25|28|5|7|4|196|30|90|23"), nDaysInArea_other, nDaysInArea)) %>%
  select(-c(nDaysInArea_other)) %>%
# Each of the ifelse statements were checked after each run, this was done by checking which row number in nDaysInArea_other associated with comment changed, and was checked to make sure it matched the corresponing row in new col eg. which(Ning$nDaysInArea_other == "2a2c") == which(Ning$Party == "2a2c")

#### BoatAccess #### 
  mutate(BoatAccess = ifelse(Year == "2015", "Yes", BoatAccess),
         BoatAccess = ifelse(Year == "2016", "Yes", BoatAccess)) %>%
#### nBoatDays ####
  mutate(nBoatDays = ifelse(nBoatDays == "All", nDaysInArea, nBoatDays)) %>%
#### nBoatDays_other ####
# Explore
# DPrum %>%
#  select(ID, Resident, nBoatDays_other, nBoatDays, nDaysInArea) %>%
#  mutate(nBoatDays_other = ifelse(nBoatDays_other == "", NA, nBoatDays_other)) %>%
#  filter(nBoatDays_other == ""|!is.na(nBoatDays_other)) %>% 
#  group_by(nBoatDays_other) %>%
#  arrange(Resident)
# Add ResBoatAvid - got to be applicable only when resident = 1 make NA
# Deal with visitor nBoatDays
# Ning <- Ning %>%
#   mutate(ResBoatAvidYr = ifelse(str_detect(nBoatDays_other, "1 formiye|Every couple of weeks|For|"), "26",
#                                 ifelse(str_detect(nBoatDays_other, ""), T, F)))
# Manual method
# unique(Ning$nBoatDays_other)
# which(Ning$nBoatDays_other == "Couple weekly")
# which(colnames(Ning) == "Resident")

#### nTimesLast12m ####
  mutate(nTimesLast12m_other = ifelse(str_detect(nTimesLast12m_other, "1|July"), "1",
                                      ifelse(str_detect(nTimesLast12m_other, "0|Zero|5-6|o"), NA, nTimesLast12m_other)),
         Comments = ifelse(str_detect(nTimesLast12m_other, "dog|house"), nTimesLast12m_other, NA),
         nTimesLast12m_other = ifelse(str_detect(nTimesLast12m_other, "dog|house"), NA, nTimesLast12m_other),
         nTimesLast12m = ifelse(nTimesLast12m_other == "1", "1", nTimesLast12m)) %>%
  select(-nTimesLast12m_other) %>%
#### nTimes19 #### 
  mutate(nTimes19_other = ifelse(str_detect(nTimes19_other, "1|Once"), "1",
                                 ifelse(str_detect(nTimes19_other, "Can’t|H|She’d|For|ould|Might"), NA,
                                        ifelse(str_detect(nTimes19_other, "Was living here"), "Resident", nTimes19_other))),
         Resident = ifelse(nTimes19_other == "Resident", "1", Resident),
         nTimes19_other = ifelse(str_detect(nTimes19_other, "Resident"), NA, nTimes19_other),
         nTimes19 = ifelse(nTimes19_other == "1", "1", nTimes19)) %>%
  select(-nTimes19_other) %>%
  #### Append useLat/Long ####
 # mutate(UseLong = ifelse(Year == "2015", jUseLong, UseLong),
  #       UseLong = ifelse(Year == "2016", jUseLong, UseLong),
   #      UseLat = ifelse(Year == "2015", jUseLat, UseLat),
    #     UseLat = ifelse(Year == "2016", jUseLat, UseLat)) %>%
  # select(-c(jUseLong, jUseLat)) %>%
#### EXTRACTIVE ####
# Every entry which had a location name had coordinates checked to se if they matched or not, if not new coord added
  mutate(UseLong = ifelse(str_detect(EXTRACTIVE, "ooper|squidding"), "114.2212", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Artificial reef|King"), "114.187083", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Marion"),"114.359287", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Exmouth reef"), "114.3667", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Five finger reef"), "113.751633", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "10k of bay"), "114.1126", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "South tanatabiddid"), "113.901809", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Ned’s"), "113.915051", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Fishing off Exmouth boat ramp|harbour"), "114.139613", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "trolled"), "113.919484", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "100m line"), "113.923947", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Helby bank|west side?"), "114.027586", UseLong),
         UseLong = ifelse(str_detect(EXTRACTIVE, "Long Island"), "114.674036", UseLong),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Peak island"), "114.509023", UseLat)) %>% 
    
  mutate(UseLat = ifelse(str_detect(EXTRACTIVE, "ooper|squidding"), "-22.0555", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Artificial reef|King"), "-21.9155", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Marion"), "-21.6583", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Exmouth reef"), "-21.8500", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Five finger reef"), "-23.1820", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "10k of bay"), "-22.4413", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "South tanatabiddid"), "-21.9977", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Ned’s"), "-21.9956", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Fishing off Exmouth boat ramp|harbour"), "-21.9559", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "trolled"), "-21.9193", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "100m line"), "21.8559", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Helby bank|west side?"), "-21.8100", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Long Island"), "-21.6144", UseLat),
         UseLat = ifelse(str_detect(EXTRACTIVE, "Peak island"), "-21.6039", UseLat)) %>%
  
  mutate(EXTRACTIVE = ifelse(str_detect(EXTRACTIVE, "ooper|squidding|Artificial reef|King|Marion|Exmouth reef|Five finger reef|10k of bay|South tanatabiddid|Ned’s|Fishing off Exmouth boat ramp|harbour|trolled|100m line|Helby bank|west side?|Long Island|Peak island|Trolling|trawled|Don’t fish|interviewed|school|Janz|inlet|Bro|I|22k|Larked|tang"), NA, EXTRACTIVE)) %>%
  mutate(YrBorn = ifelse(str_detect(EXTRACTIVE, "1982"), "1982", YrBorn),
         Postcode = ifelse(str_detect(EXTRACTIVE, "6166"), "6166", Postcode),
         Comments = ifelse(str_detect(EXTRACTIVE, "70%depredated"), "0%depredated - gone up, less fish", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Doesn’t fish deeper than 16m because of sharks"), "Doesn’t fish deeper than 16m because of sharks", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "6 spots"), "6 spots between tantabiddi br to light house outside reef", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Targeting marlin"), "Targeting marlin", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Oil"), "Oil rigs", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "They tag and release sharks"), "They tag and release sharks", Comments),
         Comments = ifelse(str_detect(EXTRACTIVE, "Couldn’t be blthard telling me about all the sites but they went to 5 sites and left every time because of sharks"), "Couldn’t be blthard telling me about all the sites but they went to 5 sites and left every time because of sharks", Comments),
         BaitLure = ifelse(str_detect(EXTRACTIVE, "bait"), "Both", BaitLure),
         BoatID = ifelse(str_detect(EXTRACTIVE, "BO 343"), "BO 343", BoatID)) %>%
  select(-EXTRACTIVE) %>%
#### exRecall ####
# DPrum <- DPrum %>% 
#  separate(exRecall, into = c("exDate", "b", "c"), sep = " ") %>%# Separate datetime var for later
#  mutate(exDate = parse_date_time(exDate, c("mdY IMS p")), # making POSIX
#         exDate = substr(exDate, 1, 10), # separating out Date
#         exDate = as.Date(exDate, "%Y-%m-%d"), # formatting as.Date to standardise format
#         exDate = as.character(exDate, "%d/%m/%Y"), # format as chacrter to be able to append 
#         Date = as.character(Date, "%Y/%m/%d"), # without changing to character comes back numeric nonsense
#         Date = ifelse(str_detect(exDate, "201"), exDate, Date), # Append date cols
#         Date = as.Date(Date, "%d/%m/%Y")) # format as date%>% # Appending year data

#### Fishing Type ####
  mutate(FishingType_other = ifelse(str_detect(FishingType_other, "we’re just squidding|Going for squid|Squidding"), "Squidding", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "ottom|Drift|Demersal"), "Demersal", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Free|Spearing"), "Spearing", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Casting|Lining"), "Casting", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Trawling|Trolling|For. Arlin|Marlin|game"), "Trolling", FishingType_other),
         FishingType_other = ifelse(str_detect(FishingType_other, "Prawning"), "Prawning", FishingType_other),
         BaitLure = ifelse(str_detect(FishingType_other, "Both|Bith"), "Both", BaitLure),
         BaitLure = ifelse(str_detect(FishingType_other, "Bait"), "Bait", BaitLure),
         MaxHook = ifelse(str_detect(FishingType_other, "80"), "80", FishingType_other),
         UseLong = ifelse(str_detect(FishingType_other, "fly"), "114.555110", UseLong),
         UseLat = ifelse(str_detect(FishingType_other, "fly"), "-21.809301", UseLat),
         exnTimes12m = ifelse(str_detect(FishingType_other, "25 times a year"), "25", exnTimes12m),
         FishingType_other = ifelse(str_detect(FishingType_other, "Both|Bith|Bait|80|fly|25|0ff|From"), NA, FishingType_other),
         FishingType = ifelse(str_detect(FishingType_other, "Squidding"), "Squidding", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Demersal"), "Demersal", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Spearing"), "Spearing", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Casting"), "Casting", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Trolling"), "Trolling", FishingType),
         FishingType = ifelse(str_detect(FishingType_other, "Prawning"), "Prawning", FishingType),
         FishingType_other = ifelse(str_detect(FishingType_other, "Squidding|Demersal|Spear|Casting|Trolling|Prawning"), NA, FishingType_other))
#### exStart ####
# Standardise and append to jexStart
#### exStop ####
# Standardise and append to jexStop
#### MaxHook ####
# Standardise and append to jMaxHook
#### KeptUndam ####
# Standardise and append other
DPrum <- DPrum %>%
  mutate(Species = ifelse(str_detect(KeptUdam_other, "uid"), "Squid", Species),
         Species = ifelse(str_detect(KeptUdam_other, "1 spangled emperor; 1 threadfin "), "Spangled1/thredfin1", Species),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "Caught|catch"), NA, 
                                 ifelse(str_detect(KeptUdam_other, "2|spangled"), "2",
                                        ifelse(str_detect(KeptUdam_other, "4"), "4",
                                               ifelse(str_detect(KeptUdam_other, "1 squid|Flatty"), "1",
                                                      ifelse(str_detect(KeptUdam_other, "litres"), "Ten litres between two people", KeptUdam_other))))),
         KeptUndam = ifelse(str_detect(KeptUdam_other, "2|5|4|6|1|3"), KeptUdam_other, KeptUndam),
         KeptUdam_other = ifelse(str_detect(KeptUdam_other, "2|5|4|6|1|3"), NA, KeptUdam_other),
         KeptUndam = as.integer(KeptUndam)) %>%
#### RelUndam ####
# Standardise and append other
 mutate(RelUndam_other = ifelse(str_detect(RelUndam_other, "Rel|Too"), "1", RelUndam_other),
        RelUndam = ifelse(str_detect(RelUndam_other, "0|1|8|4|15"), RelUndam_other, RelUndam),
        RelUndam_other = ifelse(str_detect(RelUndam_other, "0|1|8|4|15"), NA, RelUndam_other),
        RelUndam = as.integer(RelUndam)) %>%
#### jCaughtUndam ####
# Append Rel Undam + Kept Undam to 2020/21 values
  mutate(jCaughtUndam = ifelse(Year =="2020", RelUndam + KeptUndam, jCaughtUndam),
         jCaughtUndam = ifelse(Year =="2021", RelUndam + KeptUndam, jCaughtUndam)) %>%
  rename(CaughtUndam = jCaughtUndam) %>%
#### nDP ####
# standardise and append to other and jnDP
  mutate(nDP_other = ifelse(str_detect(nDP_other, "mate"), "1", nDP_other),
         nDP = ifelse(str_detect(nDP_other, "1"), nDP_other, nDP),
         nDP = ifelse(Year =="2015", jnDP, nDP),
         nDP = ifelse(Year == "2016", jnDP, nDP),
         nDP = as.integer(nDP)) %>%
  select(-jnDP) %>%
#### nHooked ####
# If Year = 2020/21, then = RelUndam + nDP + KeptUndam
 mutate(nHooked = ifelse(Year == "2020", KeptUndam + RelUndam + nDP, nHooked),
        nHooked = ifelse(Year == "2021", KeptUndam + RelUndam + nDP, nHooked)) %>%
#### perDP ####
# if Year = 2020/21, then = nDP/nHooked
 mutate(perDP = ifelse(Year == "2020", nDP/nHooked, perDP),
        perDP = ifelse(Year == "2021", nDP/nHooked, perDP)) %>%

#### Append useLat/Long ####
  mutate(UseLong = ifelse(Year == "2015", jUseLong, UseLong),
         UseLong = ifelse(Year == "2016", jUseLong, UseLong),
         UseLat = ifelse(Year == "2015", jUseLat, UseLat),
         UseLat = ifelse(Year == "2016", jUseLat, UseLat)) %>%
  select(-c(jUseLong, jUseLat)) 


##### Spatial Plots ####
w.dir <- "/Users/23088313/Documents/Spatial"
s.dir <- paste(w.dir, "shp", sep='/')
coast <- readOGR(paste(s.dir, "waCoast.shp",sep='/'))

coast <- readOGR(paste(s.dir, "/Users/23088313/Documents/Spatial/waCoast.shp"))

plot(coast)

crs1 <- proj4string(gb)
str(gb)
levels(gb$ZoneName)

```